### 2025-03-12 index.qmd
First commit

# README

Gender in authors in music psychology

What is the gender distribution is music psychology specialist journals?

The public-facing version of the repository is available at
[https://tuomaseerola.github.io/gender_in_music_psych/](https://tuomaseerola.github.io/gender_in_music_psych/)

## Data

Data comes from this source and ....

## Preprocessing

To attribute the gender for the authors, we used gender API [gender_attribution.qmd](gender_attribution.qmd).

## Analysis

Data analysis is covered in [analysis/analysis.qmd](analysis/analysis.qmd) document.

## Manuscript

The study report is available [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.

### 2025-03-12 manuscript/manuscript.qmd
Added MS

---
title             : "Gender Distribution of Authors in Music Psychology"
shorttitle        : "Gender in Music Psychology"
author: 
  - name          : "Tuomas Eerola"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Palace Green, Department of Music, Durham University, DH1 3RL Durham, United Kingdom"
    email         : "tuomas.eerola@durham.ac.uk"
  - name          : "Anna Czepiel"
    affiliation   : "2"
    address       : "Department of Psychology, University of Toronto, Mississauga, Canada"
    email         : "a.czepiel@utoronto.ca"
affiliation:
  - id            : "1"
    institution   : "Department of Music, Durham University"
  - id            : "2"
    institution   : "Department of Psychology, University of Toronto"

author_note: |
  Tuomas Eerola is at the Department of Music, Durham University, UK.
  Anna Czepiel is at the Department of Psychology, University of Toronto, Canada.

abstract: |
 blah blah blah.  
keywords          : "gender, transparency, music psychology, open science, meta-research"
wordcount         : "5555"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csl: apa7.csl
class             : "man"
bibliography: references.bib
format:
  html: default
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Introduction

Equality – Why Should We Care?

- Lack of equity impacts

- productivity
- innovation
- job satisfaction

Academia has acknowledged challenges in gender equity

- women less present in positions of power (e.g. 30% UK profs)
- pay gap is wider in academia (16%) than in other sectors (9%)
- women represent fewer than 30% authorships globally

@son2022scientific Scientific authorship by gender

@shah2021gender Gender parity in scientific authorship in a National Institute for Health Research Biomedical Research Centre: a bibliometric analysis

@rock2021quantifying Quantifying the Gender Gap in Authorship in Herpetology

@ni2021gendered The gendered nature of authorship

See @gonzalez2020women for a baseline from other disciplines:

- Hard Sciences (N=119,592)  0.148
- Bio and Social Sciences (N=262,122)   0.433
- Psychology (N=90,067)     0.452

What is the situation in music psychology?

## Aims

# Methods

## Materials and analyses

- Dataset description. All papers between 2010 and 2022 from specialist journals used in [@jakubowski2024].

- Initial attribution of genders based on first names. Automatic attribution of gender based on first name databases (Gender API) + manual fixes

- Quality control.

- Manual corrections.

# Results

- Overall proportion: First authors

- First authors prevalence by Country (Figure)

- Association with the topics (Figure?)

- Time trends (Figure)

# Conclusions

Encouraging

Compare with sub-disciplines of psychology (Alvarez 2017) Psychology: First authors (N=33,631)

Female role models (past and present)

Area that appeals to women?

Not matched by musicology: 25% female profs, 38% below???

### Funding statement

???

### Competing interests statement

There were no competing interests.

### Open practices statement

Study data, analysis scripts and supporting information is available at GitHub, <https://tuomaseerola.github.io/gender_in_music_psych>.

# References

### 2025-03-19 analysis/process.qmd
Added process.qmd and updated some minor issues in the MS

# Analyse gender in music psychology

T. Eerola, 19/3/2025

Gender analysis of the WEIRD data (Jakubowski et al., 2025).

## Required libraries

```{r}
#| id: libraries
#| warning: false
#| eval: true
#| echo: false
#| message: false
library(tidyverse)
library(stringr)
library(ggplot2)
library(ggwordcloud)
library(modelsummary)
```

## Load and preprocess data

```{r}
#| id: preprocess
#| eval: true

full_process <- FALSE
if(full_process==TRUE){
    d <- read.csv('../data/WEIRD part cleaned Feb24.csv',header = TRUE)
    source('scripts/preprocess.R')      # Routine to process the WEIRD dataset
    source('scripts/gender_analysis.R') # creates enriched dataset
}
#### Get processed data --------------------
if(full_process==FALSE){
    d <- readRDS('../data/enriched_dataset.rds')
}

D <- dplyr::filter(d,study_id=='study1') # 1360 articles
print(paste("number of unique studies:",nrow(D)))
```

## Deal with other authors (non-first authors)

This needs to be done.

1. The author first names for all authors (except for the 1st author) are not in the WEIRD data frame, so these needs to be retrieved.

2. Then the gender attribution process needs to be done with some manual checking.

3. Aggregate with the present (first author) data

## Summarise

This is now only first author data.

```{r}
#| id: summarise
#| eval: true
#| output: asis
source('scripts/summarise_gender.R') #
```

## Visualise

This is now only first author data.

```{r}
#| id: visualise
#| eval: true
source('scripts/visualise_gender.R') # creates a figure
source('scripts/visualise_gender_trends.R') # creates a figure
source('scripts/visualise_names.R')  # creates a figure
```

### 2025-03-19 manuscript/manuscript.qmd
Added process.qmd and updated some minor issues in the MS

@@
 
-- Lack of equity impacts
+- Lack of equity impacts:
 
@@
 
-Academia has acknowledged challenges in gender equity
+Academia has acknowledged serious challenges in gender equity
 
@@
 
-@son2022scientific Scientific authorship by gender
+Weave these studies into the narrative:
 
-@shah2021gender Gender parity in scientific authorship in a National Institute for Health Research Biomedical Research Centre: a bibliometric analysis
+- @son2022scientific Scientific authorship by gender
 
-@rock2021quantifying Quantifying the Gender Gap in Authorship in Herpetology
+- @shah2021gender showed how far off is the gender parity in scientific authorship using a bibliometric analysis
 
-@ni2021gendered The gendered nature of authorship
+- @rock2021quantifying quantified the gender gap in authorship in herpetology, showing that ....
+
+- @ni2021gendered presented an overview of the gendered nature of authorship
 
@@
 
-- Hard Sciences (N=119,592)  0.148
-- Bio and Social Sciences (N=262,122)   0.433
-- Psychology (N=90,067)     0.452
+- Hard Sciences (N=119,592) = 0.148
+- Bio and Social Sciences (N=262,122) =  0.433
+- Psychology (N=90,067) =  0.452
 
@@
 ## Aims
+
+Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
+What is the proportion of first authors in the published papers in the last 10 years?
+What is the situation with all authors and senior authors?
+Are there specific trends in terms of countries of the affiliations, topics, or time?
 
@@
 
+- Expansion into all other first names (requires getting more data and repeating the process above).
+
 # Results
@@
 
-Encouraging
+- Encouraging overall?
 
-Compare with sub-disciplines of psychology (Alvarez 2017) Psychology: First authors (N=33,631)
+- Compare with sub-disciplines of psychology [@gonzalez2020women] Psychology: First authors (N=33,631)
 
-Female role models (past and present)
+- Female role models (past and present) exist in music psychology (list names and their eminent roles, chairs of societies, editors of the main journals, etc.
+Diana Deutsch, Carol Krumhansl, Irene Deliege, to a more recent ICMPC and ESCOM presidents).
 
-Area that appeals to women?
+- Is this an area that appeals to women? (see comparison to sub-disciplines of psychology, where "softer" areas tend to have higher proportion)
 
-Not matched by musicology: 25% female profs, 38% below???
+- Not matched by musicology: 25% female profs, 38% below??? (these are just some UK stats, I don't think the field-wide analysis has been done)
 

### 2025-03-19 analysis/scripts/gender_analysis.R

@@ -1,0 +1,156 @@
+# gender_analysis.R
+# Filename: visualise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: Complete
+
+#### 1. attribute gender -------------------------------
+cat("Attribute gender \n")
+#remotes::install_github("lmullen/genderdata")
+#install.packages("gender",repos = "http://cran.us.r-project.org")
+
+library(gender)
+#g<-gender(x, method = "ssa")
+#d<-read.csv('WEIRD part cleaned Feb24.csv',header=TRUE)
+
+d$firstn <- d$FirstAuthorName_cleaned
+#head(d$firstn)
+d$firstn <- tolower(d$firstn)
+d$firstn <- str_replace_all(d$firstn,'-',' ')
+d$firstn <- str_replace_all(d$firstn,'^[a-z ]+, ','')
+d$firstn <- str_replace_all(d$firstn,' .*$','')
+#table(d$firstn)
+
+U <- sort(unique(d$firstn))
+g <- gender(U, method = "ssa")
+g <- data.frame(g)
+
+## which names are missing?
+
+ind <- U %in% g$name
+missing<-U[!ind]
+missing
+
+# fixes
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Arthurs')]<-'yuko'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Williamson')]<-'victoria'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Annemieke')] <- 'Annemieke'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Scalas')] <- 'Francesca'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Dowling')] <- 'Jay'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Windsor')] <- 'Luke'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Beaman')] <- 'Philip'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Nicholson')] <- 'Riley'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Springer')] <- 'Gregory'
+d$firstn[str_detect(d$FirstAuthorName_cleaned,'Windsor')] <- 'Luke'
[…]

### 2025-03-19 analysis/scripts/get_data_from_scopus.R

@@ -1,0 +1,10 @@
+# get_data_from_scopus.R
+# Filename: visualise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+
+# In order to get the full author list
+
+# we have done this already ELSEWHERE (transparency project!)

### 2025-03-19 analysis/scripts/preprocess.R

@@ -1,0 +1,589 @@
+# preprocess.R
+# WEIRD article
+# T. Eerola, 23/3/2024
+# Status: Requires more work
+
+cat("Preprocessing.....")
+
+cat(paste('\nNumber of coded studies:',nrow(d)))
+cat(paste('\nUnique first authors:',length(unique(d$FirstAuthorName_cleaned))))
+t<-table(d$FirstAuthorName_cleaned)
+t<-dplyr::arrange(data.frame(t),-Freq)
+#head(t,15)
+
+
+#### 1. Create Human studies index -------------
+cat("\n1. Studies with samples (human studies)")
+
+d$humansample <- FALSE
+d$humansample[!is.na(d$SampleSize)  ] <- TRUE
+sum(d$humansample)
+
+ind1<-which(is.na(d$SampleSize) & !is.na(d$SampleOtherDescription))
+ind2 <- which(is.na(d$SampleSize) & !is.na(d$SamplingMethodDescription))
+ind3 <- which(is.na(d$SampleSize) & !is.na(d$SampleMusicianshipDescription))
+ind4 <- unique(c(ind1, ind2, ind3))
+d[ind4,]
+
+d$humansample[ind4]<-TRUE #sum is now 1532
+sum(d$humansample) # indeed
+
+#### 2. Create index for studies with music: 1047, MusicSource column (not NA) -------
+cat("\n2. Studies with music (music studies)")
+d$musicstudies<-FALSE
+d$musicstudies[!is.na(d$MusicSource)]<-TRUE
+sum(d$musicstudies)
+
+#### 3. Create paper id from author+year --------
+cat("\n3. Create paper_ids \n")
+
+tmp <- str_replace(d$FirstAuthorName_cleaned,',.*','')
+tmp <- str_replace(tmp,' ','')
+tmp <- str_replace(tmp,'-.*','')
+tmp <- str_replace(tmp,'_.*','')
+tmp <- str_replace(tmp,' .*','')
+tmp <- str_replace(tmp,'\\..*','')
+tmp <- paste0(tolower(stringi::stri_trans_general(tmp,"latin-ascii")),d$Year,tolower(stringr::str_extract(stringi::stri_trans_general(d$PaperTitle,"latin-ascii"), "^.{2}")))
+d$paper_id <-tmp
+
[…]

### 2025-03-19 analysis/scripts/summarise_gender.R

@@ -1,0 +1,45 @@
+# Filename: summarise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+#### 1. Check for errors ------
+cat("Check for errors \n")
+
+S<-dplyr::summarise(group_by(D),n=n(),female = sum(gender=='female',na.rm = T))
+S$prop <- S$female/S$n
+S$male <- S$n - S$female
+S<-dplyr::arrange(S,-n)
+
+D<-ungroup(D)
+tmp <- D %>%
+  select(firstn,gender) %>%
+  filter(gender=='female') %>%
+  drop_na() %>%
+  group_by(firstn) %>%
+  summarise(n = n())
+tmp<-dplyr::arrange(tmp,-n)
+print(knitr::kable(head(tmp,5),caption='Top 5 names.'))
+
+tmp<-dplyr::arrange(tmp,n)
+print(knitr::kable(head(tmp,5),caption='Bottom 5 names.'))
+
+
+#### 2. Typical stats --------
+
+#% of female authors in this field
+print(datasummary(gender + 1 ~ N + Percent(),
+            data = D))
+#### 3. Number of co-authors -----------
+
+
+##### 4. Summaries ------------
+cat("Summaries \n")
+
+#install.packages("modelsummary",repos = "http://cran.us.r-project.org")
+
+#% of female authors from Non-West
+print(datasummary(FirstAuthorCountry_WEOG * gender + 1 ~ N + Percent(),
+            data = D))
+datasummary(FirstAuthorCountry_USD_Md * gender + 1 ~ N + Percent(),data = D)
+datasummary(gender * FirstAuthorCountry_USD_Md + 1 ~ N + Percent(),data = D)

### 2025-03-19 analysis/scripts/visualise_gender.R

@@ -1,0 +1,207 @@
+# Filename: visualise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+
+#D <- dplyr::filter(d,study_id=='study1') # 1360 articles (1360 correct!)
+#dim(D)
+
+#install.packages("x",repos='http://cran.us.r-project.org', dependencies=TRUE)
+#devtools::install_github("nsgrantham/ggdark")
+
+#### 1. Top countries ------------------
+cat("Top countries \n")
+
+#% female Country specific 
+S<-dplyr::summarise(group_by(D,FirstAuthorCountry),n=n(),female = sum(gender=='female'))
+S$prop <- S$female/S$n
+S$male <- - S$n - S$female
+head(S)
+SS<-dplyr::arrange(S,-n)
+head(SS,20)
+library(ggplot2)
+g1<-ggplot(data = SS[1:10,], aes(x = FirstAuthorCountry,y=prop-.5)) + 
+  geom_col() + 
+  coord_flip() +
+  theme(text = element_text(size=16)) +
+  scale_y_continuous(breaks=seq(-.5,.5,.1),labels=seq(-.5,.5,.10)+.5,limits = c(-.25,.25)) + 
+  scale_fill_brewer(palette="Dark2")+
+  theme_linedraw()
+print(g1)
+
+# applied to my data
+D2<-dplyr::filter(D,FirstAuthorCountry=='US' | 
+                    FirstAuthorCountry=='UK' |
+                    FirstAuthorCountry=='Australia' |
+                    FirstAuthorCountry=='Germany' |
+                    FirstAuthorCountry=='Canada' |
+                    FirstAuthorCountry=='Finland' |
+                    FirstAuthorCountry=='Israel' |
+                    FirstAuthorCountry=='Spain' |
+                    FirstAuthorCountry=='Italy' |
+                    FirstAuthorCountry=='France')
+D2$FirstAuthorCountry<-factor(D2$FirstAuthorCountry,levels = rev(c("US","UK","Australia","Germany","Canada","Finland","Israel","Spain","Italy","France")))
+
+g2<-ggplot(data=D2,aes(x=FirstAuthorCountry,fill=gender)) + 
[…]

### 2025-03-19 analysis/scripts/visualise_gender_trends.R

@@ -1,0 +1,23 @@
+# Filename: visualise_gender_trends.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+tmp <- D %>%
+  select(Year,gender) %>%
+  drop_na() %>%
+  group_by(Year) %>%
+  summarise(Gender = sum(gender=='female'), n = n(),prop=Gender/n)
+head(tmp)
+
+g1<-ggplot(tmp,aes(Year,prop))+
+  geom_smooth(method = lm,formula = y ~ splines::bs(x, 4),se=FALSE,color='lightblue2') +
+  geom_point(size=3)+
+  #  scale_y_continuous(breaks = seq(0,1,by=0.25),limits = c(0.3,1))+
+  scale_y_continuous(limits=c(0.40,.65),breaks = seq(0,1,by=.1),expand = c(0.005,0.005),labels = seq(0,1,by=.1)*100)+
+  scale_x_continuous(breaks = seq(2010,2022,by=2),expand = c(0.005,0.2))+
+  ylab('% Female')+
+  theme_bw(base_size = 15)
+
+print(g1)
+

### 2025-03-19 analysis/scripts/visualise_names.R

@@ -1,0 +1,51 @@
+# Filename: visualise_names.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+plotflag <- TRUE
+
+D$firstn <- tolower(D$firstn)
+D$firstn<-str_to_title(D$firstn)
+
+men <- D %>%
+  filter(gender=='male') %>%
+  group_by(firstn) %>%
+  summarise(n=n())
+
+men<-dplyr::arrange(men,-n)
+men$gender<-'male'
+head(men)
+
+which(men$firstn=='Petri')
+
+women <- D %>%
+  filter(gender=='female') %>%
+  group_by(firstn) %>%
+  summarise(n=n())
+
+women<-dplyr::arrange(women,-n)
+women$gender<-'female'
+head(women)
+top<-30
+both<-rbind(men[1:top,],women[1:top,])
+head(both,30)
+
+
+if(plotflag==TRUE){
+  library(ggwordcloud)
+  set.seed(42)
+
+  g1 <-ggplot(both, aes(label = firstn,size=n,color=gender)) +
+    geom_text_wordcloud(area_corr = TRUE,shape='square') +
+    scale_size_area(max_size = 42) +
+    facet_wrap(.~gender)+
+    scale_radius(range = c(8, 38), limits = c(0, NA)) +
+    scale_color_brewer(palette="Set2")+
+    theme_bw()
+  print(g1)
+
+
+}
+
+#rm(D,data)

### 2025-03-19 data/gonzalez.R

@@ -1,0 +1,7 @@
+# Gonzalez et al 2009
+# TABLE 2
+Location <- c("US","UK","Germany","Canada","Netherlands","Australia","Spain","Italy","France","Belgium","Switzerland","Japan","Sweden","Israel","Brazil","Finland","China","Norway","New Zealand","Austria","Turkey","Portugal","South Africa","Others")
+prop <- c(46.5,43.6,36.5,48.0,41.5,49.1,45.0,52.8,46.1,34.7,35.9,29.8,42.3,49.1,56.1,55.8,36.8,37.8,44.2,42.2,56.9,47.9,50.2,46.1)
+n <- c(35980,6280,4726,4406,3205,2866,2566,1529,1500,1036,1006,793,726,689,610,600,473,473,468,365,348,303,281,3184)
+gonzalez <- data.frame(Country=Location,prop,n=n)
+

### 2025-08-05 analysis.qmd
Updated the analysis and the manuscript and the data

# Analyse gender in music psychology

T. Eerola, 1/8/2025

Gender analysis of the music psychology journals since 2000 (Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science).

## Required libraries

```{r}
#| id: libraries
#| warning: false
#| eval: true
#| echo: false
#| message: false
library(tidyverse)
library(stringr)
library(ggplot2)
library(bib2df)
library(countrycode)
library(modelsummary)
#library(DescTools)
library(easystats)
library(here)
```

## Load and preprocess data

```{r}
#| id: preprocess
#| eval: true
#| output: asis
source('scripts/load_scopus_datasets.R') # Creates full_names dataset
source('scripts/attribute_country.R') # Add country affiliations
source('scripts/attribute_gender.R') # Adds gender attributions (precalculated with API)
source('scripts/create_keys.R') # Add keys for authors and study+author
source('scripts/clean_citations_and_OA.R') # Process citations and OA status
source('scripts/attribute_gender_diagnostics.R') # Process citations and OA status
source('scripts/export_gender_data.R')
```

## Summarise

```{r}
#| id: summarise
#| eval: true
#| output: asis
#| warning: false
#| messages: false
source('scripts/summarise_gender.R') # different summaries
```

## Quantify

```{r}
#| id: quantify
#| eval: true
#| warning: false
#| messages: false
#| output: asis
source('scripts/quantify_authorship.R') # use Odds
source('scripts/citations.R')           # Citations and gender
print(knitr::kable(citestats_all, digits = 2, caption = 'Citations across all authors'))
print(knitr::kable(stats_all))

source('scripts/open_access.R')         # OA and gender
print(knitr::kable(author_OA, digits = 2, caption = 'Open access across all authors'))
```

## Geographical

```{r}
#| id: geographical
#| eval: true
#| output: asis

source('scripts/country_table.R') # OK
source('scripts/visualise_gender.R') # OK

```

```{r}
#| id: geographical2
#| eval: true
#| echo: true
#| output: asis
#source('scripts/table_continent.R') # IN progress
```

Table: Female author odds ratios across continents.

|Continent |    N| Odds_ratio| CI_low| CI_high|
|:---------|----:|----------:|------:|-------:|
|Europe    | 4985|       0.82|   0.75|    0.89|
|Americas  | 2356|       1.12|   1.02|    1.23|
|Asia      |  931|       0.89|   0.78|    1.02|
|Oceania   |  742|       1.66|   1.43|    1.93|
|Africa    |   52|       0.91|   0.52|    1.59|

### 2025-08-05 index.qmd
added placeholders for analyses
Updated the analysis and the manuscript and the data

@@
-# README
+# Gender in music psychology
 
-Gender in authors in music psychology
-
-What is the gender distribution is music psychology specialist journals?
+What is the gender distribution of publishing authors is music psychology?
+Has there been a change over the years in authorship positions (first, coauthor, last) and how is this discipline compared to for instance to psychology?
 
@@
 
-Data comes from this source and ....
-
-## Preprocessing
-
-To attribute the gender for the authors, we used gender API [gender_attribution.qmd](gender_attribution.qmd).
+Data comes from this Scopus with the specific syntax: `SRCTITLE ( JOURNAL NAME ) AND PUBYEAR > 1999`.
+We focussed on 5 highest ranked journals in Music Psychology, namely _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
+To attribute the gender for the authors, we used genderize API from the extracted first names.
+We manually checked a subset of the data and attributed some of those names that the API could not resolve.
 
@@
 
-Data analysis is covered in [analysis/analysis.qmd](analysis/analysis.qmd) document.
+Data analysis is covered in [analysis.qmd](analysis.qmd) document.
 

### 2025-08-05 manuscript/manuscript.qmd
Updated the analysis and the manuscript and the data

@@
 knitr::opts_chunk$set(warning=FALSE, message=FALSE)
+library(here)
+here::i_am("manuscript/manuscript.qmd")
 ```
@@
 # Introduction
+
+In recent years, meta-scientific attention has increasingly turned toward music psychology, with a focus on how WEIRD [@jakubowski2024] and transparent [@eerola2023transparency] the discipline is.
+In this study, we examine gender equality in authorship as an indicator of how equitably gender is represented within the field.
+Disparities in gender distribution may signal structural barriers and disciplinary cultures that constrain the diversity of contributors and hinder optimal knowledge production [@ni2021gendered].
+Authorship visibility is also critical for academic career progression, influencing recognition, funding, and advancement opportunities.
+Establishing a clear understanding of gender representation in music psychology is therefore essential for supporting a more inclusive and sustainable academic community.
+Previous studies across various disciplines—including psychology—have consistently revealed persistent gender inequalities in authorship, though the extent of these disparities varies by field [@son2022scientific; @shah2021gender; @rock2021quantifying; @gonzalez2020women].
 
@@
 
-- @son2022scientific Scientific authorship by gender
+- @ni2021gendered presented an overview of the gendered nature of authorship
 
-- @shah2021gender showed how far off is the gender parity in scientific authorship using a bibliometric analysis
+- @son2022scientific Scientific authorship by gender and they concluded that...
 
-- @rock2021quantifying quantified the gender gap in authorship in herpetology, showing that ....
+- @shah2021gender showed how far off is the gender parity in scientific authorship using a bibliometric analysis.
 
-- @ni2021gendered presented an overview of the gendered nature of authorship
+- @rock2021quantifying quantified the gender gap in authorship in herpetology, showing that .... (we can omit this)
 
@@
 Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
-What is the proportion of first authors in the published papers in the last 10 years?
-What is the situation with all authors and senior authors?
+What is the proportion of different types of authorships (lead, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
 Are there specific trends in terms of countries of the affiliations, topics, or time?
@@
 
-- Dataset description. All papers between 2010 and 2022 from specialist journals used in [@jakubowski2024].
+We retrieved bibliographic information for all articles published between 2000 and June 2025 from five specialist journals, resulting in 3,373 unique articles: Musicae Scientiae (N = 639), Psychology of Music (N = 1,231), Music Perception (N = 675), Journal of New Music Research (N = 563), and Music & Science (N = 265).
+These journals have also been used in previous meta-science studies to characterise research practices in music psychology [@jakubowski2024].
+Author affiliations were extracted automatically and converted into country-level data.
+However, these were not manually verified for each entry, as affiliations are not always clearly matched to individual authors due to variations in reporting conventions, such as multiple or partial affiliations.
 
-- Initial attribution of genders based on first names. Automatic attribution of gender based on first name databases (Gender API) + manual fixes
+In total, the dataset included 9,066 authors, of whom 5,573 were unique.
+Author affiliations spanned 63 countries.
+We also extracted citation counts and Open Access status from Scopus.
+Information on joint first authorship was not available in the data.
 
-- Quality control.
+Gender attribution was initially based on first names.
+We recognise that treating gender as a binary category is inherently problematic.
+Gender is a complex and multidimensional social construct.
+However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
+We used the genderizeR API [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
+This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
+Only 89 names had a low attribution probability (< 0.55).
+Unattributed cases were then checked manually, resulting in 185 manual corrections.
+After this process, 32 names remained ambiguous and 27 were unknown—some likely due to data entry errors in Scopus (e.g.
+only initials or surnames).
+These 59 cases were excluded from the dataset.
+It is likely that the due to challenge of attributing gender correctly not all the gender attributions are correct.
+However, this error has been previously been shown to be non-biased, i.e.
+showing similar quantify of mistakes in both directions [ref].
 
-- Manual corrections.
+We defined three types of authorship positions: first authorship, coauthorship, and last authorship.
+Coauthorship includes all positions other than first and last.
+These positions carry different academic prestige: first authorship is typically associated with the primary contributor, while last authorship is often held by senior researchers with established reputations [@tscharntke2007author].
 
-- Expansion into all other first names (requires getting more data and repeating the process above).
+For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
+Our analysis focused on the female author proportion (FAP) and odds ratios comparing the likelihood of occupying each authorship position by gender.
+
+- Manual corrections (to be done!).
 
@@
 
-- Overall proportion: First authors
+```{r}
+#| id: results
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+source(here("scripts","load_gender_data.R"))
+```
 
-- First authors prevalence by Country (Figure)
+Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),2)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
+To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of occupying different authorship positions.
+We first investigate the authorship types across gender and then examine the author order and number in more detail.
 
-- Association with the topics (Figure?)
+## Authorship positions
 
-- Time trends (Figure)
+The overall odds ratio for women as first authors was 1.32 [95% CI: 1.21–1.44], indicating that the odds of first authorship are 32% higher for females compared to the male authors.
+For coauthorship positions, the OR was 0.99 [0.90–1.08], suggesting no substantial gender difference.
+In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower in females.
+For the first and last authorship position, the odds ratios are significant at _p_ < .001 level.
+
+To illustrate the trends across time in the authorship positions across the gender, we looked at 5-year-windows of the authorship types across gender, shown in Figure 1.
+The higher odds for female authors at first author position seems to be established in 2010 onwards.
+Female first author position exhibits a positive 5-year growth rate of +11.1%.
+The coauthor positions for female authors are relative stable at the odds around 1, indicating relatively equal odds between female and male authors.
+However, the last author positions have become relatively more rare after 2010, showing statistically significantly lower odds ratios from 2010 onwards and negative 5-year growth rate (-9.5%).
+
+{{< include _figure1.qmd >}}
+
+```{r}
+#| id: counts
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+library(dplyr)
+s <- summarise(group_by(df, BIBTEXKEY), n = n())
+df <- mutate(group_by(df, BIBTEXKEY), author_N = max(author_order))
+```
+The range of the number of authors in varies considerable.
+Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`,SD=`{r} round(sd(s$n),2)`).
+To illustrate how the actual number of authors in a paper impacts female first author likelihood, we calculated the odds ratios across the number of authors in paper, pooling the papers with 7 or more authors together (N=`{r} sum(df$author_N>6)`,  `{r} round(sum(df$author_N>6)/nrow(df)*100,1)`%).
+We added the single-authored papers to the plot to indicate the odds of being female to be single authors compared to all studies with multiple authors.
+The results are shown in Figure 2.
+We can observe that while female authors are not more likely to be single authors than men, they tend to be more likely to be first authors than males in papers with 2 or 3 authors; the odds ratios for first authorship position for female authors for papers with 2 authors is 1.55 [130-1.85], and for 3 authors the odds ratio is 1.61 [1.35-1.92], both significantly different from equal proportion.
+The last authorship positions for female authors are consequently lower, odds ratio of 0.65 [0.54–0.77] for 2 authors and 0.59 [0.49–0.71] for 3 authors, both also significant at _p_<.001 level.
+At four and five authors, the higher odds ratio for female is still evident (OR of 1.37 both both), but the last authorship positions are no longer relative rare for papers with 4 and 5 authors (OR 0.90 and 0.95, respectively).
+At six authors and above, the odds ratios indicate the same trend (higher ratio for first author, lower for last) although due to lower number of observations, these are no longer significantly different.
+
+{{< include _figure2.qmd >}}
+
+## Citations and Open Access
+```{r}
+#| id: citations
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+source(here("scripts","citations.R"))
+options(scipen = 9999)
+```
+One potential difference established in previous bibliometric analyses of gendered authorship is citations [@west2013role;@chatterjee2021gender].
+Here we tested whether the citations -- as indexed by Scopus -- show differences across gender.
+The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_=`{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
+The same comparison of citations for studies with last authors yields similar results `{r} citestats_last$Md[1]` [`{r} citestats_last$CI_lower[1]`–`{r} citestats_last$CI_upper[1]`], and for males, the numbers are similar (Md=`{r} citestats_last$Md[2]` [`{r} citestats_last$CI_lower[2]`–`{r} citestats_last$CI_upper[2]`]).
+Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_=`{r} round(stats_last$p.value,3)`).
+If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_=`{r} round(stats_all$p.value,4)`).
+
+```{r}
+#| id: openaccess
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+source(here("scripts","open_access.R"))
+options(scipen = 9999)
+t <- table(df$OA[df$author_order==1])
+```
+
+We also explored whether the open access status of the articles is associated with gender.
+Out of `{r} as.numeric(sum(t))` articles, `{r} round(as.numeric(t[1]/sum(t))*100,2)`% are Open Access in this sample, as indexed by Scopus.
+For first-authored articles, female odds ratio for Open Access is `{r} round(author_OA$Odds_ratio[1],2)` [`{r} round(author_OA$CI_low[1],2)`–`{r} round(author_OA$CI_high[1],2)`], suggesting over 50% higher odds associated with publishing open access by women as compared to men.
+If we observe only the last authorship status of the publications, the difference vanishes with the odds ratio indisguishable from even division (1.00 and the confidence interval), OR = `{r} round(author_OA$Odds_ratio[3],2)` [`{r} round(author_OA$CI_low[3],2)`–`{r} round(author_OA$CI_high[3],2)`].
+
+## Geographical differences
+
+Past studies have identified consistent geographical patterns in female authorship.
+For example, a large-scale analysis of psychology publications found that 46.5% of authorships were by women, with European countries showing a lower average proportion of 42.8% [@gonzalez2020women].
+To examine geographical differences more closely, we calculated the odds of female authorship by country, as shown in Figure 3.
+
+```{r}
+#| id: countrywise
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| output: asis
+#| fig-width: 9
+#| fig-height: 6
+#| fig.cap: > 
+#|   Female author odds ratio across more prolific countries publishing music psychology.
+
+source(here("scripts","country_table.R"))
+```
+
+OPTIONAL
+
+Table: Female author odds ratios across continents.
+
+|Continent |    N| Odds_ratio| CI_low| CI_high|
+|:---------|----:|----------:|------:|-------:|
+|Europe    | 4985|       0.82|   0.75|    0.89|
+|Americas  | 2356|       1.12|   1.02|    1.23|
+|Asia      |  931|       0.89|   0.78|    1.02|
+|Oceania   |  742|       1.66|   1.43|    1.93|
+|Africa    |   52|       0.91|   0.52|    1.59|
+
+## More analyses?
+
+- Association with the keywords (similar)
 
@@
 
-- Encouraging overall?
+- Encouraging overall
 
@@
 
+- Senior authorship positions
+
 ### Funding statement
 
-???
+Authors received no funding for this research.
 

### 2025-08-05 contents.R

@@ -1,0 +1,37 @@
+# contents
+# T. Eerola
+# 01/08/2025
+# Gender analysis of MusicPsych authors
+
+#### LIBRARIES ----------------------------
+library(tidyverse)
+library(stringr)
+library(ggplot2)
+library(bib2df)
+library(countrycode)
+#library(modelsummary)
+library(DescTools)
+library(here)
+
+#### DATA ----------------------------------
+rm(list = ls())
+
+#### GET A NEW DATASET WITH FULL FIRST NAMES --------
+source('scripts/load_scopus_datasets.R') # Creates full_names dataset
+source('scripts/attribute_country.R') # Add country affiliations
+source('scripts/attribute_gender.R') # Adds gender attributions (precalcutaed with API)
+source('scripts/create_keys.R') # Add keys for authors and study+author
+source('scripts/clean_citations_and_OA.R') # Process citations and OA status
+source('scripts/attribute_gender_diagnostics.R') # Calculate how reliable is the attribution
+source('scripts/export_gender_data.R') # save data (remove emails)
+
+#### Analyse --------
+source('scripts/load_gender_data.R') # Loads the data
+source('scripts/summarise_gender.R') # Requires work
+source('scripts/quantify_authorship.R') # Requires work
+source('scripts/visualise_gender.R') # Requires work
+source('scripts/visualise_continent.R') # Requires work
+
+#source('scripts/visualise_gender_trends.R') # OK
+#source('scripts/visualise_names.R') # OK
+source('scripts/keyword_analysis.R') # creates figure (OK)

### 2025-08-05 scripts/attribute_country.R

@@ -1,0 +1,81 @@
+# attribute_country.R
+
+# Attempt attribute country from Affiliation using ISO 3166-1alpha-2 codes
+library(countrycode)
+
+# tactic 1; list all country names first
+#names_of_countries <- countrycode::get_dictionary("exiobase3")
+#save(names_of_countries, file = "data/names_of_countries.rda")
+load(file = "data/names_of_countries.rda")
+
+# fix some country names
+names_of_countries$country.name[names_of_countries$country.name=="Hong Kong SAR China"] <- "Hong Kong"
+names_of_countries$country.name[names_of_countries$country.name=="Czechia"] <- "Czech"
+
+#### patch those countries that are not listed directly but mention city
+source('scripts/missing_countries_in_affiliations.R') # 
+
+# read missing affiliations
+miss <- read.csv("data/missing_affiliations.txt", stringsAsFactors = FALSE,header = TRUE)
+# For manual checking the papers via DOI
+#tmp <- df[df$BIBTEXKEY==miss$BIBTEXKEY[126],]$DOI
+#browseURL(paste0("https://doi.org/",tmp[1]))
+
+delete_studies<- unique(miss$BIBTEXKEY[miss$FIX=="DELETE_ME"])
+df <- dplyr::filter(df, !BIBTEXKEY %in% delete_studies)
+miss <- dplyr::filter(miss, !BIBTEXKEY %in% delete_studies)
+
+for (k in 1:nrow(miss)) {
+  df$AFFILIATIONS[df$BIBTEXKEY==miss$BIBTEXKEY[k]] <- miss$FIX[k]
+}
+
+# attempt to estimate the countries from affiliations, but due to 
+# unattributed / multiple attributions, this is not accurate
+uniq <- unique(df$BIBTEXKEY)
+df$Affiliation_country <- NA
+
+#k<-which(uniq=="Platte2024")
+
+for (k in 1:length(uniq)) {
+#  print(paste0("Processing ", k, " of ", length(uniq), " BIBTEXKEYs: ", uniq[k]))
+  index <- df$BIBTEXKEY==uniq[k]
+  tmp_strings <- stringr::str_split(df$AFFILIATIONS[which(index)[1]], ";")
+  tmp_expanded <- tmp_strings[[1]]
+  tmp_expanded
+  if(!is.na(tmp_expanded[1])) {
+  
+    tmp_names <- NULL
+    for (i in 1:length(tmp_expanded)) {
[…]

### 2025-08-05 scripts/attribute_gender.R

@@ -1,0 +1,71 @@
+# install.packages("genderizeR")
+#suppressPackageStartupMessages(library(genderizeR))
+
+library(modelsummary)
+
+run_genderize <- FALSE
+
+if (run_genderize == TRUE) {
+  # genderize API
+  # Note: Has API limits for free usage
+  my_api_key <- 'xxxx'
+
+  first_names <- data.frame(
+    Name = df$first_name,
+    Country_id = df$Affiliation_country_code
+  )
+  head(first_names)
+  write.csv(first_names, file = "data/first_names.csv", row.names = FALSE)
+
+  # process run in website API because the function was not working in R
+}
+# read the results obtained from the website API
+first_names_resolved <- read.csv('data/first_names_resolved.csv', header = TRUE)
+df$Gender <- first_names_resolved$Gender
+df$Gender.Probability <- first_names_resolved$Gender.Probability
+df$Gender.Count <- first_names_resolved$Gender.Count
+rm(first_names_resolved)
+
+#hist(df$Gender.Probability)
+
+# explore the unknowns
+unattributed <- df$first_name[df$Gender == 'unknown']
+write.csv(
+  unattributed,
+  file = "data/unattributed_first_names.csv",
+  row.names = FALSE
+)
+
+# read data
+inferred <- readr::read_csv(
+  'data/first_names_inferred.csv',
+  show_col_types = FALSE
+)
+#table(inferred$InferredGender)
+
+for (k in 1:nrow(inferred)) {
+  #  print(paste0("Processing ", k, " of ", nrow(inferred), " first names: ", inferred$Name[k]))
+  index <- df$first_name == inferred$Name[k]
+  if (sum(index) > 0) {
+    df$Gender[index] <- inferred$InferredGender[k]
+    df$Gender.Probability[index] <- 0.75
+    df$Gender.Count[index] <- 1
+  }
+}
+
+# Remove books
+df <- dplyr::filter(df, CATEGORY != "BOOK")
+
+# summarise
+print(datasummary(Gender + 1 ~ N + Percent(), data = df))
+
+# Remove unattributed gender
+df <- dplyr::filter(df, Gender != 'ambiguous' & Gender != 'unknown')
+
+cat(paste(
+  "\nEntries in the author-expanded database after filtering books/unattributed:",
+  nrow(df)
+)) 
+
+detach("package:modelsummary", unload=TRUE)
+rm(unattributed, inferred, k, index, run_genderize)

### 2025-08-05 scripts/attribute_gender_diagnostics.R

@@ -1,0 +1,58 @@
+# attribute_gender_diagnostics.R
+library(modelsummary)
+#### Simple diagnostics
+
+cat(paste('\nUnique studies:', length(unique(df$BIBTEXKEY)),'\n'))
+
+cat(paste(
+  "\nUnique Entries as studies and author combinations:",
+  length(unique(df$paper_id_author_id)),'\n'
+))
+
+cat(paste('\n Number of rows:', nrow(df),'\n'))
+
+print(datasummary(JOURNAL + 1 ~ N + Percent(), data = dplyr::filter(df,author_order==1),output='markdown',title='Number of studies across journals'))
+
+cat("\n\n")
+
+# API finds are based on
+#print(median(df$Gender.Count))
+#hist(df$Gender.Count)
+#hist(df$Gender.Probability)
+
+# the majority is confidently attributed
+cat(paste("Prop. over .90 conf. =",sum(df$Gender.Probability>.90)/nrow(df),"/n"))
+cat(paste("Prop. over .95 conf. =",sum(df$Gender.Probability>.95)/nrow(df),"/n"))
+cat(paste("Prop. under .55 conf. =",sum(df$Gender.Probability<.55)/nrow(df),"/n"))
+cat(paste("Prop. under .55 conf. =",sum(df$Gender.Probability<.55),"/n/n"))
+
+cat("\n***")
+
+print(datasummary(Mean + Median ~ Gender.Probability * Gender, data = df,output='markdown',title='Probability of API attributed Gender.'))
+
+cat("\n***")
+
+upperCI<-function(x, conf.level = 0.95) {
+  f <- DescTools::MeanCI(x, conf.level = conf.level)
+  #lowerCI <- f[1]
+  upperCI <- f[3]
+  return(upperCI)
+}
+lowerCI<-function(x, conf.level = 0.95) {
+  f <- DescTools::MeanCI(x, conf.level = conf.level)
+  lowerCI <- f[2]
+  #upperCI <- f[2]
+  return(lowerCI)
+}
+
+print(datasummary(Mean + upperCI + lowerCI ~ Citations * Gender, data = df,output='markdown',title='Citation counts across Gender'))
+
+print(broom::glance(t.test(Citations ~ Gender, data=df)))
+print(broom::glance(wilcox.test(Citations ~ Gender, data=df)))
+
+detach("package:modelsummary", unload=TRUE)
+
+
+#### Gender distribution of authors -----------
[…]

### 2025-08-05 scripts/citations.R

@@ -1,0 +1,56 @@
+# citations.R
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'First'
+df$authortype[df$author_last == 1] <- 'Other'
+#table(df$authortype)
+
+tmp<-dplyr::filter(df,authortype=='First')
+citestats_first <- summarise(group_by(tmp, Gender), Md = median(Citations), M = mean(Citations),Q75=quantile(Citations,0.75))
+stats_first <- broom::glance(kruskal.test(Citations ~ Gender, data=tmp))
+citestats_first$CI_lower<-NULL
+citestats_first$CI_upper<-NULL
+citestats_first$CI_lower[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[2]
+citestats_first$CI_upper[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[3]
+citestats_first$CI_lower[2]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='male'], conf.level = 0.95)[2]
+citestats_first$CI_upper[2]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='male'], conf.level = 0.95)[3]
+
+# classify order: Coauthor
+df$authortype <- 'Coauthor'
+df$authortype[df$author_order == 1] <- 'Other'
+df$authortype[df$author_last == 1] <- 'Other'
+
+tmp <- dplyr::filter(df,authortype=='Coauthor')
+citestats_coauthor <- summarise(group_by(tmp, Gender), Md = median(Citations), M = mean(Citations),Q75=quantile(Citations,0.75))
+stats_coauthor <- broom::glance(kruskal.test(Citations ~ Gender, data=tmp))
+citestats_coauthor$CI_lower<-NULL
+citestats_coauthor$CI_upper<-NULL
+citestats_coauthor$CI_lower[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[2]
+citestats_coauthor$CI_upper[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[3]
+citestats_coauthor$CI_lower[2]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='male'], conf.level = 0.95)[2]
[…]

### 2025-08-05 scripts/clean_citations_and_OA.R

@@ -1,0 +1,20 @@
+# aggregate_citations_and_OA.R
+
+df$OA <- "Undefined"
+df$OA[stringr::str_detect(df$NOTE, ';.*Access')] <- df$NOTE[stringr::str_detect(
+  df$NOTE,
+  ';.*Access'
+)]
+df$OA <- stringr::str_replace_all(df$OA, 'Cite.*; ', '')
+table(df$OA)
+# Here consider all Open Access as OA
+df$OA[df$OA != "Undefined"] <- 'Open Access'
+
+#### Citations ------------------
+
+df$Citations <- df$NOTE
+df$Citations <- stringr::str_replace_all(df$Citations, ';.*$', '')
+df$Citations <- stringr::str_replace_all(df$Citations, 'Cited by: ', '')
+df$Citations <- as.numeric(df$Citations)
+#table(df$Citations)
+#hist(df$Citations)

### 2025-08-05 scripts/create_keys.R

@@ -1,0 +1,27 @@
+# create_keys.R
+
+#### Add Author ids--------
+
+library(dplyr)
+df <- df %>%
+  group_by(BIBTEXKEY) %>%
+  mutate(author_id = paste0("author", row_number()))
+
+df$paper_id_author_id <- paste0(df$BIBTEXKEY, '_', df$author_id)
+
+#### Add indices for the author orders -------
+
+df$author_order <- as.numeric(stringr::str_replace(df$author_id, 'author', ''))
+df$author_order
+df <- mutate(group_by(df, BIBTEXKEY), Max = max(author_order))
+df$author_last <- 0
+df$author_last[df$author_order == df$Max] <- 1 # if the author rank equals max
+df$author_last[df$author_order == 1] <- 0 # except when the rank is 1
+
+df <- dplyr::select(df,-Max) # delete this column, not needed
+
+#### Unique author ids ----------
+df$unique_name <- paste0(str_trim(str_to_lower(df$last_name, locale = "en")),"_", str_trim(str_to_lower(df$first_name, locale = "en")))
+
+cat(paste("\n Unique authors count:",length(unique(df$unique_name))))
+cat(paste("\n Unique countries count:",length(unique(df$Affiliation_country))))

### 2025-08-05 scripts/export_gender_data.R

@@ -1,0 +1,8 @@
+# export_gender_data.R
+run_only_once <- FALSE
+if (run_only_once) {
+  index <- which(sapply(df, is.list))
+  names(df)[index]
+  df <- dplyr::select(df,-EDITOR,-CATEGORY,-CORRESPONDENCE_ADDRESS,-NOTE)
+  write.csv(df,file='data/gender_in_mp_exported.csv', row.names = FALSE)
+}

### 2025-08-05 scripts/load_gender_data.R

@@ -1,0 +1,3 @@
+# load_gender_data.R
+
+df <- read.csv(file=here("data","gender_in_mp_exported.csv"), header = TRUE)

### 2025-08-05 scripts/load_scopus_datasets.R

@@ -1,0 +1,64 @@
+# load_scopus_datasets.R
+
+df1 <- bib2df::bib2df("data/MetS2000_onwards.bib",separate_names = TRUE)
+stopifnot(nrow(df1)==298)
+df2 <- bib2df::bib2df("data/JNMR2000_onwards.bib",separate_names = TRUE)
+stopifnot(nrow(df2)==597)
+df3 <- bib2df::bib2df("data/MP2000_onwards.bib",separate_names = TRUE)
+stopifnot(nrow(df3)==745)
+df4 <- bib2df::bib2df("data/MS2000_onwards.bib",separate_names = TRUE)
+stopifnot(nrow(df4)==682)
+df5 <- bib2df::bib2df("data/POM2000_onwards.bib",separate_names = TRUE)
+stopifnot(nrow(df5)==1604)
+d<-rbind(df1,df2,df3,df4,df5)
+rm(df1,df2,df3,df4,df5)
+
+cat(paste("\nEntries in the merged database:", nrow(d)))
+
+#### Deal with authors, now nested into AUTHOR column 
+library(tidyr)
+library(dplyr)
+
+df <- d %>%
+  unnest(AUTHOR)
+
+head(df)
+cat(paste("\nEntries in the author-expanded database:", nrow(df))) # 9922 authors
+
+# check what type of articles have missing names
+df$TYPE[is.na(df$first_name)]
+
+# filter Erratum
+df <- df %>%
+  filter(!str_detect(TYPE, "Erratum|Corrigendum|Retraction|Announcement|Editorial"))
+cat(paste("\nEntries in the author-expanded database after filtering editorials etc:", nrow(df))) # 9662 authors
+
+# Filter if full name is NA
+df <- df %>%
+  filter(!is.na(full_name))
+cat(paste("\nEntries in the author-expanded database after filtering empty:", nrow(df))) # 9651 authors
+
+# maybe remove those that are just initials
+
+df$AFFILIATIONS <- df$AFFILIATIONS %>%
+  stringr::str_replace_all("Australia, Australia", "Australia") %>%
+  stringr::str_replace_all("Germany, Germany", "Germany") %>%
+  stringr::str_replace_all("Australia,Australia", "Australia") %>%
+  stringr::str_replace_all("Germany,Germany", "Germany") %>%
+  stringr::str_replace_all("United States, Brunswick", "") %>%
+  stringr::str_replace_all("United States, Georgia", "") %>%
[…]

### 2025-08-05 scripts/quantify_authorship.R

@@ -1,0 +1,165 @@
+# quantify_authorship.R
+
+# Use Odds ratios to quantify the probabilities of gendered authorships.
+
+# classify order: First author
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'First'
+df$authortype[df$author_last == 1] <- 'Other'
+t<-table(df$authortype,df$Gender)
+
+or1 <- effectsize::oddsratio(t) # 1.32
+#print(or1)
+
+# classify order: Coauthor
+df$authortype <- 'Coauthor'
+df$authortype[df$author_order == 1] <- 'Other'
+df$authortype[df$author_last == 1] <- 'Other'
+t<-table(df$authortype,df$Gender)
+t
+or2 <- effectsize::oddsratio(t) # 0.99
+#print(or2)
+
+# classify order: Last
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'Other'
+df$authortype[df$author_last == 1] <- 'Last'
+t<-table(df$authortype,df$Gender)
+t
+or3 <- effectsize::oddsratio(t) # 0.73
+
+or <- rbind(or1,or2,or3)
+or$name<-c("First","Coauthor","Last")
+print(knitr::kable(or))
+
+
+#### Annually ---------
+
+# introduce 5 year bins
+df$YEAR_5 <- cut(df$YEAR, breaks=seq(2000, 2025, by=5), labels=c("2000-2004", "2005-2009", "2010-2014", "2015-2019", "2020-2024"), right=FALSE)
+table(df$YEAR,df$YEAR_5)
+table(df$YEAR_5)
+
+# classify order: First author
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'First'
+df$authortype[df$author_last == 1] <- 'Other'
+
+output <- NULL
+U<-levels(df$YEAR_5)
+for (k in 1:length(U)){
+  tmp<-dplyr::filter(df,YEAR_5==U[k])
+  t <- table(tmp$authortype,tmp$Gender)
+  output <- rbind(output,effectsize::oddsratio(t))
+}
+output$YearRange <- U
+first <- data.frame(output)
+first
+# classify order: Coauthor
+df$authortype <- 'Coauthor'
+df$authortype[df$author_order == 1] <- 'Other'
+df$authortype[df$author_last == 1] <- 'Other'
+
+U<-levels(df$YEAR_5)
[…]

### 2025-08-05 scripts/summarise_gender.R

@@ -1,0 +1,38 @@
+# Filename: summarise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+invisible(library(modelsummary))
+
+#### 1. Gender distribution ------
+
+# % of female authors in this field
+print(datasummary(Gender + 1 ~ N + Percent(), data = df,output='markdown'))
+
+tmp <- df %>%
+  select(first_name, Gender) %>%
+  filter(Gender == 'female') %>%
+  drop_na() %>%
+  group_by(first_name) %>%
+  summarise(n = n())
+tmp <- dplyr::arrange(tmp, -n)
+print(knitr::kable(head(tmp, 5), caption = 'Top 5 names.'))
+rm(tmp)
+
+#### 3. Number of co-authors -----------
+#print(datasummary(factor(author_order) + 1 ~ Gender + N + Percent(), data = df))
+
+##### NEW SUMMARIES ------
+# How many author on average
+cat("\nNumber of coauthors:\n")
+library(dplyr)
+s <- summarise(group_by(df, BIBTEXKEY), n = n())
+cat(paste("\n median:",median(s$n)))
+cat(paste("\n mean:",round(mean(s$n),3)))
+cat(paste("\n sd:",round(sd(s$n),3)))
+cat(paste("\n max:",round(max(s$n),3)))
+cat("\n")
+rm(s)
+
+
+detach("package:modelsummary", unload=TRUE)

### 2025-08-05 scripts/visualise_gender.R

@@ -1,0 +1,123 @@
+# Filename: visualise_gender.R
+# Author: T. Eerola, 6/4/2024
+# Project: gender in music psychology
+# Status: in progress
+
+#### 1. Top countries ------------------
+cat("Top countries \n")
+
+#% female Country specific
+S <- dplyr::summarise(
+  group_by(df, Affiliation_country),
+  n = n(),
+  female = sum(Gender == 'female')
+)
+S$prop <- S$female / S$n
+S$male <- -S$n - S$female
+head(S)
+SS <- dplyr::arrange(S, -n)
+head(SS, 20)
+
+library(ggplot2)
+g1 <- ggplot(data = SS[1:10, ], aes(x = Affiliation_country, y = prop - .5)) +
+  geom_col() +
+  coord_flip() +
+  theme(text = element_text(size = 16)) +
+  scale_y_continuous(
+    breaks = seq(-.5, .5, .1),
+    labels = seq(-.5, .5, .10) + .5,
+    limits = c(-.25, .25)
+  ) +
+  scale_fill_brewer(palette = "Dark2") +
+  theme_linedraw()
+print(g1)
+
+# applied to my data
+D2 <- dplyr::filter(
+  df,
+  Affiliation_country == 'US' |
+    Affiliation_country == 'UK' |
+    Affiliation_country == 'Australia' |
+    Affiliation_country == 'Germany' |
+    Affiliation_country == 'Canada' |
+    Affiliation_country == 'Finland' |
+    Affiliation_country == 'Israel' |
+    Affiliation_country == 'Spain' |
+    Affiliation_country == 'Italy' |
+    Affiliation_country == 'France'
+)
+D2$Affiliation_country <- factor(
+  D2$Affiliation_country,
+  levels = rev(c(
+    "US",
+    "UK",
+    "Australia",
+    "Germany",
+    "Canada",
+    "Finland",
+    "Israel",
+    "Spain",
+    "Italy",
+    "France"
+  ))
+)
+
+g2 <- ggplot(data = D2, aes(x = Affiliation_country, fill = Gender)) +
+  geom_bar(data = subset(D2, Gender == "female"), color = 'black') +
+  geom_bar(
+    data = subset(D2, Gender == "male"),
+    aes(y = after_stat(count) * (-1)),
+    color = 'black'
+  ) +
+  # scale_y_continuous(
[…]

### 2025-08-05 index.qmd
Updated the analysis and the manuscript and the data

@@
 Data comes from this Scopus with the specific syntax: `SRCTITLE ( JOURNAL NAME ) AND PUBYEAR > 1999`.
-We focussed on 5 highest ranked journals in Music Psychology, namely _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
-To attribute the gender for the authors, we used genderize API from the extracted first names.
-We manually checked a subset of the data and attributed some of those names that the API could not resolve.
+We focussed on 5 journals in Music Psychology, namely _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
+To attribute the gender for the authors, we used [genderize API](https://genderize.io/).
+The details of the process are given in the manuscript.
 
@@
 
-Data analysis is covered in [analysis.qmd](analysis.qmd) document.
+Data analysis steps are covered in the [analysis.qmd](analysis.qmd) document.
 
@@
 
-The study report is available [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.
+The study report is available at [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.

### 2025-08-07 manuscript/manuscript-apaquarto.qmd
Added a version that compiles MS in pdf

--- 2025-08-05 manuscript/manuscript.qmd
@@
     affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Palace Green, Department of Music, Durham University, DH1 3RL Durham, United Kingdom"
-    email         : "tuomas.eerola@durham.ac.uk"
+    corresponding: true
+    email: tuomas.eerola@durham.ac.uk
+    url: https://tuomaseerola.github.io/
+    roles:
+      - conceptualization
+      - methodology
+      - formal Analysis
+      - writing
+    affiliations:
+      - id: id1
+        name: "Durham University"
+        department: Department of Music
+        address: Palace Green
+        city: Durham
+        country: United Kingdom
+        postal-code: DH1 3DA
   - name          : "Anna Czepiel"
-    affiliation   : "2"
-    address       : "Department of Psychology, University of Toronto, Mississauga, Canada"
-    email         : "a.czepiel@utoronto.ca"
-affiliation:
-  - id            : "1"
-    institution   : "Department of Music, Durham University"
-  - id            : "2"
-    institution   : "Department of Psychology, University of Toronto"
-
-author_note: |
-  Tuomas Eerola is at the Department of Music, Durham University, UK.
-  Anna Czepiel is at the Department of Psychology, University of Toronto, Canada.
-
-abstract: |
- blah blah blah.  
+    roles:
+      - data curation
+      - formal analysis
+      - writing
+    affiliations   :
+      - id: id2
+        name       : "Department of Psychology, University of Toronto"
+        city: Toronto
+        country: Canada
+author_note:
+      data-sharing: "https://tuomaseerola.github.io/gender_in_music_psych"
+abstract: "This study examines the gender distribution of the authors in music psychology."
 keywords          : "gender, transparency, music psychology, open science, meta-research"
-wordcount         : "5555"
-figsintext        : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : no
-mask              : no
-csl: apa7.csl
-class             : "man"
+impact-statement: ~
+floatsintext: true
+# Numbered lines (.pdf and .docx only)
+numbered-lines: false
+# File with references
 bibliography: references.bib
+# Suppress title page
+suppress-title-page: false
+# Link citations to references
+link-citations: true
+# Masks references that appear in the masked-citations list
+mask: false
+# If true, adds today's date below author affiliations. If text, can be any value.
+# This is not standard APA format, but it is convenient.
+# Works with docx, html, and typst. 
+draft-date: false
+# Language options. See https://quarto.org/docs/authoring/language.html
+lang: en
+language:
+  citation-last-author-separator: "and"
+  citation-masked-author: "Masked Citation"
+  citation-masked-date: "n.d."
+  citation-masked-title: "Masked Title"
+  email: "Email"
+  title-block-author-note: "Author Note"
+  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
+  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
+  title-impact-statement: "Impact Statement"
+  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
 format:
-  html: default
+#  apaquarto-docx: default
+#  apaquarto-html: default
+  apaquarto-typst: 
+    keep-typ: true
+    list-of-figures: false
+    list-of-tables: false
 ---
@@
 Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),2)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
-To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of occupying different authorship positions.
-We first investigate the authorship types across gender and then examine the author order and number in more detail.
+To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of females occupying different authorship positions.
+We first investigate the authorship types across gender and then examine the author order and the overall number of coauthors in more detail.
 

### 2025-08-07 analysis/scripts/gender_analysis.R

@@ -1,156 +1,0 @@
-# gender_analysis.R
-# Filename: visualise_gender.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: Complete
-
-#### 1. attribute gender -------------------------------
-cat("Attribute gender \n")
-#remotes::install_github("lmullen/genderdata")
-#install.packages("gender",repos = "http://cran.us.r-project.org")
-
-library(gender)
-#g<-gender(x, method = "ssa")
-#d<-read.csv('WEIRD part cleaned Feb24.csv',header=TRUE)
-
-d$firstn <- d$FirstAuthorName_cleaned
-#head(d$firstn)
-d$firstn <- tolower(d$firstn)
-d$firstn <- str_replace_all(d$firstn,'-',' ')
-d$firstn <- str_replace_all(d$firstn,'^[a-z ]+, ','')
-d$firstn <- str_replace_all(d$firstn,' .*$','')
-#table(d$firstn)
-
-U <- sort(unique(d$firstn))
-g <- gender(U, method = "ssa")
-g <- data.frame(g)
-
-## which names are missing?
-
-ind <- U %in% g$name
-missing<-U[!ind]
-missing
-
-# fixes
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Arthurs')]<-'yuko'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Williamson')]<-'victoria'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Annemieke')] <- 'Annemieke'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Scalas')] <- 'Francesca'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Dowling')] <- 'Jay'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Windsor')] <- 'Luke'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Beaman')] <- 'Philip'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Nicholson')] <- 'Riley'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Springer')] <- 'Gregory'
-d$firstn[str_detect(d$FirstAuthorName_cleaned,'Windsor')] <- 'Luke'
[…]

### 2025-08-07 analysis/scripts/get_data_from_scopus.R

@@ -1,10 +1,0 @@
-# get_data_from_scopus.R
-# Filename: visualise_gender.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: in progress
-
-
-# In order to get the full author list
-
-# we have done this already ELSEWHERE (transparency project!)

### 2025-08-07 analysis/scripts/preprocess.R

@@ -1,589 +1,0 @@
-# preprocess.R
-# WEIRD article
-# T. Eerola, 23/3/2024
-# Status: Requires more work
-
-cat("Preprocessing.....")
-
-cat(paste('\nNumber of coded studies:',nrow(d)))
-cat(paste('\nUnique first authors:',length(unique(d$FirstAuthorName_cleaned))))
-t<-table(d$FirstAuthorName_cleaned)
-t<-dplyr::arrange(data.frame(t),-Freq)
-#head(t,15)
-
-
-#### 1. Create Human studies index -------------
-cat("\n1. Studies with samples (human studies)")
-
-d$humansample <- FALSE
-d$humansample[!is.na(d$SampleSize)  ] <- TRUE
-sum(d$humansample)
-
-ind1<-which(is.na(d$SampleSize) & !is.na(d$SampleOtherDescription))
-ind2 <- which(is.na(d$SampleSize) & !is.na(d$SamplingMethodDescription))
-ind3 <- which(is.na(d$SampleSize) & !is.na(d$SampleMusicianshipDescription))
-ind4 <- unique(c(ind1, ind2, ind3))
-d[ind4,]
-
-d$humansample[ind4]<-TRUE #sum is now 1532
-sum(d$humansample) # indeed
-
-#### 2. Create index for studies with music: 1047, MusicSource column (not NA) -------
-cat("\n2. Studies with music (music studies)")
-d$musicstudies<-FALSE
-d$musicstudies[!is.na(d$MusicSource)]<-TRUE
-sum(d$musicstudies)
-
-#### 3. Create paper id from author+year --------
-cat("\n3. Create paper_ids \n")
-
-tmp <- str_replace(d$FirstAuthorName_cleaned,',.*','')
-tmp <- str_replace(tmp,' ','')
-tmp <- str_replace(tmp,'-.*','')
-tmp <- str_replace(tmp,'_.*','')
-tmp <- str_replace(tmp,' .*','')
-tmp <- str_replace(tmp,'\\..*','')
-tmp <- paste0(tolower(stringi::stri_trans_general(tmp,"latin-ascii")),d$Year,tolower(stringr::str_extract(stringi::stri_trans_general(d$PaperTitle,"latin-ascii"), "^.{2}")))
-d$paper_id <-tmp
-
[…]

### 2025-08-07 analysis/scripts/summarise_gender.R

@@ -1,45 +1,0 @@
-# Filename: summarise_gender.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: in progress
-
-#### 1. Check for errors ------
-cat("Check for errors \n")
-
-S<-dplyr::summarise(group_by(D),n=n(),female = sum(gender=='female',na.rm = T))
-S$prop <- S$female/S$n
-S$male <- S$n - S$female
-S<-dplyr::arrange(S,-n)
-
-D<-ungroup(D)
-tmp <- D %>%
-  select(firstn,gender) %>%
-  filter(gender=='female') %>%
-  drop_na() %>%
-  group_by(firstn) %>%
-  summarise(n = n())
-tmp<-dplyr::arrange(tmp,-n)
-print(knitr::kable(head(tmp,5),caption='Top 5 names.'))
-
-tmp<-dplyr::arrange(tmp,n)
-print(knitr::kable(head(tmp,5),caption='Bottom 5 names.'))
-
-
-#### 2. Typical stats --------
-
-#% of female authors in this field
-print(datasummary(gender + 1 ~ N + Percent(),
-            data = D))
-#### 3. Number of co-authors -----------
-
-
-##### 4. Summaries ------------
-cat("Summaries \n")
-
-#install.packages("modelsummary",repos = "http://cran.us.r-project.org")
-
-#% of female authors from Non-West
-print(datasummary(FirstAuthorCountry_WEOG * gender + 1 ~ N + Percent(),
-            data = D))
-datasummary(FirstAuthorCountry_USD_Md * gender + 1 ~ N + Percent(),data = D)
-datasummary(gender * FirstAuthorCountry_USD_Md + 1 ~ N + Percent(),data = D)

### 2025-08-07 analysis/scripts/visualise_gender.R

@@ -1,207 +1,0 @@
-# Filename: visualise_gender.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: in progress
-
-
-#D <- dplyr::filter(d,study_id=='study1') # 1360 articles (1360 correct!)
-#dim(D)
-
-#install.packages("x",repos='http://cran.us.r-project.org', dependencies=TRUE)
-#devtools::install_github("nsgrantham/ggdark")
-
-#### 1. Top countries ------------------
-cat("Top countries \n")
-
-#% female Country specific 
-S<-dplyr::summarise(group_by(D,FirstAuthorCountry),n=n(),female = sum(gender=='female'))
-S$prop <- S$female/S$n
-S$male <- - S$n - S$female
-head(S)
-SS<-dplyr::arrange(S,-n)
-head(SS,20)
-library(ggplot2)
-g1<-ggplot(data = SS[1:10,], aes(x = FirstAuthorCountry,y=prop-.5)) + 
-  geom_col() + 
-  coord_flip() +
-  theme(text = element_text(size=16)) +
-  scale_y_continuous(breaks=seq(-.5,.5,.1),labels=seq(-.5,.5,.10)+.5,limits = c(-.25,.25)) + 
-  scale_fill_brewer(palette="Dark2")+
-  theme_linedraw()
-print(g1)
-
-# applied to my data
-D2<-dplyr::filter(D,FirstAuthorCountry=='US' | 
-                    FirstAuthorCountry=='UK' |
-                    FirstAuthorCountry=='Australia' |
-                    FirstAuthorCountry=='Germany' |
-                    FirstAuthorCountry=='Canada' |
-                    FirstAuthorCountry=='Finland' |
-                    FirstAuthorCountry=='Israel' |
-                    FirstAuthorCountry=='Spain' |
-                    FirstAuthorCountry=='Italy' |
-                    FirstAuthorCountry=='France')
-D2$FirstAuthorCountry<-factor(D2$FirstAuthorCountry,levels = rev(c("US","UK","Australia","Germany","Canada","Finland","Israel","Spain","Italy","France")))
-
-g2<-ggplot(data=D2,aes(x=FirstAuthorCountry,fill=gender)) + 
[…]

### 2025-08-07 analysis/scripts/visualise_gender_trends.R

@@ -1,23 +1,0 @@
-# Filename: visualise_gender_trends.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: in progress
-
-tmp <- D %>%
-  select(Year,gender) %>%
-  drop_na() %>%
-  group_by(Year) %>%
-  summarise(Gender = sum(gender=='female'), n = n(),prop=Gender/n)
-head(tmp)
-
-g1<-ggplot(tmp,aes(Year,prop))+
-  geom_smooth(method = lm,formula = y ~ splines::bs(x, 4),se=FALSE,color='lightblue2') +
-  geom_point(size=3)+
-  #  scale_y_continuous(breaks = seq(0,1,by=0.25),limits = c(0.3,1))+
-  scale_y_continuous(limits=c(0.40,.65),breaks = seq(0,1,by=.1),expand = c(0.005,0.005),labels = seq(0,1,by=.1)*100)+
-  scale_x_continuous(breaks = seq(2010,2022,by=2),expand = c(0.005,0.2))+
-  ylab('% Female')+
-  theme_bw(base_size = 15)
-
-print(g1)
-

### 2025-08-07 analysis/scripts/visualise_names.R

@@ -1,51 +1,0 @@
-# Filename: visualise_names.R
-# Author: T. Eerola, 6/4/2024
-# Project: gender in music psychology
-# Status: in progress
-
-plotflag <- TRUE
-
-D$firstn <- tolower(D$firstn)
-D$firstn<-str_to_title(D$firstn)
-
-men <- D %>%
-  filter(gender=='male') %>%
-  group_by(firstn) %>%
-  summarise(n=n())
-
-men<-dplyr::arrange(men,-n)
-men$gender<-'male'
-head(men)
-
-which(men$firstn=='Petri')
-
-women <- D %>%
-  filter(gender=='female') %>%
-  group_by(firstn) %>%
-  summarise(n=n())
-
-women<-dplyr::arrange(women,-n)
-women$gender<-'female'
-head(women)
-top<-30
-both<-rbind(men[1:top,],women[1:top,])
-head(both,30)
-
-
-if(plotflag==TRUE){
-  library(ggwordcloud)
-  set.seed(42)
-
-  g1 <-ggplot(both, aes(label = firstn,size=n,color=gender)) +
-    geom_text_wordcloud(area_corr = TRUE,shape='square') +
-    scale_size_area(max_size = 42) +
-    facet_wrap(.~gender)+
-    scale_radius(range = c(8, 38), limits = c(0, NA)) +
-    scale_color_brewer(palette="Set2")+
-    theme_bw()
-  print(g1)
-
-
-}
-
-#rm(D,data)

### 2025-08-08 manuscript/manuscript.qmd
fixed minor things

@@
 
-abstract: |
- blah blah blah.  
+abstract: "This study examines the gender distribution of the authors in music psychology. N=XXXX published papers between years 2000 to 2025 from 5 journals were ..."  
 keywords          : "gender, transparency, music psychology, open science, meta-research"
@@
 
+The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.
+While overall women hold smaller number of authorship overall (xx%),
+
 - Encouraging overall

### 2025-08-08 manuscript/manuscript.qmd
Introduction added a segment

@@
 Previous studies across various disciplines—including psychology—have consistently revealed persistent gender inequalities in authorship, though the extent of these disparities varies by field [@son2022scientific; @shah2021gender; @rock2021quantifying; @gonzalez2020women].
+
+Academia suffers from issues regarding equity.
+A lack equity impacts productivity, innovation, and job satisfaction in the workplace, which hugely impact academic progress.
+One aspect of this inequity is that gender.
+Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power, a pay gap remains between men and women in academia (16%) which is wider than other sectors (9%).
+In a “publish or perish” culture of science and academia (Kiai, 2019), one crucial aspect to focus on is authorship.
+The number and quality (e.g., journal reputation) of authorships an individual determines crucial career development such as grant funding and future (permanent) positions.
 

### 2025-08-08 manuscript/manuscript.qmd
add another paragraph to introduction

@@
 The number and quality (e.g., journal reputation) of authorships an individual determines crucial career development such as grant funding and future (permanent) positions.
+
+Research across several scientific disciplines that show that women tend to be less represented in authors compared to men (Banks et al., 2025; @son2022scientific].
+Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving.
+From around 1960 to 2021, studies show a general improvement [@gonzalez2020women]; Ioannadis et al., 2023; Jemielniak & Wilamowski, 2025).
+However, there is evidence that this improvement is plateauing (Jemielniak & Wilamowski, 2025) and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender].
+There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorships can be unclear.
+What counts as an authorship may also differ between academic disciplines.
+Women are more likely to experience authorship disagreements (Ni et al., 2021).
+There are general disagreements of when to decide authorship - women preferring to discuss in earlier stages, but men choosing authorship at the final stage.
+Another reason could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement (Neilsen et al 2023).
+Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit (Banks et al., 2025) as well as creating equitable environments in academic science (Martinez Acosta et al., 2025).
 

### 2025-08-11 contents.R
Checking the gender added

@@ -20,6 +20,7 @@ rm(list = ls())
 source('scripts/load_scopus_datasets.R') # Creates full_names dataset
 source('scripts/attribute_country.R') # Add country affiliations
 source('scripts/attribute_gender.R') # Adds gender attributions (precalcutaed with API)
+source('scripts/name_gender_fixes.R') # Fixes the gender attribution of some names manually 
 source('scripts/create_keys.R') # Add keys for authors and study+author
 source('scripts/clean_citations_and_OA.R') # Process citations and OA status
 source('scripts/attribute_gender_diagnostics.R') # Calculate how reliable is the attribution

### 2025-08-11 data/gonzalez.R

@@ -5,3 +5,39 @@ prop <- c(46.5,43.6,36.5,48.0,41.5,49.1,45.0,52.8,46.1,34.7,35.9,29.8,42.3,49.1,
 n <- c(35980,6280,4726,4406,3205,2866,2566,1529,1500,1036,1006,793,726,689,610,600,473,473,468,365,348,303,281,3184)
 gonzalez <- data.frame(Country=Location,prop,n=n)
 
+gonzalez$male<- round((100 - gonzalez$prop)/100 * gonzalez$n)
+gonzalez$female<- round((gonzalez$prop)/100 * gonzalez$n)
+head(gonzalez)
+
+total_male <- sum(gonzalez$male)
+total_female <- sum(gonzalez$female)
+total_female
+#### First position
+# if males were in the first position, from p. 5
+male <- 23128 
+female <- 8804
+t <- matrix(NA,2,2)
+t[1,1]<-female; t[1,2] <- male
+t[2,1]<-total_female-female; t[2,2]<- total_male-male
+colnames(t)<-c("female","male")
+rownames(t)<-c("first","other")
+t
+or <- effectsize::oddsratio(t) # 0.73
+or
+chisq.test(t,simulate.p.value = FALSE)
+
+# if the papers were signed by a women the first position, from p. 5
+male<- 11473 
+female <- 20589
+t <- matrix(NA,2,2)
+t[1,1]<-female; t[1,2] <- male
+t[2,1]<-total_female-female; t[2,2]<- total_male-male
+colnames(t)<-c("female","male")
+rownames(t)<-c("first","other")
+t
+or <- effectsize::oddsratio(t) # 0.73
+or
+chisq.test(t,simulate.p.value = FALSE)
+
+
+

### 2025-08-11 scripts/attribute_gender_diagnostics.R

@@ -56,3 +56,7 @@ detach("package:modelsummary", unload=TRUE)
 #### Gender distribution of authors -----------
 dist1 <- DescTools::MultinomCI(as.numeric(table(df$Gender)),conf.level=0.95,method="sisonglaz")
 print(knitr::kable(dist1,digits=3,caption='Gender Distribution'))
+
+
+
+

### 2025-08-11 scripts/create_keys.R

@@ -18,10 +18,29 @@ df$author_last <- 0
 df$author_last[df$author_order == df$Max] <- 1 # if the author rank equals max
 df$author_last[df$author_order == 1] <- 0 # except when the rank is 1
 
+#### Check all author roles -----------
+table(df$author_last,df$author_order)
+table(df$author_order==1,df$author_order) # 
+table(df$author_order==1,df$Max) # 
+
+df$author_single <- 0
+df$author_single[df$author_order==1 & df$Max==1] <- 1 # if the author is the only one in the paper
+table(df$author_single)
+table(df$author_last)
+table(df$author_single,df$author_order)
+table(df$author_single,df$author_last)
+
+sum(df$author_single==1 & df$author_last==1)
+
+total_authors <- sum(df$author_single) + sum(df$author_last)
+sum(df$author_order==1)
+
 df <- dplyr::select(df,-Max) # delete this column, not needed
 
 #### Unique author ids ----------
 df$unique_name <- paste0(str_trim(str_to_lower(df$last_name, locale = "en")),"_", str_trim(str_to_lower(df$first_name, locale = "en")))
 
+cat(paste("\n Single authors count:",sum(df$author_single)))
+cat(paste("\n Single + last authors count:",sum(df$author_single)+sum(df$author_last)))
 cat(paste("\n Unique authors count:",length(unique(df$unique_name))))
 cat(paste("\n Unique countries count:",length(unique(df$Affiliation_country))))

### 2025-08-11 scripts/name_gender_fixes.R

@@ -1,0 +1,28 @@
+# Filename: name_gender_fixes.R
+# Author: T. Eerola, 9/8/2025
+# Project: gender in music psychology
+# Status: Complete
+
+#### 1. Manual fixes to gender by authors (TE and AC) --------------
+cat("Manual fixes\n")
+
+# Fix wrongly attributed
+# from the first name only
+df$Gender[df$first_name=='Simha'] <- 'male' # TE
+df$Gender[df$first_name=='Zohar']<-'male'   # TE
+
+# from a full name
+df$Gender[df$full_name=='Ashley D. Vanstone']<-'male' # TE
+df$Gender[df$full_name=='Andrea Schiavio']<-'male'    # TE
+
+#### Diagnostic output ------
+df$composite_name <- paste(df$first_name, df$last_name, sep = " ")
+df %>%
+  group_by(Gender, composite_name) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(Gender,desc(n)) -> output
+
+write_file <- FALSE
+if(write_file==TRUE){
+  write.csv(output,'check_gender.csv',row.names = FALSE)
+}

### 2025-08-11 scripts/quantify_authorship.R

@@ -1,37 +1,50 @@
 # quantify_authorship.R
-
+# T. Eerola
+# 09/08/2025
+# Gender analysis of MusicPsych authors
 # Use Odds ratios to quantify the probabilities of gendered authorships.
 
-# classify order: First author
+#### Classify order: Single author ----------
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'Other'
+df$authortype[df$author_last == 1] <- 'Other'
+df$authortype[df$author_single == 1] <- 'Single' # FOCUS
+t<-table(df$authortype,df$Gender)
+or1 <- effectsize::oddsratio(t) # 1.10
+#print(or1)
+
+#### Classify order: First author ----------
 df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'First'
 df$authortype[df$author_last == 1] <- 'Other'
+df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-
-or1 <- effectsize::oddsratio(t) # 1.32
+or2 <- effectsize::oddsratio(t) # 1.32
 #print(or1)
 
-# classify order: Coauthor
+#### classify order: Coauthor --------------
 df$authortype <- 'Coauthor'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
+df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-t
-or2 <- effectsize::oddsratio(t) # 0.99
+
+or3 <- effectsize::oddsratio(t) # 1.01
 #print(or2)
 
-# classify order: Last
+#### Classify order: Last -------------
 df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Last'
+df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
 t
-or3 <- effectsize::oddsratio(t) # 0.73
+or4 <- effectsize::oddsratio(t) # 0.73
 
-or <- rbind(or1,or2,or3)
-or$name<-c("First","Coauthor","Last")
+or <- rbind(or1,or2,or3,or4)
+or$name<-c("Single","First","Coauthor","Last")
[…]

### 2025-08-19 manuscript/manuscript.qmd
checked female->male names in name_gender_fixes. Added to intro. Figure update authorships by year.

--- 2025-08-08 manuscript/manuscript.qmd
@@
 
+Equality – Why Should We Care?
 Academia suffers from issues regarding equity.
@@
 One aspect of this inequity is that gender.
-Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power, a pay gap remains between men and women in academia (16%) which is wider than other sectors (9%).
-In a “publish or perish” culture of science and academia (Kiai, 2019), one crucial aspect to focus on is authorship.
+Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power (e.g.
+30% UK profs- REF NEEDED?), a pay gap remains between men and women in academia (16%) which is wider than other sectors (9%) (REF also needed).
+In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
 The number and quality (e.g., journal reputation) of authorships an individual determines crucial career development such as grant funding and future (permanent) positions.
 
-Equality – Why Should We Care?
-
-- Lack of equity impacts:
-
-- productivity
-- innovation
-- job satisfaction
-
-Academia has acknowledged serious challenges in gender equity
-
-- women less present in positions of power (e.g. 30% UK profs)
-- pay gap is wider in academia (16%) than in other sectors (9%)
-- women represent fewer than 30% authorships globally
-
-Weave these studies into the narrative:
-
-- @ni2021gendered presented an overview of the gendered nature of authorship
-
-- @son2022scientific Scientific authorship by gender and they concluded that...
-
-- @shah2021gender showed how far off is the gender parity in scientific authorship using a bibliometric analysis.
-
-- @rock2021quantifying quantified the gender gap in authorship in herpetology, showing that .... (we can omit this)
+Research across several scientific disciplines that show that women tend to be less represented in authors compared to men [@banks2025women]; @son2022scientific].
+Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving.
+From around 1960 to 2021, studies show a general improvement [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender].
+However, there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender].
+There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorships can be unclear.
+What counts as an authorship may also differ between academic disciplines.
+Women are more likely to experience authorship disagreements [@ni2021gendered].
+There are general disagreements of when to decide authorship - women preferring to discuss in earlier stages, but men choosing authorship at the final stage.
+Another reason could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
+Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
 
 See @gonzalez2020women for a baseline from other disciplines:
-
 - Hard Sciences (N=119,592) = 0.148
@@
 What is the situation in music psychology?
+To improve certain academic equity issues, it is worthwhile focusing within disciplines and seeing how more local academic communities can provide more equity.
+To do this, it is worth giving a snapshot of specific fields.
+Indeed, instead of looking broadly at journal articles in STEM, or in several different kinds of journals, it can be worthwhile giving an idea of a more specific.
+For example, one paper explores authorship in Herpetology, showing that male authors outnumber female, but female authorship is increasing steadily across the last decade [@rock2021quantifying].
+The current paper also focuses on music psychology as a discipline: within music psychology, there are many efforts to improve the field.
+There have been papers that provide a snapshot of the field to provide a rough overview of the state of affairs (e.g., @anglada2019visualizing) with suggestions of future directions to improve the field.
+[@eerola2023transparency] showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
+@jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is, which the view to see how – as a field – we need to diversify participant samples and stimuli to gain a more realistic understanding.
+In a similar way to these papers, the current paper aims to give a current overview of authorship in the field of music psychology.
 

### 2025-08-19 scripts/name_gender_fixes.R

@@ -15,6 +15,34 @@ df$Gender[df$first_name=='Zohar']<-'male'   # TE
 df$Gender[df$full_name=='Ashley D. Vanstone']<-'male' # TE
 df$Gender[df$full_name=='Andrea Schiavio']<-'male'    # TE
 
+df$Gender[df$full_name=='Andrea Ravignani']<-'male' # AMC
+df$Gender[df$full_name=='Alexis Kirke']   <-'male' # AMC
+df$Gender[df$full_name=='Toni Bechtold']  <-'male' # AMC
+df$Gender[df$full_name=='Gopala Koduri']  <-'male' # AMC
+df$Gender[df$full_name=='Hani Yehia']     <-'male' # AMC
+df$Gender[df$full_name=='Jolan Kegelars'] <-'male' # AMC
+df$Gender[df$full_name=='Li Hu']          <-'male' # AMC
+df$Gender[df$full_name=='N. Ruiz-Reyes']  <-'male' # N. -> Nicolas? AMC
+df$Gender[df$full_name=='Nori Jacoby']    <-'male' #  AMC
+df$Gender[df$full_name=='Archia Cont']    <-'male' #  AMC
+df$Gender[df$full_name=='Ashlin Roy']     <-'male' #  AMC
+df$Gender[df$full_name=='Chuansheng Chen']<-'male' #  AMC
+df$Gender[df$full_name=='Clude Alain']    <-'male' # typo, should be Claude Alain AMC
+df$Gender[df$full_name=='Hao-Min Liu']    <-'male' #  AMC
+df$Gender[df$full_name=='Haruka Shoda']   <-'male' #  AMC
+df$Gender[df$full_name=='Hirokata Fukushima']<-'male' #  AMC
+df$Gender[df$full_name=='Ji Kim']  <-'male' #  AMC
+df$Gender[df$full_name=='Junqi Deng']<-'male' #  AMC
+df$Gender[df$full_name=='Li Su']<-'male' #  AMC
+df$Gender[df$full_name=='Linglan Zhu']<-'male' #  AMC
+df$Gender[df$full_name=='Linh Nghiem']<-'male' #  AMC
+df$Gender[df$full_name=='Nery Borges']<-'male' #  AMC
+df$Gender[df$full_name=='Vesa Välimäki']<-'male' #  AMC
+df$Gender[df$full_name=='Yi-Hsuan Yang']<-'male' #  AMC
[…]

### 2025-08-19 scripts/quantify_authorship.R

@@ -103,15 +103,36 @@ coauthor$Type<-'Co'
 last$Type<-'Last'
 
 annual<-rbind(first,coauthor,last)
+annual$Type = factor(annual$Type, levels = c('First','Co', 'Last'))
 
-fig1 <- ggplot(annual,aes(x=YearRange,y=Odds_ratio,color=Type))+
+# Create a new facet variable that group INCREASE / DECREASE of OR 
+annual <- annual %>%
+  mutate(TypeGroup = case_when(Type %in% c("Co", "Last") ~ "Co and Last",TRUE ~ Type))
+annual$TypeGroup = as.factor(annual$TypeGroup)
+annual$TypeGroup = factor(annual$TypeGroup, levels = c( "First" ,"Co and Last"))
+  
+fig1 <- ggplot(annual,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type))+
+  #facet_wrap(~TypeGroup,nrow = 2)+
   geom_hline(yintercept=1, linetype="dashed", color = "black") +
   geom_point(position = position_dodge(width = 0.4))+
   geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
+  geom_line(alpha = 0.2,position = position_dodge(width = 0.4))+
   scale_color_brewer(palette = "Set1") +
   xlab('Year range')+
   ylab('Female Authorship Odds Ratio')+
   theme_classic()
+
+# fig1 <- ggplot(annual,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type))+
+#   facet_wrap(~TypeGroup,nrow = 2)+
+#   geom_hline(yintercept=1, linetype="dashed", color = "black") +
+#   geom_point(position = position_dodge(width = 0.4))+
+#   geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
+#   geom_line(alpha = 0.2)+
+#   scale_color_brewer(palette = "Set1") +
+#   xlab('Year range')+
+#   ylab('Female Authorship Odds Ratio')+
+#   theme_classic()
+
 print(fig1)
 
 #### Average growth rate ----------

### 2025-08-19 scripts/visualise_gender.R

@@ -79,7 +79,7 @@ g2 <- ggplot(data = D2, aes(x = Affiliation_country, fill = Gender)) +
   xlab('') +
   ylab('N') +
   theme_linedraw(base_size = 16)
-print(g2)
+#print(g2)
 #ggsave("gender_top10_countries_normal.png", g1, bg="transparent")
 
 ### %

### 2025-08-19 manuscript/_figure1.qmd
Revised analyses, added references, additions to MS

```{r}
#| label: fig1
#| fig-width: 9
#| fig-height: 4.75
#| echo: false
#| warning: false
#| message: false
#| fig.cap: > 
#|   Authorship types across time.

source(here("scripts","load_gender_data.R"))
library(ggplot2,quietly = TRUE)
library(RColorBrewer)

#### Classify order: Single author ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Single' # FOCUS
t<-table(df$authortype,df$Gender)
or1 <- effectsize::oddsratio(t) # 1.00
#print(or1)

#### Classify order: First author ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'First'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
t<-table(df$authortype,df$Gender)
or2 <- effectsize::oddsratio(t) # 1.41
#print(or2)

#### Classify order: Coauthor --------------
df$authortype <- 'Coauthor'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
t<-table(df$authortype,df$Gender)
or3 <- effectsize::oddsratio(t) # 1.00
#print(or3)

#### Classify order: Last -------------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Last'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
t<-table(df$authortype,df$Gender)
or4 <- effectsize::oddsratio(t) # 0.73
#print(or4)

or <- rbind(or1,or2,or3,or4)
or$name<-c("Single","First","Coauthor","Last")
or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
#print(knitr::kable(or,digits = 2))

#### Annually ---------

# introduce 5 year bins
df$YEAR_5 <- cut(df$YEAR, breaks=seq(2000, 2025, by=5), labels=c("2000-2004", "2005-2009", "2010-2014", "2015-2019", "2020-2024"), right=FALSE)
#table(df$YEAR,df$YEAR_5)
#table(df$YEAR_5)

#### Classify order: Single author ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Single' # FOCUS

output <- NULL
U<-levels(df$YEAR_5)
for (k in 1:length(U)){
  tmp<-dplyr::filter(df,YEAR_5==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}
output$YearRange <- U
single <- data.frame(output)
#single

#### Classify order: First author ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'First'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!

output <- NULL
U<-levels(df$YEAR_5)
for (k in 1:length(U)){
  tmp<-dplyr::filter(df,YEAR_5==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}
output$YearRange <- U
first <- data.frame(output)
#first

#### Classify order: Coauthor --------------
df$authortype <- 'Coauthor'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!

U<-levels(df$YEAR_5)
output <- NULL
for (k in 1:length(U)){
  tmp<-dplyr::filter(df,YEAR_5==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}
output$YearRange <- U
coauthor <- data.frame(output)
#coauthor

#### Classify order: Last -------------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Last'
df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!

U<-levels(df$YEAR_5)
output <- NULL
for (k in 1:length(U)){
  tmp<-dplyr::filter(df,YEAR_5==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}
output$YearRange <- U
last <- data.frame(output)

single$Type<-'Single'
first$Type<-'First'
coauthor$Type<-'Co'
last$Type<-'Last'

annual<-rbind(single,first,coauthor,last)
annual$Type = factor(annual$Type, levels = c('Single','First','Co', 'Last'))
#annual

# Create a new facet variable that group INCREASE / DECREASE of OR 
#annual <- annual %>%
#  mutate(TypeGroup = case_when(Type %in% c("Co", "Last") ~ "Co and Last",TRUE ~ Type))
#annual$TypeGroup = as.factor(annual$TypeGroup)
#annual$TypeGroup = factor(annual$TypeGroup, levels = c( "First" ,"Co and Last"))

# add overall
or$name = factor(or$name, levels = c('Single','First','Coauthor', 'Last'), labels = c('Single','First','Co', 'Last'))

tmp <- data.frame(Odds_ratio = or$Odds_ratio, CI=0.95,CI_low = or$CI_low, CI_high = or$CI_high, YearRange = "Overall", Type = or$name)
annual2<- rbind(tmp,annual)
annual2$separator<-"Yearly"
annual2$separator[annual2$YearRange=="Overall"]<-"Overall"
annual2$YearRange <- factor(annual2$YearRange, levels = c("Overall", "2000-2004", "2005-2009", "2010-2014", "2015-2019", "2020-2024"))

fig1 <- ggplot(annual2,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator))+
  #facet_wrap(~TypeGroup,nrow = 2)+
  geom_hline(yintercept=1, linetype="dashed", color = "black") +
  geom_point(position = position_dodge(width = 0.4),show.legend=FALSE,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
#  geom_line(alpha = 0.2,position = position_dodge(width = 0.4))+
  scale_shape_manual(values = c(22, 16)) +
  scale_color_brewer(name="Authorship\ntype",palette = "Set1",direction=-1) +
  xlab('Year range')+
  ylab('Female Authorship Odds Ratio (95%CI)')+
  theme_classic(base_size = 15)
print(fig1)

```

### 2025-08-19 manuscript/_figure2.qmd
Revised analyses, added references, additions to MS

```{r}
#| label: fig2
#| fig-width: 9
#| fig-height: 4
#| echo: false
#| warning: false
#| message: false
#| fig.cap: > 
#|   Authorship types across author counts.

source(here("scripts","load_gender_data.R"))
library(ggplot2,quietly = TRUE)
library(RColorBrewer)

df$author_arbitrary <- cut(
  df$author_order,
  include.lowest = TRUE,
  breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8,40),
  labels = c("1", "2", "3", "4", "5", "6", "7", "8","9+")
)

df <- dplyr::mutate(dplyr::group_by(df, BIBTEXKEY), author_N = max(author_order))
#table(df$author_N)

df$author_N_cropped <- cut(
  df$author_N,
  include.lowest = TRUE,
  breaks = c(0, 1, 2, 3, 4, 5, 6, 40),
  labels = c("1", "2", "3", "4", "5", "6","7+")
)
#table(df$author_N_cropped)

#### classify order: First author -------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'First'
df$authortype[df$author_last == 1] <- 'Other'
#table(df$authortype,df$Gender)

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 2:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[2:length(U)]
first <- data.frame(output)
#first

#### classify order: Coauthor ----------
df$authortype <- 'Coauthor'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
#table(df$authortype,df$Gender)

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 3:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[3:length(U)]
coauthor <- data.frame(output)

#### classify order: Last ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Last'

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 2:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
  t <- table(tmp$authortype,tmp$Gender)
  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[2:length(U)]
last <- data.frame(output)

first$Type<-'First'
coauthor$Type<-'Co'
last$Type<-'Last'

author_N<-rbind(first,coauthor,last)
author_N$Type<-factor(author_N$Type,levels=c("First","Co","Last"))

# Can we count this for single-authored papers?
tmp<-dplyr::filter(df,author_N==1)
t<-table(tmp$Gender)
tmp2<-dplyr::filter(df,author_N>1)
t2<-table(tmp2$Gender)
T<-data.frame(rbind(single=t,multi=t2))
single<-data.frame(effectsize::oddsratio(T),author_N_cropped=1,Type="Single")
author_N <- rbind(single,author_N)
author_N$Type<-factor(author_N$Type,levels=c("Single","First","Co","Last"))

custom_palette <- brewer.pal(n = 4, name = "Set1")
custom_palette<-c(custom_palette[4],custom_palette[1:3])
fig2 <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
  geom_hline(yintercept=1, linetype="dashed", color = "black") +
  geom_point(position = position_dodge(width = 0.4),shape=15,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  scale_color_manual(name="Authorship\ntype",values = custom_palette) +
  xlab('Author N')+
  ylab('Female Authorship Odds Ratio')+
  theme_classic(base_size = 15)
print(fig2)
```

### 2025-08-19 manuscript/manuscript.qmd
Revised analyses, added references, additions to MS

@@
 One aspect of this inequity is that gender.
-Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power (e.g.
-30% UK profs- REF NEEDED?), a pay gap remains between men and women in academia (16%) which is wider than other sectors (9%) (REF also needed).
+Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
+For instance, in the UK, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the academia [@harris2025gender].
+A a pay gap remains between men and women in academia (female-male wage ratio 0.85 in 2020) [@quadlin2023higher].
 In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
@@
 However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
-We used the genderizeR API [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
+We used the _genderize API_ [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
 This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
@@
 It is likely that the due to challenge of attributing gender correctly not all the gender attributions are correct.
-However, this error has been previously been shown to be non-biased, i.e.
-showing similar quantify of mistakes in both directions [ref].
+Although genderize.io has been shown to achieve 96.6% accuracy in a diverse multinational test database without using a country of origin information -- 98% with the information included -- [@vanhelene2024inferring], the tool underperforms in Asian names (82% accuracy).
+However, the small error has been previously been shown to be non-biased, i.e.
+showing similar number of mistakes for both genders [@sebo2021performance;@vanhelene2024inferring].
+
+We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full first name coded with initials, or the second name used as the first name.
+This led to `XXXX` corrections.
 
@@
 Our analysis focused on the female author proportion (FAP) and odds ratios comparing the likelihood of occupying each authorship position by gender.
-
-- Manual corrections (to be done!).
 
@@
 The range of the number of authors in varies considerable.
-Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`,SD=`{r} round(sd(s$n),2)`).
+Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`, SD=`{r} round(sd(s$n),2)`).
 To illustrate how the actual number of authors in a paper impacts female first author likelihood, we calculated the odds ratios across the number of authors in paper, pooling the papers with 7 or more authors together (N=`{r} sum(df$author_N>6)`,  `{r} round(sum(df$author_N>6)/nrow(df)*100,1)`%).
@@
 #| fig.cap: > 
-#|   Female author odds ratio across more prolific countries publishing music psychology.
+#|   Female author odds ratio across the prolific countries publishing music psychology (with more than sixty publications).
 

### 2025-08-19 contents.R

@@ -9,7 +9,6 @@ library(stringr)
 library(ggplot2)
 library(bib2df)
 library(countrycode)
-#library(modelsummary)
 library(DescTools)
 library(here)
 
@@ -28,11 +27,11 @@ source('scripts/export_gender_data.R') # save data (remove emails)
 
 #### Analyse --------
 source('scripts/load_gender_data.R') # Loads the data
-source('scripts/summarise_gender.R') # Requires work
-source('scripts/quantify_authorship.R') # Requires work
-source('scripts/visualise_gender.R') # Requires work
-source('scripts/visualise_continent.R') # Requires work
+source('scripts/summarise_gender.R') # removed modelsummary
+source('scripts/quantify_authorship.R') # Added overall. Add psych data?
+source('scripts/visualise_gender.R') # Which representation we prefer?
+#source('scripts/visualise_continent.R') # Is this needed?
 
 #source('scripts/visualise_gender_trends.R') # OK
 #source('scripts/visualise_names.R') # OK
-source('scripts/keyword_analysis.R') # creates figure (OK)
+#source('scripts/keyword_analysis.R') # creates figure (OK)

### 2025-08-19 scripts/attribute_gender.R

@@ -1,8 +1,6 @@
 # install.packages("genderizeR")
 #suppressPackageStartupMessages(library(genderizeR))
 
-library(modelsummary)
-
 run_genderize <- FALSE
 
 if (run_genderize == TRUE) {
@@ -57,7 +55,18 @@ for (k in 1:nrow(inferred)) {
 df <- dplyr::filter(df, CATEGORY != "BOOK")
 
 # summarise
-print(datasummary(Gender + 1 ~ N + Percent(), data = df))
+
+gender_with_total <- df %>%
+  count(Gender) %>%
+  mutate(percentage = round(n / sum(n) * 100, 1)) %>%
+  bind_rows(
+    summarise(., 
+              Gender = "Total", 
+              n = sum(n), 
+              percentage = sum(percentage))
+  )
+
+print(knitr::kable(gender_with_total))
 
 # Remove unattributed gender
 df <- dplyr::filter(df, Gender != 'ambiguous' & Gender != 'unknown')
@@ -67,5 +76,4 @@ cat(paste(
   nrow(df)
 )) 
 
-detach("package:modelsummary", unload=TRUE)
-rm(unattributed, inferred, k, index, run_genderize)
+rm(unattributed, inferred, k, index, run_genderize,gender_with_total)

### 2025-08-19 scripts/attribute_gender_diagnostics.R

@@ -1,5 +1,4 @@
 # attribute_gender_diagnostics.R
-library(modelsummary)
 #### Simple diagnostics
 
 cat(paste('\nUnique studies:', length(unique(df$BIBTEXKEY)),'\n'))
@@ -11,7 +10,19 @@ cat(paste(
 
 cat(paste('\n Number of rows:', nrow(df),'\n'))
 
-print(datasummary(JOURNAL + 1 ~ N + Percent(), data = dplyr::filter(df,author_order==1),output='markdown',title='Number of studies across journals'))
+journal_with_total <- df %>%
+  count(JOURNAL) %>%
+  mutate(percentage = round(n / sum(n) * 100, 1)) %>%
+  bind_rows(
+    summarise(., 
+              JOURNAL = "Total", 
+              n = sum(n), 
+              percentage = sum(percentage))
+  )
+
+print(knitr::kable(journal_with_total))
+
+
 
 cat("\n\n")
 
@@ -28,8 +39,8 @@ cat(paste("Prop. under .55 conf. =",sum(df$Gender.Probability<.55),"/n/n"))
 
 cat("\n***")
 
-print(datasummary(Mean + Median ~ Gender.Probability * Gender, data = df,output='markdown',title='Probability of API attributed Gender.'))
-
+S <- dplyr::summarise(group_by(df,Gender),M=mean(Gender.Probability),Md=median(Gender.Probability))
+print(knitr::kable(S,digits = 2))
 cat("\n***")
 
 upperCI<-function(x, conf.level = 0.95) {
@@ -44,19 +55,14 @@ lowerCI<-function(x, conf.level = 0.95) {
   #upperCI <- f[2]
   return(lowerCI)
 }
+# citations across gender
+S <- dplyr::summarise(group_by(df,Gender),M=mean(Citations),Md=median(Citations),LCI=lowerCI(Citations),UCI=upperCI(Citations))
+print(knitr::kable(S,digits = 2))
 
-print(datasummary(Mean + upperCI + lowerCI ~ Citations * Gender, data = df,output='markdown',title='Citation counts across Gender'))
-
+# get the stats of citations across gender
 print(broom::glance(t.test(Citations ~ Gender, data=df)))
 print(broom::glance(wilcox.test(Citations ~ Gender, data=df)))
-
-detach("package:modelsummary", unload=TRUE)
-
 
 #### Gender distribution of authors -----------
[…]

### 2025-08-19 scripts/country_table.R

@@ -33,6 +33,6 @@ g1<-ggplot(DATA,aes(x=reorder(Country,Odds_ratio),y=Odds_ratio,label=paste0("n="
   theme_classic(base_size = 14)+
   geom_hline(yintercept = 1,linetype='dashed',color='grey20')+
   xlab('Country') +
-  ylab('Female Authorship Odds Ratio') +
+  ylab('Female Authorship Odds Ratio (95%CI)') +
   coord_flip()
 print(g1)

### 2025-08-19 scripts/name_gender_fixes.R

@@ -1,7 +1,7 @@
 # Filename: name_gender_fixes.R
 # Author: T. Eerola, 9/8/2025
 # Project: gender in music psychology
-# Status: Complete
+# Status: To be checked
 
 #### 1. Manual fixes to gender by authors (TE and AC) --------------
 cat("Manual fixes\n")
@@ -14,6 +14,32 @@ df$Gender[df$first_name=='Zohar']<-'male'   # TE
 # from a full name
 df$Gender[df$full_name=='Ashley D. Vanstone']<-'male' # TE
 df$Gender[df$full_name=='Andrea Schiavio']<-'male'    # TE
+df$Gender[df$full_name=='J. Riikka Ahokas']<-'female'    # TE 
+df$Gender[df$full_name=='J. Ginsborg']<-'female'    # TE
+df$Gender[df$last_name=="graham-jackson"] <- "female" # TE
+
+# few names encoded as initials (Ahokas and Ginsborg), but others as well
+# but the gender of these are correct
+df$full_name[df$last_name=="Schellenberg"]
+df$full_name[df$last_name=="Dowling"]
+df$full_name[df$last_name=="Neill"]
+df$full_name[df$last_name=="Springer"]
+df$full_name[df$last_name=="Anta"]
+df$full_name[df$last_name=="McAuley"]
+df$full_name[df$last_name=="Ter Bogt"]
+df$full_name[df$last_name=="Fitch"]
+df$full_name[df$last_name=="Windsor"]
+df$full_name[df$last_name=="de Haas"]
+df$full_name[df$full_name=="B. S. Gupta"]
+df$full_name[df$last_name=="Fung"] # several variants and two persons
+df$full_name[df$last_name=="Beaman"] 
+df$full_name[df$last_name=="Yates"] 
+df$full_name[df$last_name=="Zhang"] 
+df$full_name[df$last_name=="Gaydecki"] 
+df$full_name[df$last_name=="Ali"] 
+df$full_name[df$last_name=="Lloyd"] 
+df$full_name[df$last_name=="graham-jackson"] 
+
 
 df$Gender[df$full_name=='Andrea Ravignani']<-'male' # AMC
 df$Gender[df$full_name=='Alexis Kirke']   <-'male' # AMC

### 2025-08-19 scripts/quantify_authorship.R

@@ -10,7 +10,7 @@ df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Single' # FOCUS
 t<-table(df$authortype,df$Gender)
-or1 <- effectsize::oddsratio(t) # 1.10
+or1 <- effectsize::oddsratio(t) # 1.00
 #print(or1)
 
 #### Classify order: First author ----------
@@ -19,18 +19,17 @@ df$authortype[df$author_order == 1] <- 'First'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-or2 <- effectsize::oddsratio(t) # 1.32
-#print(or1)
+or2 <- effectsize::oddsratio(t) # 1.41
+#print(or2)
 
-#### classify order: Coauthor --------------
+#### Classify order: Coauthor --------------
 df$authortype <- 'Coauthor'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-
-or3 <- effectsize::oddsratio(t) # 1.01
-#print(or2)
+or3 <- effectsize::oddsratio(t) # 1.00
+#print(or3)
 
 #### Classify order: Last -------------
 df$authortype <- 'Other'
@@ -38,13 +37,13 @@ df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Last'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-t
 or4 <- effectsize::oddsratio(t) # 0.73
+#print(or4)
 
 or <- rbind(or1,or2,or3,or4)
 or$name<-c("Single","First","Coauthor","Last")
 or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
-print(knitr::kable(or))
+print(knitr::kable(or,digits = 2))
 
 #### Annually ---------
 
@@ -53,10 +52,28 @@ df$YEAR_5 <- cut(df$YEAR, breaks=seq(2000, 2025, by=5), labels=c("2000-2004", "2
[…]

### 2025-08-19 scripts/summarise_gender.R

@@ -2,22 +2,31 @@
 # Author: T. Eerola, 6/4/2024
 # Project: gender in music psychology
 # Status: in progress
-invisible(library(modelsummary))
 
 #### 1. Gender distribution ------
 
-# % of female authors in this field
-print(datasummary(Gender + 1 ~ N + Percent(), data = df,output='markdown'))
+gender_with_total <- df %>%
+  count(Gender) %>%
+  mutate(percentage = round(n / sum(n) * 100, 1)) %>%
+  bind_rows(
+    summarise(., 
+              Gender = "Total", 
+              n = sum(n), 
+              percentage = sum(percentage))
+  )
 
-tmp <- df %>%
-  select(first_name, Gender) %>%
-  filter(Gender == 'female') %>%
-  drop_na() %>%
-  group_by(first_name) %>%
-  summarise(n = n())
-tmp <- dplyr::arrange(tmp, -n)
-print(knitr::kable(head(tmp, 5), caption = 'Top 5 names.'))
-rm(tmp)
+print(knitr::kable(gender_with_total))
+
+
+# tmp <- df %>%
+#   select(first_name, Gender) %>%
+#   filter(Gender == 'female') %>%
+#   drop_na() %>%
+#   group_by(first_name) %>%
+#   summarise(n = n())
+# tmp <- dplyr::arrange(tmp, -n)
+# print(knitr::kable(head(tmp, 5), caption = 'Top 5 names.'))
+# rm(tmp)
 
 #### 3. Number of co-authors -----------
 #print(datasummary(factor(author_order) + 1 ~ Gender + N + Percent(), data = df))
@@ -32,7 +41,4 @@ cat(paste("\n mean:",round(mean(s$n),3)))
 cat(paste("\n sd:",round(sd(s$n),3)))
 cat(paste("\n max:",round(max(s$n),3)))
 cat("\n")
-rm(s)
-
-
-detach("package:modelsummary", unload=TRUE)
+rm(s,gender_with_total)

### 2025-08-19 scripts/quantify_authorship.R
Revised analyses, added references, additions to MS

@@ -128,13 +128,7 @@ annual<-rbind(single,first,coauthor,last)
 annual$Type = factor(annual$Type, levels = c('Single','First','Co', 'Last'))
 annual
 
-# Create a new facet variable that group INCREASE / DECREASE of OR 
-#annual <- annual %>%
-#  mutate(TypeGroup = case_when(Type %in% c("Co", "Last") ~ "Co and Last",TRUE ~ Type))
-#annual$TypeGroup = as.factor(annual$TypeGroup)
-#annual$TypeGroup = factor(annual$TypeGroup, levels = c( "First" ,"Co and Last"))
-
-# add overall
+# Add overall numbers
 or$name = factor(or$name, levels = c('Single','First','Coauthor', 'Last'), labels = c('Single','First','Co', 'Last'))
 
 tmp <- data.frame(Odds_ratio = or$Odds_ratio, CI=0.95,CI_low = or$CI_low, CI_high = or$CI_high, YearRange = "Overall", Type = or$name)

### 2025-08-19 manuscript/manuscript.qmd
Revised analyses, added references, additions to MS

@@
 
-Equality – Why Should We Care?
 Academia suffers from issues regarding equity.
@@
 Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
-For instance, in the UK, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the academia [@harris2025gender].
-A a pay gap remains between men and women in academia (female-male wage ratio 0.85 in 2020) [@quadlin2023higher].
+For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
+Similar situation holds for italy [@Filandri03082021], 34% in the US [@huang2020].
+A a pay gap remains between men and women in academia (female-male wage ratio of 0.85 in 2020 in the UK) [@quadlin2023higher].
 In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.

### 2025-08-19 manuscript/manuscript.qmd
Revised text to MS

@@
 For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
-Similar situation holds for italy [@Filandri03082021], 34% in the US [@huang2020].
-A a pay gap remains between men and women in academia (female-male wage ratio of 0.85 in 2020 in the UK) [@quadlin2023higher].
+A similar situation holds for italy (24% female professors) [@Filandri03082021], and to some extent for the US (14%) [@spoon2023].
+A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the UK) [@quadlin2023higher].
 In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
@@
 
-Research across several scientific disciplines that show that women tend to be less represented in authors compared to men [@banks2025women]; @son2022scientific].
+Research across several scientific disciplines that show that women tend to be less represented in authors compared to men [@banks2025women; @son2022scientific].
 Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving.
@@
 
-See @gonzalez2020women for a baseline from other disciplines:
-- Hard Sciences (N=119,592) = 0.148
-- Bio and Social Sciences (N=262,122) =  0.433
-- Psychology (N=90,067) =  0.452
+It is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
+For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%).
+In psychology, which is perhaps the closest benchmark for the present focus on music psychology, female authors accounted for 45.2% of the sample (N = 90,067).
 
 What is the situation in music psychology?
-To improve certain academic equity issues, it is worthwhile focusing within disciplines and seeing how more local academic communities can provide more equity.
-To do this, it is worth giving a snapshot of specific fields.
-Indeed, instead of looking broadly at journal articles in STEM, or in several different kinds of journals, it can be worthwhile giving an idea of a more specific.
-For example, one paper explores authorship in Herpetology, showing that male authors outnumber female, but female authorship is increasing steadily across the last decade [@rock2021quantifying].
-The current paper also focuses on music psychology as a discipline: within music psychology, there are many efforts to improve the field.
-There have been papers that provide a snapshot of the field to provide a rough overview of the state of affairs (e.g., @anglada2019visualizing) with suggestions of future directions to improve the field.
+
+To improve certain academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
+The past efforts have already provided a snapshot a rough overview of the state of affairs (e.g., @anglada2019visualizing) with suggestions of future directions to improve the field.
 [@eerola2023transparency] showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
 @jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is, which the view to see how – as a field – we need to diversify participant samples and stimuli to gain a more realistic understanding.
-In a similar way to these papers, the current paper aims to give a current overview of authorship in the field of music psychology.
+In a similar way to these papers, the current paper aims to give a current overview of authorship and gender patterns in the field of music psychology.
 
@@
 Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
-What is the proportion of different types of authorships (lead, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
+What is the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
 Are there specific trends in terms of countries of the affiliations, topics, or time?

### 2025-08-19 analysis.qmd
Added keyword visualisations

@@
 source('scripts/attribute_gender.R') # Adds gender attributions (precalculated with API)
+source('scripts/name_gender_fixes.R') # Fixes the gender attribution of some names manually 
 source('scripts/create_keys.R') # Add keys for authors and study+author
@@
 
-```{r}
-#| id: geographical2
-#| eval: true
-#| echo: true
-#| output: asis
-#source('scripts/table_continent.R') # IN progress
-```
-
 Table: Female author odds ratios across continents.
@@
 |Africa    |   52|       0.91|   0.52|    1.59|
+
+### Keywords and gender
+
+```{r}
+#| id: keywords
+#| eval: true
+#| echo: true
+#| fig.width: 10
+#| fig.height: 6
+source('scripts/keyword_analysis.R') 
+```

### 2025-08-19 contents.R

@@ -32,6 +32,5 @@ source('scripts/quantify_authorship.R') # Added overall. Add psych data?
 source('scripts/visualise_gender.R') # Which representation we prefer?
 #source('scripts/visualise_continent.R') # Is this needed?
 
-#source('scripts/visualise_gender_trends.R') # OK
-#source('scripts/visualise_names.R') # OK
-#source('scripts/keyword_analysis.R') # creates figure (OK)
+#source('scripts/visualise_names.R') # Bonus feature
+source('scripts/keyword_analysis.R') # creates 3 alternative figures

### 2025-08-19 scripts/attribute_gender_diagnostics.R

@@ -11,6 +11,7 @@ cat(paste(
 cat(paste('\n Number of rows:', nrow(df),'\n'))
 
 journal_with_total <- df %>%
+  ungroup() %>%
   count(JOURNAL) %>%
   mutate(percentage = round(n / sum(n) * 100, 1)) %>%
   bind_rows(
@@ -21,8 +22,6 @@ journal_with_total <- df %>%
   )
 
 print(knitr::kable(journal_with_total))
-
-
 
 cat("\n\n")
 

### 2025-08-19 scripts/citations.R

@@ -7,8 +7,8 @@ df$authortype[df$author_last == 1] <- 'Other'
 tmp<-dplyr::filter(df,authortype=='First')
 citestats_first <- summarise(group_by(tmp, Gender), Md = median(Citations), M = mean(Citations),Q75=quantile(Citations,0.75))
 stats_first <- broom::glance(kruskal.test(Citations ~ Gender, data=tmp))
-citestats_first$CI_lower<-NULL
-citestats_first$CI_upper<-NULL
+citestats_first$CI_lower<-c(0,0)
+citestats_first$CI_upper<-c(0,0)
 citestats_first$CI_lower[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[2]
 citestats_first$CI_upper[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[3]
 citestats_first$CI_lower[2]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='male'], conf.level = 0.95)[2]
@@ -22,8 +22,8 @@ df$authortype[df$author_last == 1] <- 'Other'
 tmp <- dplyr::filter(df,authortype=='Coauthor')
 citestats_coauthor <- summarise(group_by(tmp, Gender), Md = median(Citations), M = mean(Citations),Q75=quantile(Citations,0.75))
 stats_coauthor <- broom::glance(kruskal.test(Citations ~ Gender, data=tmp))
-citestats_coauthor$CI_lower<-NULL
-citestats_coauthor$CI_upper<-NULL
+citestats_coauthor$CI_lower<-c(0,0)
+citestats_coauthor$CI_upper<-c(0,0)
 citestats_coauthor$CI_lower[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[2]
 citestats_coauthor$CI_upper[1]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='female'], conf.level = 0.95)[3]
 citestats_coauthor$CI_lower[2]<-DescTools::MedianCI(tmp$Citations[tmp$Gender=='male'], conf.level = 0.95)[2]
@@ -37,8 +37,8 @@ df$authortype[df$author_last == 1] <- 'Last'
 tmp <- dplyr::filter(df,authortype=='Last')
[…]

### 2025-08-19 scripts/keyword_analysis.R

@@ -1,0 +1,231 @@
+# keyword_analysis.R
+# gender in music psych
+# T. Eerola, 19/8/2025
+# Status: In progress
+
+#### Custom functions -------------------
+source('scripts/count2category.R')
+plotflag <- TRUE
+
+## remove studies, keep articles (to avoid duplicating keywords)
+D <- dplyr::filter(df, author_id == 'author1') # 3373
+
+if (!dim(D)[1] == 3373) {
+  print("incorrect number of observations! (Needs to be run at the level of studies, N=1360)")
+  break
+}
+
+#### keyword analysis ------------------
+KW <- NULL
+d_index <- NULL
+GENDER_index <- NULL
+year_index <- NULL # year
+
+for (k in 1:nrow(D)) {
+  tmp <- D$AUTHOR_KEYWORDS[k]
+  GENDER <- as.character(D$Gender[k])
+  y <- D$YEAR[k]
+  tmp <- stringi::stri_trans_general(tmp, "latin-ascii")
+  if (!is.na(tmp)) {
+    kw <- str_split(tmp, '[;,]', simplify = TRUE)
+    KW <- c(KW, kw)
+    d_index <- c(d_index, rep(k, length(kw)))
+    # u_index <- c(u_index, rep(u, length(kw)))
+    # o_index <- c(o_index, rep(o, length(kw)))
+    # m_index <- c(m_index, rep(m, length(kw)))
+    # a_index <- c(a_index, rep(a, length(kw)))
+    # g_index <- c(g_index, rep(g, length(kw)))
+    year_index <- c(year_index, rep(y, length(kw)))
+    GENDER_index <- c(GENDER_index, rep(GENDER, length(kw)))
+  }
+}
+
+length(KW)
+length(GENDER_index)
+length(year_index)
+
+KW <- stringi::stri_trans_general(KW, "latin-ascii")
+KW
+KW<-stringr::str_trim(KW, side = "both")
+KW<-tolower(KW)
+
+#source('scripts/simplify_keywords.R')
+
+data <- data.frame(KW,
+                   GENDER_index,
+                   year_index)
+x1 <- count2category(data,
+                     index = "GENDER_index",
+                     str1 = "female",
+                     str2 = "male")
+
+FROM <- 1
+TO <- 25
+
+#### Figure 1 ----------
+
+figure1 <- ggplot(data = x1[FROM:TO, ], aes(
[…]

### 2025-08-19 scripts/summarise_gender.R

@@ -6,6 +6,7 @@
 #### 1. Gender distribution ------
 
 gender_with_total <- df %>%
+  ungroup() %>%
   count(Gender) %>%
   mutate(percentage = round(n / sum(n) * 100, 1)) %>%
   bind_rows(

### 2025-08-19 manuscript/manuscript.qmd
Small tweaks to MS

@@
 
-What is the situation in music psychology?
-
 To improve certain academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
@@
 
+```{r}
+#| id: results
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+source(here("scripts","load_gender_data.R"))
+```
+
 We retrieved bibliographic information for all articles published between 2000 and June 2025 from five specialist journals, resulting in 3,373 unique articles: Musicae Scientiae (N = 639), Psychology of Music (N = 1,231), Music Perception (N = 675), Journal of New Music Research (N = 563), and Music & Science (N = 265).
@@
 
-In total, the dataset included 9,066 authors, of whom 5,573 were unique.
-Author affiliations spanned 63 countries.
+In total, the dataset included `{r} nrow(df)` authors, of whom `{r} length(unique(df$unique_name))` were unique.
+Author affiliations spanned `{r} length(unique(df$Affiliation_country_code))` countries.
 We also extracted citation counts and Open Access status from Scopus.
@@
 This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
-Only 89 names had a low attribution probability (< 0.55).
-Unattributed cases were then checked manually, resulting in 185 manual corrections.
-After this process, 32 names remained ambiguous and 27 were unknown—some likely due to data entry errors in Scopus (e.g.
+Only `89` names had a low attribution probability (< 0.55).
+Unattributed cases were then checked manually, resulting in `185` manual corrections.
+After this process, `32` names remained ambiguous and `27` were unknown—some likely due to data entry errors in Scopus (e.g.
 only initials or surnames).
-These 59 cases were excluded from the dataset.
+These `59` cases were excluded from the dataset.
 It is likely that the due to challenge of attributing gender correctly not all the gender attributions are correct.
-Although genderize.io has been shown to achieve 96.6% accuracy in a diverse multinational test database without using a country of origin information -- 98% with the information included -- [@vanhelene2024inferring], the tool underperforms in Asian names (82% accuracy).
-However, the small error has been previously been shown to be non-biased, i.e.
+Although genderize.io has been shown to achieve 96.6% accuracy in a diverse multinational test database without using a country of origin information -- 98% with the information included -- it also underperforms (82% accuracy) in Asian names [@vanhelene2024inferring].
+However, the error rates of these processes have been previously been shown to be non-biased, i.e.
 showing similar number of mistakes for both genders [@sebo2021performance;@vanhelene2024inferring].
@@
 We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full first name coded with initials, or the second name used as the first name.
-This led to `XXXX` corrections.
+This led to `32` corrections.
 
-We defined three types of authorship positions: first authorship, coauthorship, and last authorship.
+We defined four types of authorship positions: single authorship, first authorship (which does not include papers with a single author), coauthorship, and last authorship.
 Coauthorship includes all positions other than first and last.
@@
 
-```{r}
-#| id: results
-#| eval: true
-#| echo: false
-#| warning: false
-#| message: false
-source(here("scripts","load_gender_data.R"))
-```
-
-Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),2)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
+Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
 To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of females occupying different authorship positions.
@@
 
-The overall odds ratio for women as first authors was 1.32 [95% CI: 1.21–1.44], indicating that the odds of first authorship are 32% higher for females compared to the male authors.
-For coauthorship positions, the OR was 0.99 [0.90–1.08], suggesting no substantial gender difference.
-In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower in females.
+The overall odds ratio for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors.
+For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference.
+In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men.
+Being a single author has odds ratio of 1.11 [0.95-1.29] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not substantial (95% crosses 1.0).
 For the first and last authorship position, the odds ratios are significant at _p_ < .001 level.
@@
 To illustrate the trends across time in the authorship positions across the gender, we looked at 5-year-windows of the authorship types across gender, shown in Figure 1.
+The overall numbers across the whole data are illustrated in the first column of the graph and the remaining five columns portray the chronology in 5-year windows.
 The higher odds for female authors at first author position seems to be established in 2010 onwards.
@@
 However, the last author positions have become relatively more rare after 2010, showing statistically significantly lower odds ratios from 2010 onwards and negative 5-year growth rate (-9.5%).
+The odds ratio for female single authored papers hovers around 1.0, showing no meaningful trend across the time.
 
@@
 We can observe that while female authors are not more likely to be single authors than men, they tend to be more likely to be first authors than males in papers with 2 or 3 authors; the odds ratios for first authorship position for female authors for papers with 2 authors is 1.55 [130-1.85], and for 3 authors the odds ratio is 1.61 [1.35-1.92], both significantly different from equal proportion.
-The last authorship positions for female authors are consequently lower, odds ratio of 0.65 [0.54–0.77] for 2 authors and 0.59 [0.49–0.71] for 3 authors, both also significant at _p_<.001 level.
+The last authorship positions for female authors are consequently lower, odds ratio of 0.65 [0.54–0.77] for 2 authors and 0.59 [0.49–0.71] for 3 authors, both also significant at _p_ < .001 level.
 At four and five authors, the higher odds ratio for female is still evident (OR of 1.37 both both), but the last authorship positions are no longer relative rare for papers with 4 and 5 authors (OR 0.90 and 0.95, respectively).
@@
 Here we tested whether the citations -- as indexed by Scopus -- show differences across gender.
-The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_=`{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
+The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_ = `{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
 The same comparison of citations for studies with last authors yields similar results `{r} citestats_last$Md[1]` [`{r} citestats_last$CI_lower[1]`–`{r} citestats_last$CI_upper[1]`], and for males, the numbers are similar (Md=`{r} citestats_last$Md[2]` [`{r} citestats_last$CI_lower[2]`–`{r} citestats_last$CI_upper[2]`]).
-Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_=`{r} round(stats_last$p.value,3)`).
-If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_=`{r} round(stats_all$p.value,4)`).
+Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_ = `{r} round(stats_last$p.value,3)`).
+If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_ < `{r} round(stats_all$p.value,4)`).
 
@@
 
-OPTIONAL
+## Gendered topics through analysis of keywords
 
-Table: Female author odds ratios across continents.
-
-|Continent |    N| Odds_ratio| CI_low| CI_high|
-|:---------|----:|----------:|------:|-------:|
-|Europe    | 4985|       0.82|   0.75|    0.89|
-|Americas  | 2356|       1.12|   1.02|    1.23|
-|Asia      |  931|       0.89|   0.78|    1.02|
-|Oceania   |  742|       1.66|   1.43|    1.93|
-|Africa    |   52|       0.91|   0.52|    1.59|
-
-## More analyses?
-
-- Association with the keywords (similar)
+[to be done].
 
@@
 
-- Senior authorship positions
+- Senior authorship positions!
 

### 2025-08-26 scripts/keyword_analysis.R
work on keyword analysis: grouping together similar words

@@ -1,6 +1,6 @@
 # keyword_analysis.R
 # gender in music psych
-# T. Eerola, 19/8/2025
+# T. Eerola, 19/8/2025, edited also by A. M. Czepiel
 # Status: In progress
 
 #### Custom functions -------------------
@@ -54,6 +54,432 @@ KW<-tolower(KW)
 data <- data.frame(KW,
                    GENDER_index,
                    year_index)
+
+
+# Clean up similar keywords
+
+## Words related to music cognition
+idx = grep("cognition", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("music cognition",
+                           "music and cognition", 
+                           "musical and cognition",
+                           "language and music cognition",
+                           "music and social cognition") 
+                     ~ "music cognition",TRUE ~ KW))
+
+## Words related to recognition
+idx = grep("recognition", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("recognition",
+                           "optical music recognition", 
+                           "optical music recognition (omr)",
+                           "omr (optical music recognition)", 
+                           "content-based generalized sound recognition", 
+                           "tempo recognition", 
+                           "melody recognition",
+                           "serial recognition",
+                           "recognition of scale and key by experts",
+                           "emotion recognition",
+                           "recognition task" ,
+                           "auditory recognition", 
+                           "pattern recognition",
+                           "sound recognition",
+                           "recognition performance") 
+                 ~ "music recognition",TRUE ~ KW))
+
+## Words related to music emotion
+idx = grep("emotion", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("emotional music experience",
+                           "music and emotions", 
+                           "music emotion recognition",
+                           "emotional response to music",
+                           "music emotion classification",
+                           "emotional valence",
+                           "music-perceived emotions",
+                           "musical emotions",
+                           "music induced emotions",
+                           "music-induced emotion",
+                           "music-evoked emotion",
+                           "negative emotion in music",
+                           "musical emotion mechanisms",
+                           "emotional use of music",
+                           "music emotion identification" ,
+                           "musical emotional contagion",  
+                           "emotion in music",
+                           "emotions during listening" ) 
+                     ~ "music emotions",TRUE ~ KW)
+)
+
[…]

### 2025-08-26 scripts/keyword_analysis.R
work on keyword analysis: continued work

@@ -129,7 +129,7 @@ data = data %>% mutate(
                                "embodied knowledge",
                                "embodied",
                                "embodied mind",
-                               "4e cognition") 
+                               "4e cognition‌") 
                      ~ "embodied music cognition",TRUE ~ KW)
 )
 
@@ -175,8 +175,6 @@ data = data %>% mutate(
                  ~ "historical performance practice",TRUE ~ KW))
 
 ## Words related to performance anxiety
-idx = grep("performa", data$KW)
-unique(data$KW[idx])
 data = data %>% mutate(
   KW = case_when(KW %in% c("performance anxiety",
                            "anxiety-performance relationship",
@@ -477,6 +475,60 @@ data = data %>% mutate(
                            "social collaboration")
                  ~ "social",TRUE ~ KW))
 
+
+# Mental health
+idx = grep("mental health", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("mental health",
+                           "mental health and the music industry",
+                           "mental health interventions",
+                           "mental health help-seeking" ,
+                           "mental health care")
+                 ~ "mental health",TRUE ~ KW))
+
+
+# health
+idx = grep("health", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("health musician" ,
+                           "health care",
+                           "health" ,
+                           "health behaviors"  ,
+                           "healthcare",
+                           "healthy ageing" ,
+                           "health-promoting behaviours" ,
+                           "health promotion"  ,
+                           "health-oriented preventive behavior",
+                           "health-literacy" ,
+                           "physical health"     ,
+                           "health psychology",
+                           "healthcare services" ,
+                           "health outcomes"  ,
+                           "public health"  ,
+                           "health responsibility" ,
+                           "occupational health"  ,
+                           "general health" ,
+                           "healthy adults"    ,
+                           "job-related health problems")
+                 ~ "health",TRUE ~ KW))
+
+# art/music health
+idx = grep("health", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("arts in health",
+                           "music and health"  ,
+                           "arts-and-health",
+                           "healthy music use",
+                           "unhealthy music use",
+                           "vocal health" ,
+                           "playing-related health problems",
+                           "musician's health"  ,
+                           "arts and health",
+                           "health musicking" )
+                 ~ "music health",TRUE ~ KW))
 
 
 #-------

### 2025-08-26 manuscript/manuscript.qmd
additional text to MS, keywords analysis altered, figures revised

@@
 
-abstract: "This study examines the gender distribution of the authors in music psychology. N=XXXX published papers between years 2000 to 2025 from 5 journals were ..."  
+abstract: "This study examines the gender distribution of authors in music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
+In addition, patterns related to citations, open-access publishing, and keywords were explored across gender.
+Overall, female authors were in the minority, accounting for 40.2% of all authors.
+However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
+These patterns have become more pronounced over the past 15 years.
+No substantial gender differences in citation counts were observed.
+Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
+Finally, an examination of keywords suggested broad thematic preferences differing by gender."
 keywords          : "gender, transparency, music psychology, open science, meta-research"
@@
 df <- mutate(group_by(df, BIBTEXKEY), author_N = max(author_order))
+
+s2 <- summarise(group_by(df, BIBTEXKEY,Gender), n = n(),.groups="drop")
+
+S <- s2 %>%
+  dplyr::group_by(Gender) %>%
+  dplyr::summarise(N=n(),m=mean(n,na.rm = TRUE),sd=sd(n,na.rm = TRUE),.groups = 'drop') %>%
+  dplyr::mutate(se=sd/sqrt(N),LCI=m+qnorm(0.025)*se,UCI=m+qnorm(0.975)*se)
+
+T <- t.test(s2$n~s2$Gender)
+
 ```
-The range of the number of authors in varies considerable.
+The variation of the number of authors across the papers is considerable.
 Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`, SD=`{r} round(sd(s$n),2)`).
+Previous research has calculated the collaboration index across gender [@gonzalez2020women], suggesting that male author tend to have larger number of coauthors (M = 1.5) than female (M = 1.34) in psychology.
+Here the average author numbers for male and female authors (disregarding authorship positions) indicate similar difference, higher mean author numbers for males (M = `{r} round(S$m[2],2)`, [`{r} round(S$LCI[2],2)`-`{r} round(S$UCI[2],2)`]) than for females (M = `{r} round(S$m[1],2)`, [`{r} round(S$LCI[1],2)`-`{r} round(S$UCI[1],2)`]), _t_(4640.58) = 7.21, _p_ < .001, Cohen's _d_ = -0.21, 95% CI
+[-0.26, -0.15]. This analysis, however, does not allow us to distinguish the authorship roles in the counts.
+
 To illustrate how the actual number of authors in a paper impacts female first author likelihood, we calculated the odds ratios across the number of authors in paper, pooling the papers with 7 or more authors together (N=`{r} sum(df$author_N>6)`,  `{r} round(sum(df$author_N>6)/nrow(df)*100,1)`%).
@@
 
+For comparison, the closest disciplinary benchmark comes from psychology, where @gonzalez2020women analyzed the gender distribution of 74,413 authors publishing in psychology journals in 2009.
+Women accounted for 45.2% of the sample, indicating near gender parity overall.
+However, substantial variation was observed across sub-disciplines: developmental psychology (57.0%) and educational psychology (53.8%) showed higher proportions of female authors, whereas biological psychology (39.4%), experimental psychology (37.5%), and mathematical psychology (25.8%) showed lower representation.
+Across the full dataset, women were slightly overrepresented in first authorship positions (OR = 1.10) but underrepresented in last authorship positions (OR = 0.86).
+This pattern resembles the trends identified in the present study.
+
 ## Citations and Open Access
+
 ```{r}
@@
 If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_ < `{r} round(stats_all$p.value,4)`).
+For reference, prior study of citation patterns across gender in psychology has reported a small effect favoring men (M = 16.57) over women (M = 15.90) [@gonzalez2020women].
 
@@
 Out of `{r} as.numeric(sum(t))` articles, `{r} round(as.numeric(t[1]/sum(t))*100,2)`% are Open Access in this sample, as indexed by Scopus.
-For first-authored articles, female odds ratio for Open Access is `{r} round(author_OA$Odds_ratio[1],2)` [`{r} round(author_OA$CI_low[1],2)`–`{r} round(author_OA$CI_high[1],2)`], suggesting over 50% higher odds associated with publishing open access by women as compared to men.
+For first-authored articles, female odds ratio for Open Access is `{r} round(author_OA$Odds_ratio[1],2)` [`{r} round(author_OA$CI_low[1],2)`–`{r} round(author_OA$CI_high[1],2)`], suggesting nearly 50% higher odds associated with publishing open access by women as compared to men.
 If we observe only the last authorship status of the publications, the difference vanishes with the odds ratio indisguishable from even division (1.00 and the confidence interval), OR = `{r} round(author_OA$Odds_ratio[3],2)` [`{r} round(author_OA$CI_low[3],2)`–`{r} round(author_OA$CI_high[3],2)`].
@@
 For example, a large-scale analysis of psychology publications found that 46.5% of authorships were by women, with European countries showing a lower average proportion of 42.8% [@gonzalez2020women].
-To examine geographical differences more closely, we calculated the odds of female authorship by country, as shown in Figure 3.
+To examine geographical differences more closely, we calculated the odds of female authorship by the countries that have published at least 60 publications in the sample, as shown in Figure 3.
 
@@
 
+- If we want to add the continents at this figure, that's possible.
+
 ## Gendered topics through analysis of keywords
 
-[to be done].
+To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (15,316 in total, 6730 unique).
+To reduce variant keywords, we manually simplified the variants (e.g., "arts in health", "healthy music use", "health musicking" were converted to "music health" and "music cognition" to "cognition" and "music perception" to "perception") and eliminated single use of the term "music".
+We aggregated the counts of keywords across all authors in the articles and Figure 4 shows the odds ratios for female authors to be associated (in any author position) with the 40 most frequently used keywords.
+The keywords signifying related to development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
+At the opposite end of this continuum, topics such as music analysis, machine learning, harmony, cognition, rhythm, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+
+```{r}
+#| id: keywords
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 9
+#| fig-height: 8
+#| fig.cap: > 
+#|   Female author odds ratio across 40 most frequent keywords in all articles as aggregated across all authors.
+library(stringr)
+source(here("scripts","count2category.R"))
+source(here("scripts","keyword_analysis.R"))
+
+```
 
@@
 The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.
-While overall women hold smaller number of authorship overall (xx%),
+While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are more likely to hold first author positions in coauthored publications but less likely likely to hold last author positions than men.
+For solo publications and being a coauthor, there are no significant gender differences.
 
-- Encouraging overall
+There was small but significant effect of gender on citations that favoured men when the citations were aggregated all authors in the papers.
+There was no, however, differences in citations for female first or last  authors.
+The aggregated results are in line with previous findings in psychology that favored men [@gonzalez2020women].
+We also observed that female first authored articles tend to be more likely to be published as open access as male first authors.
+The past surveys of experiences and attitudes towards open access publishing has revealed gender differences typically in the opposite direction [@zhu2017support].
 
-- Compare with sub-disciplines of psychology [@gonzalez2020women] Psychology: First authors (N=33,631)
+While the present analysis cannot determine the underlying causes of the observed trends, we can point to several factors that may help explain why this discipline appears to be more equitably distributed by gender than, for example, psychology.
 
-- Female role models (past and present) exist in music psychology (list names and their eminent roles, chairs of societies, editors of the main journals, etc.
-Diana Deutsch, Carol Krumhansl, Irene Deliege, to a more recent ICMPC and ESCOM presidents).
+(1) Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch and Edward Carterette for ICMPC in 1989, Irène Deliège for ESCOM in 1991), to a more recent ICMPC and ESCOM presidents?
+Editor-in-chiefs of the journals?).
 
-- Is this an area that appeals to women? (see comparison to sub-disciplines of psychology, where "softer" areas tend to have higher proportion)
+(2) Taking sub-disciplinary differences in psychology as an indicator of topic areas more often favoured by women, @gonzalez2020women observed that “nurturing” disciplines such as education and developmental psychology had higher proportions of female authors.
+In music psychology, however, a similar explanation may not apply.
+Keywords typically associated with nurturing domains (e.g., education, development, social, therapy, well-being, mental health, emotion regulation) account for only 17.9% of the total keyword counts among the top 40 keywords.
 
-- Not matched by musicology: 25% female profs, 38% below??? (these are just some UK stats, I don't think the field-wide analysis has been done)
+- It would have been interesting to analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018]
 
-- Senior authorship positions!
+- Current findings present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
+While we have comparable accounts from psychology, musicology and other disciplines of music research (popular music, ethnomusicology) tends to have less equal division gender; a recent analysis of the last 25 years of articles in 15 musicology journals amounting to 3,515 articles suggests that female authors are far more infrequent (34.8% [33.3%-36.3%]) than males (65.2% [63.7%-66.7%]) [@eerola2026inprep].
+
+- Senior/last authorship position and why such a large inbalance exists.
 

### 2025-08-26 contents.R

@@ -19,7 +19,7 @@ rm(list = ls())
 source('scripts/load_scopus_datasets.R') # Creates full_names dataset
 source('scripts/attribute_country.R') # Add country affiliations
 source('scripts/attribute_gender.R') # Adds gender attributions (precalcutaed with API)
-source('scripts/name_gender_fixes.R') # Fixes the gender attribution of some names manually 
+source('scripts/name_gender_fixes.R') # Fixes the gender attribution of some names manually
 source('scripts/create_keys.R') # Add keys for authors and study+author
 source('scripts/clean_citations_and_OA.R') # Process citations and OA status
 source('scripts/attribute_gender_diagnostics.R') # Calculate how reliable is the attribution

### 2025-08-26 manuscript/tmp.R

@@ -1,0 +1,20 @@
+library(here)
+here::i_am("manuscript/manuscript.qmd")
+
+
+source(here("scripts","load_gender_data.R"))
+library(dplyr)
+s <- summarise(group_by(df, BIBTEXKEY), n = n())
+df <- mutate(group_by(df, BIBTEXKEY), author_N = max(author_order))
+
+s
+
+s2 <- summarise(group_by(df, BIBTEXKEY,Gender), n = n(),.groups="drop")
+
+# get CI
+S <- s2 %>%
+  dplyr::group_by(Gender) %>%
+  dplyr::summarise(N=n(),m=mean(n,na.rm = TRUE),sd=sd(n,na.rm = TRUE),.groups = 'drop') %>%
+  dplyr::mutate(se=sd/sqrt(N),LCI=m+qnorm(0.025)*se,UCI=m+qnorm(0.975)*se)
+
+T <- t.test(s2$n~s2$Gender)

### 2025-08-26 scripts/keyword_analysis.R

@@ -4,8 +4,9 @@
 # Status: In progress
 
 #### Custom functions -------------------
-source('scripts/count2category.R')
-plotflag <- TRUE
+#source('scripts/count2category.R')
+plotflag <- FALSE
+duplication <- TRUE
 
 ## remove studies, keep articles (to avoid duplicating keywords)
 D <- dplyr::filter(df, author_id == 'author1') # 3373
@@ -15,58 +16,92 @@ if (!dim(D)[1] == 3373) {
   break
 }
 
-#### keyword analysis ------------------
-KW <- NULL
-d_index <- NULL
-GENDER_index <- NULL
-year_index <- NULL # year
+#### keyword analysis for first authors only ------------------
+if(duplication==FALSE){
+  KW <- NULL
+  d_index <- NULL
+  GENDER_index <- NULL
+  year_index <- NULL # year
 
-for (k in 1:nrow(D)) {
-  tmp <- D$AUTHOR_KEYWORDS[k]
-  GENDER <- as.character(D$Gender[k])
-  y <- D$YEAR[k]
-  tmp <- stringi::stri_trans_general(tmp, "latin-ascii")
-  if (!is.na(tmp)) {
-    kw <- str_split(tmp, '[;,]', simplify = TRUE)
-    KW <- c(KW, kw)
-    d_index <- c(d_index, rep(k, length(kw)))
-    # u_index <- c(u_index, rep(u, length(kw)))
-    # o_index <- c(o_index, rep(o, length(kw)))
-    # m_index <- c(m_index, rep(m, length(kw)))
-    # a_index <- c(a_index, rep(a, length(kw)))
-    # g_index <- c(g_index, rep(g, length(kw)))
-    year_index <- c(year_index, rep(y, length(kw)))
-    GENDER_index <- c(GENDER_index, rep(GENDER, length(kw)))
+  for (k in 1:nrow(D)) {
+    tmp <- D$AUTHOR_KEYWORDS[k]
+    GENDER <- as.character(D$Gender[k])
+    y <- D$YEAR[k]
+    tmp <- stringi::stri_trans_general(tmp, "latin-ascii")
+    if (!is.na(tmp)) {
+      kw <- str_split(tmp, '[;,]', simplify = TRUE)
+      KW <- c(KW, kw)
+      d_index <- c(d_index, rep(k, length(kw)))
+      # u_index <- c(u_index, rep(u, length(kw)))
+      # o_index <- c(o_index, rep(o, length(kw)))
[…]

### 2025-08-26 manuscript/manuscript.qmd
additional text to MS, keywords analysis altered, figures revised

@@
 
-The overall odds ratio for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors.
-For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference.
-In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men.
-Being a single author has odds ratio of 1.11 [0.95-1.29] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not substantial (95% crosses 1.0).
-For the first and last authorship position, the odds ratios are significant at _p_ < .001 level.
+The overall odds ratio for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
+For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
+In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
+Being a single author has odds ratio of 0.90 [0.78-1.05] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not significant (_p_ = .21).
 

### 2025-08-26 scripts/create_keys.R

@@ -20,8 +20,8 @@ df$author_last[df$author_order == 1] <- 0 # except when the rank is 1
 
 #### Check all author roles -----------
 table(df$author_last,df$author_order)
-table(df$author_order==1,df$author_order) # 
-table(df$author_order==1,df$Max) # 
+table(df$author_order==1,df$author_order) #
+table(df$author_order==1,df$Max) #
 
 df$author_single <- 0
 df$author_single[df$author_order==1 & df$Max==1] <- 1 # if the author is the only one in the paper

### 2025-08-26 scripts/quantify_authorship.R

@@ -9,18 +9,35 @@ df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Single' # FOCUS
-t<-table(df$authortype,df$Gender)
-or1 <- effectsize::oddsratio(t) # 1.00
-#print(or1)
+# t<-table(df$authortype,df$Gender)
+# t
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("Single", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+t
+or1 <- effectsize::oddsratio(t) # 0.83       | [0.71, 0.96]
+print(or1)
+fisher.test(t)
 
 #### Classify order: First author ----------
 df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'First'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
-or2 <- effectsize::oddsratio(t) # 1.41
-#print(or2)
+# t<-table(df$authortype,df$Gender)
+# t
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("First", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+t
+
+or2 <- effectsize::oddsratio(t) # 1.27       | [1.03, 1.56]
+print(or2)
+fisher.test(t)
 
 #### Classify order: Coauthor --------------
 df$authortype <- 'Coauthor'
@@ -28,8 +45,17 @@ df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
 t<-table(df$authortype,df$Gender)
-or3 <- effectsize::oddsratio(t) # 1.00
-#print(or3)
+t
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("Coauthor", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+t
+
+or3 <- effectsize::oddsratio(t) # 1.06       | [0.77, 1.46]
+print(or3)
+fisher.test(t)
 
[…]

### 2025-08-26 manuscript/_figure1.qmd
additional text to MS, keywords analysis altered, figures revised

@@
 df$authortype[df$author_single == 1] <- 'Single' # FOCUS
-t<-table(df$authortype,df$Gender)
-or1 <- effectsize::oddsratio(t) # 1.00
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("Single", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+
+or1 <- effectsize::oddsratio(t) # 0.90
 #print(or1)
@@
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
+# t<-table(df$authortype,df$Gender)
+# t
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("First", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
 or2 <- effectsize::oddsratio(t) # 1.41
@@
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("Coauthor", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+
 or3 <- effectsize::oddsratio(t) # 1.00
@@
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
+t <- table(
+  factor(df$authortype, levels = c("Last", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+
 or4 <- effectsize::oddsratio(t) # 0.73
@@
 or$name<-c("Single","First","Coauthor","Last")
-or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
-#print(knitr::kable(or,digits = 2))
+#_or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
+#_print(knitr::kable(or,digits = 2))
 
@@
   tmp<-dplyr::filter(df,YEAR_5==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("Single", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
   output <- rbind(output,effectsize::oddsratio(t))
@@
   tmp<-dplyr::filter(df,YEAR_5==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("First", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
   output <- rbind(output,effectsize::oddsratio(t))
@@
   tmp<-dplyr::filter(df,YEAR_5==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("Coauthor", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
   output <- rbind(output,effectsize::oddsratio(t))
@@
   tmp<-dplyr::filter(df,YEAR_5==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("Last", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
   output <- rbind(output,effectsize::oddsratio(t))
@@
 
+custom_palette <- brewer.pal(n = 4, name = "Set1")
+custom_palette<-c(custom_palette[4],custom_palette[1:3])
+
 fig1 <- ggplot(annual2,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator))+
@@
   geom_hline(yintercept=1, linetype="dashed", color = "black") +
-  geom_point(position = position_dodge(width = 0.4),show.legend=FALSE,size=3)+
+  geom_point(position = position_dodge(width = 0.4),show.legend=TRUE,size=3)+
   geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
 #  geom_line(alpha = 0.2,position = position_dodge(width = 0.4))+
-  scale_shape_manual(values = c(22, 16)) +
-  scale_color_brewer(name="Authorship\ntype",palette = "Set1",direction=-1) +
+  scale_shape_manual(name="Authorship\ntype",values = c(22, 16),guide = "none") +
+  scale_color_manual(name="Authorship\ntype",values=custom_palette) +
   xlab('Year range')+

### 2025-08-26 manuscript/_figure2.qmd
additional text to MS, keywords analysis altered, figures revised

@@
 df$authortype[df$author_last == 1] <- 'Other'
+df$authortype[df$author_single == 1] <- 'Other'
 #table(df$authortype,df$Gender)
@@
   tmp<-dplyr::filter(df,author_N_cropped==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+#  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("First", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
   output <- rbind(output,effectsize::oddsratio(t))
@@
   tmp<-dplyr::filter(df,author_N_cropped==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+#  t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("Coauthor", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
+
   output <- rbind(output,effectsize::oddsratio(t))
@@
 df$authortype[df$author_last == 1] <- 'Last'
+df$authortype[df$author_single == 1] <- 'Other'
 
@@
   tmp<-dplyr::filter(df,author_N_cropped==U[k])
-  t <- table(tmp$authortype,tmp$Gender)
+  #t <- table(tmp$authortype,tmp$Gender)
+  t <- table(
+    factor(tmp$authortype, levels = c("Last", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
+  )
+
   output <- rbind(output,effectsize::oddsratio(t))

### 2025-08-26 scripts/quantify_authorship.R

@@ -44,8 +44,8 @@ df$authortype <- 'Coauthor'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Other'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
-t
+#t<-table(df$authortype,df$Gender)
+#t
 # Explicit order of the rows and columns!
 t <- table(
   factor(df$authortype, levels = c("Coauthor", "Other")),
@@ -62,8 +62,8 @@ df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'Other'
 df$authortype[df$author_last == 1] <- 'Last'
 df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
-t<-table(df$authortype,df$Gender)
-t
+#t<-table(df$authortype,df$Gender)
+#t
 # Explicit order of the rows and columns!
 t <- table(
   factor(df$authortype, levels = c("Last", "Other")),

### 2025-08-28 manuscript/manuscript.qmd
updated manuscript (abstract) started discussion. added more to keyword analysis

@@
 
-abstract: "This study examines the gender distribution of authors in music psychology.
+abstract: "Academia suffers from issues regarding equity, with one  crucial aspect being academic authorship.
+Many studies have shown that there is a gender gap in publishing, where women are often underrepresented.
+As there are differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authors within music psychology.
 A total of 3,373 papers published between 2000 and 2025 across five journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
@@
 Overall, female authors were in the minority, accounting for 40.2% of all authors.
-However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
+However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
 These patterns have become more pronounced over the past 15 years.
@@
 Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
-Finally, an examination of keywords suggested broad thematic preferences differing by gender."
+Finally, an examination of keywords suggested broad thematic preferences differing by gender.
+While overall some of this work shows that the field of music psychology is more balanced than some other field, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."
 keywords          : "gender, transparency, music psychology, open science, meta-research"
@@
 From around 1960 to 2021, studies show a general improvement [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender].
-However, there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorships can be unclear.
+Yet there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender].
+There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
 What counts as an authorship may also differ between academic disciplines.
-Women are more likely to experience authorship disagreements [@ni2021gendered].
-There are general disagreements of when to decide authorship - women preferring to discuss in earlier stages, but men choosing authorship at the final stage.
-Another reason could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
+Women are more likely to experience authorship disagreements, for example, when to decide authorship - women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
+Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
 Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
@@
 We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full first name coded with initials, or the second name used as the first name.
-This led to `32` corrections.
+This led to `53` corrections.
 
@@
 For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
-Our analysis focused on the female author proportion (FAP) and odds ratios comparing the likelihood of occupying each authorship position by gender.
+Our analysis focused on the female author proportion (FAP) and odds ratios (ORs) comparing the likelihood of occupying each authorship position by gender.
 
@@
 
-The overall odds ratio for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
+The overall OR for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
 For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
@@
 We aggregated the counts of keywords across all authors in the articles and Figure 4 shows the odds ratios for female authors to be associated (in any author position) with the 40 most frequently used keywords.
-The keywords signifying related to development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
+The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
 At the opposite end of this continuum, topics such as music analysis, machine learning, harmony, cognition, rhythm, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
@@
 ```
+
+# Discussion
+Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
+However, there are several unequities within this aspect of academia, with women often being underrepresented.
+As such representations differ across different disciplines, it is worthwhile understanding how The goal of the current paper was to provide a snapshot of authorship across gender.
+While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
+Overall, we see a relatively encouraging pattern within
 

### 2025-08-28 scripts/keyword_analysis.R

@@ -259,7 +259,6 @@ data = data %>% mutate(
 
 
 # Dance related
-## Words related to music expression
 idx = grep("dance", data$KW)
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -333,7 +332,7 @@ data = data %>% mutate(
                            "tertiary music education" )
                  ~ "music education",TRUE ~ KW))
 
-# religion
+# Religion
 idx = grep("religi", data$KW)
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -384,6 +383,32 @@ data = data %>% mutate(
                            "second language learning")
                  ~ "language",TRUE ~ KW))
 
+# Speech
+idx = grep("speech", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("speech",
+                           "speech-in-noise",
+                           "speech prosody",
+                           "musicalized speech",
+                           "music and speech",
+                           "speech and oral archives",
+                           "expressive speech" ,
+                           "sung speech",
+                           "speech perception",
+                           "joint speech" ,
+                           "speech rhythm",
+                           "perception of speech in noise",
+                           "speech intonation",
+                           "nonpropositional and propositional speech",
+                           "pre-speech development" ,
+                           "private speech",
+                           "speech-to-song" ,
+                           "l2 speech",
+                           "dialogic speech",
+                           "thai speech" )
+                 ~ "speech",TRUE ~ KW))
+
 # Prosocial
 idx = c(grep("prosocial", data$KW),grep("pro-social", data$KW))
 unique(data$KW[idx])
@@ -400,7 +425,7 @@ data = data %>% mutate(
                            "music with prosocial lyrics",
                            "prosocial information"  ,
                            "pro-social behaviour" )
-                 ~ "prosocial",TRUE ~ KW))
+                 ~ "speech",TRUE ~ KW))
 
 # Para social
 idx = grep("parasocial", data$KW)
@@ -566,9 +591,127 @@ data = data %>% mutate(
                            "health musicking" )
                  ~ "music health",TRUE ~ KW))
 
+## Words related to children
+idx = grep("child", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("children's music" ,
+                           "children's musical preferences" ,
+                           "children",
+                           "child development" ,
+                           "middle childhood",
+                           "children's audience" ,
+                           "children's use of singing voice" ,
+                           "children's singing" ,
[…]

### 2025-08-28 scripts/country_table.R
Anna's new improvements taken to ms

@@ -14,11 +14,11 @@ for (k in 1:length(U)) {
   prominent$target_country <- 'Untarget'
   prominent$target_country[prominent$Affiliation_country == U[k]] <- 'Target'
   t<-table(prominent$target_country,prominent$Gender)
-  or <- effectsize::oddsratio(t) 
+  or <- effectsize::oddsratio(t)
   or <- data.frame(or)
   or$Country<- U[k]
   or$N<- sum(t[1,])
-  DATA <- rbind(DATA, or)  
+  DATA <- rbind(DATA, or)
 }
 DATA<-dplyr::select(DATA,Country,N,-CI,Odds_ratio,CI_low,CI_high)
 DATA<-dplyr::arrange(DATA,-N)

### 2025-08-29 manuscript/manuscript.qmd
Added more to keyword analysis

@@
 
-abstract: "Academia suffers from issues regarding equity, with one  crucial aspect being academic authorship.
+abstract: "Academia suffers from issues regarding equity, where academic authorship is one aspect where inequality occurs.
 Many studies have shown that there is a gender gap in publishing, where women are often underrepresented.
@@
 Finally, an examination of keywords suggested broad thematic preferences differing by gender.
-While overall some of this work shows that the field of music psychology is more balanced than some other field, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."
+While overall some of this work shows that the field of music psychology is more balanced than some other fields, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."
+
 keywords          : "gender, transparency, music psychology, open science, meta-research"

### 2025-08-29 scripts/keyword_analysis.R

@@ -91,7 +91,7 @@ if(duplication==TRUE){
 
 
 
-
+# length(unique(data$KW))
 
 ## Words related to music cognition
 idx = grep("cognition", data$KW)
@@ -331,6 +331,50 @@ data = data %>% mutate(
                            "general music education" ,
                            "tertiary music education" )
                  ~ "music education",TRUE ~ KW))
+
+# teaching
+idx = grep("teaching", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("music teaching",
+                           "instrumental teaching and learning",
+                           "teaching beliefs" ,
+                           "instrumental music teaching",
+                           "teaching behaviours",
+                           "teaching styles",
+                           "studio music teaching" ,
+                           "scholarship of teaching and learning",
+                           "teaching methods" ,
+                           "instrumental learning and teaching",
+                           "teaching effectiveness",
+                           "teaching" ,
+                           "piano teaching",
+                           "conceptions of teaching" ,
+                           "teaching experience" ,
+                           "instrumental teaching" ,
+                           "applied teaching",
+                           "instrumental learning/teaching" ,
+                           "peer teaching and learning" ,
+                           "teaching techniques"      )
+                 ~ "teaching",TRUE ~ KW))
+
+# learning
+idx = c( grep("audio-visual", data$KW),  grep("audiovisual", data$KW), grep("audio/v", data$KW))
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("audio-visual",
+                           "audio-visuals",
+                           "audio-visual perception",
+                           "audio-visual processing",
+                           "audio-visual aids",
+                           "audiovisual interaction",
+                           "audiovisual processing" ,
+                           "audiovisual perception",
+                           "audiovisual" ,
+                           "audiovisual integration",
+                           "audio/visual" )
+                 ~ "audio-visual",TRUE ~ KW))
+
 
 # Religion
 idx = grep("religi", data$KW)
@@ -666,7 +710,7 @@ data = data %>% mutate(
                  ~ "audience",TRUE ~ KW))
 
 
-## Words related to audience
+## Words related to concert
 idx = grep("concert", data$KW)
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -687,30 +731,412 @@ data = data %>% mutate(
                            "concert going"  )
                  ~ "concert",TRUE ~ KW))
 
-## Words related to audience
-idx = grep(" sing", data$KW)
+## Words related to memory
[…]

### 2025-09-12 manuscript/manuscript.qmd
Added more to keyword analysis, added to manuscript, added continent analysis, added new references on gender/authorship in different countries, regions and about different topics. Fixed(?) quantify_gender

@@
     affiliation   : "2"
-    address       : "Department of Psychology, University of Toronto, Mississauga, Canada"
+    address       : "Department of Psychology, University of Toronto Mississauga, Canada"
     email         : "a.czepiel@utoronto.ca"
@@
   - id            : "2"
-    institution   : "Department of Psychology, University of Toronto"
+    institution   : "Department of Psychology, University of Toronto Mississauga"
 
@@
   Tuomas Eerola is at the Department of Music, Durham University, UK.
-  Anna Czepiel is at the Department of Psychology, University of Toronto, Canada.
+  Anna Czepiel is at the Department of Psychology, University of Toronto Mississauga, Canada.
 
-abstract: "Academia suffers from issues regarding equity, where academic authorship is one aspect where inequality occurs.
-Many studies have shown that there is a gender gap in publishing, where women are often underrepresented.
-As there are differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authors within music psychology.
-A total of 3,373 papers published between 2000 and 2025 across five journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
-In addition, patterns related to citations, open-access publishing, and keywords were explored across gender.
+abstract: "Academia suffers from various issues regarding equity.
+One aspect is gender inequality in academic authorship, where many studies show that women are often underrepresented.
+As there are authorship differences in across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five journals that represent the core journals in topics of music psychology were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
+In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
 Overall, female authors were in the minority, accounting for 40.2% of all authors.
@@
 Finally, an examination of keywords suggested broad thematic preferences differing by gender.
-While overall some of this work shows that the field of music psychology is more balanced than some other fields, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."
+Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture that hopefully inspires discussions and initiatives to further promote fairness and equity in academia."
 
@@
 
-Academia suffers from issues regarding equity.
-A lack equity impacts productivity, innovation, and job satisfaction in the workplace, which hugely impact academic progress.
-One aspect of this inequity is that gender.
+Academia suffers various equity issues, impacting productivity, innovation, and job satisfaction in the workplace, which-in turn-hugely impact academic progress.
+One aspect of this inequity is gender.
 Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
 For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
-A similar situation holds for italy (24% female professors) [@Filandri03082021], and to some extent for the US (14%) [@spoon2023].
-A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the UK) [@quadlin2023higher].
+A similar situation holds for Italy (24% female professors) [@Filandri03082021], and to some extent for the United States (14%) [@spoon2023].
+A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the United Kingdom) [@quadlin2023higher].
 In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
@@
 From around 1960 to 2021, studies show a general improvement [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender].
-Yet there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender].
+Yet there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
 There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
-What counts as an authorship may also differ between academic disciplines.
-Women are more likely to experience authorship disagreements, for example, when to decide authorship - women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
+What counts as an authorship may differ between academic disciplines.
+Women are more likely to experience authorship disagreements, for example, when to decide authorship: women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
 Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
 Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
+
+-> to add bits about gender authorships and country/continent: 
+- @ioannidis2023gender : compared high-income vs.
+non-high income countries.
+Higher number of authors in high-income countires, where men out-numbered women more so in high-income, though uncertain gender assignment more common in non-high income.
+- @sanchez2024analysis : looked at relationship between human development index (HDI) and female authorships
+- @son2022scientific: high female representation in Africa, lowest in Latin America and Caribbean.
+- @bruck2023bibliometric : proportion of women author varied by country (9.1% in Austria, 0% in South Korea)
+
+-> to add about topics/keywords:
+- @bruck2023bibliometric : in an analysis of medical journals, research keyword varied by gender.
+For men, it was phase II-III trials, oncology, immunotherapy (etc..), while for mena, it was more related to healthcare-related themes like patient involvement, insurance, quality-of-care, access.
+- @sebo2020gender : women more liekly to publish qualitative studies compared to experimental and systematic review articles, perhaps as females demonstrate better relational skill sinvolved in qualitative studies, like prolongeed relationships and emotional ties with research participants.
+Women were also more likely to publish in primary healthcare journals.
+- @ashmos2011gendering : men are under-represented, but women are over-represented in qualitative studies. 
+- @thelwall2019gender : females more likely to use exploratory and qualitative rather than quantitative methods.
+Females more interested in cell biology and veterinary science compared to males, who prefer abstraction, patients, and power/control field
+- @vogel2012he : in computer science (computational linguistics), used XX to look at topics that female / male authors published on.
+Female more likely to publish on dialog, discourse, and sentiment, while men more likely to publish in formal semantic and finite state models.
 
@@
 Gender is a complex and multidimensional social construct.
-However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
+However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and that non-binary authorship generally is less than 0.1% [@son2022scientific], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
 We used the _genderize API_ [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
@@
 These `59` cases were excluded from the dataset.
-It is likely that the due to challenge of attributing gender correctly not all the gender attributions are correct.
+It is likely that the due to challenge of attributing gender accurately not all the gender attributions are correct.
 Although genderize.io has been shown to achieve 96.6% accuracy in a diverse multinational test database without using a country of origin information -- 98% with the information included -- it also underperforms (82% accuracy) in Asian names [@vanhelene2024inferring].
@@
 
-The overall OR for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
-For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
-In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
-Being a single author has odds ratio of 0.90 [0.78-1.05] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not significant (_p_ = .21).
+The overall OR for women as first authors was 1.408 [95% CI: 1.284–1.544], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
+For coauthorship positions, the OR was 0.997 [0.913–1.090], suggesting no substantial gender difference (_p_ = .96).
+In contrast, the odds ratio for women as last authors was 0.733 [0.667–0.806], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
+Being a single author has odds ratio of 0.905 [0.778-1.052] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not significant (_p_ = .21).
 
@@
 The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
-At the opposite end of this continuum, topics such as music analysis, machine learning, harmony, cognition, rhythm, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+At the opposite end of this continuum, topics such as music analysis, machine learning, harmony, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
 
@@
 Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
-However, there are several unequities within this aspect of academia, with women often being underrepresented.
+However, there are several inequities within this aspect of academia, with women often being underrepresented.
 As such representations differ across different disciplines, it is worthwhile understanding how The goal of the current paper was to provide a snapshot of authorship across gender.

### 2025-09-12 scripts/country_table.R

@@ -1,5 +1,6 @@
 # Country_table.R
 
+
 # analyse only countries with N observations or more
 country_counts <- table(df$Affiliation_country)
 index <- country_counts > 60
@@ -10,7 +11,7 @@ prominent <- df[df$Affiliation_country %in% names(index)[index==TRUE],]
 U<-unique(prominent$Affiliation_country)
 
 DATA <- NULL
-for (k in 1:length(U)) {
+for (k in 1:length(U)) { 
   prominent$target_country <- 'Untarget'
   prominent$target_country[prominent$Affiliation_country == U[k]] <- 'Target'
   t<-table(prominent$target_country,prominent$Gender)
@@ -36,3 +37,50 @@ g1<-ggplot(DATA,aes(x=reorder(Country,Odds_ratio),y=Odds_ratio,label=paste0("n="
   ylab('Female Authorship Odds Ratio (95%CI)') +
   coord_flip()
 print(g1)
+
+
+
+## By continent? 
+
+
+# Assign continents
+library(countrycode) 
+df$assigned_continent = countrycode(df$Affiliation_country, origin = "country.name", destination = "continent")
+
+
+# analyse only countries with N observations or more
+continent_counts <- table(df$assigned_continent)
+#table(prominent$Affiliation_country)
+#length(table(prominent$Affiliation_country))
+continent_count = df 
+
+U<-unique(continent_count$assigned_continent)
+
+
+DATA <- NULL
+for (k in 1:length(U)) { # k = 1
+  continent_count$target_continent <- 'Untarget'
+  continent_count$target_continent[continent_count$assigned_continent == U[k]] <- 'Target'
+  t<-table(continent_count$target_continent,continent_count$Gender)
+  or <- effectsize::oddsratio(t)
+  or <- data.frame(or)
+  or$Continent<- U[k]
+  or$N<- sum(t[1,])
+  DATA <- rbind(DATA, or)
+}
+DATA<-dplyr::select(DATA,Continent,N,-CI,Odds_ratio,CI_low,CI_high)
+DATA<-dplyr::arrange(DATA,-N)
+#print(knitr::kable(DATA,digits = 2,caption='Make this into two-column table in future.'))
+
+#head(DATA)
+g2<-ggplot(DATA,aes(x=reorder(Continent,Odds_ratio),y=Odds_ratio,label=paste0("n=",N)))+
+  geom_point(shape=15)+
[…]

### 2025-09-12 scripts/keyword_analysis.R

@@ -88,7 +88,7 @@ if(duplication==TRUE){
 }
 
 
-
+kw_num = data %>% count(KW)
 
 
 # length(unique(data$KW))
@@ -254,6 +254,14 @@ data = data %>% mutate(
   KW = case_when(KW %in% c("new interfaces for musical expression",
                            "new interface for musical expression" )
                  ~ "new interface for musical expression",TRUE ~ KW))
+
+idx = grep("absolute", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("absolute pitch",
+                           "absolute pitch shift",
+                           "absolute pitch (ap)")
+                 ~ "absolute pitch",TRUE ~ KW))
 
 
 
@@ -955,7 +963,7 @@ data = data %>% mutate(
                             "cross-modal integration")
                  ~ "cross-modal",TRUE ~ KW))
 
-## Words related to computing
+## Words related to evolution
 idx = grep("evolution", data$KW)
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -974,7 +982,7 @@ data = data %>% mutate(
                  ~ "evolution",TRUE ~ KW))
 
 
-## Words related to computing
+## Words related to genetics
 idx = grep("genet", data$KW)
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -1072,7 +1080,7 @@ data = data %>% mutate(
                            "expertise development")
                  ~ "expertise",TRUE ~ KW))
 
-## Words related to predictability
+## Words related to  modeling
 idx = c(grep("modeling", data$KW),grep("modelling", data$KW))
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -1113,18 +1121,8 @@ data = data %>% mutate(
 
 
 
-## Words related to 
+## Words related to individual differences
 idx = c(grep("individual", data$KW))
-unique(data$KW[idx])
-data = data %>% mutate(
-  KW = case_when(KW %in% c("individual differences",
-                           "individual difference",
-                           "individual_differences")
-                 ~ "individual differences",TRUE ~ KW))
-
-
-## Words related to 
-idx = c(grep("analysis", data$KW))
 unique(data$KW[idx])
 data = data %>% mutate(
   KW = case_when(KW %in% c("individual differences",
@@ -1183,6 +1181,13 @@ data = data %>% mutate(
 head(data)
[…]

### 2025-09-12 scripts/quantify_authorship.R

@@ -77,7 +77,7 @@ or4 <- effectsize::oddsratio(t) # 0.73
 or <- rbind(or1,or2,or3,or4)
 or$name<-c("Single","First","Coauthor","Last")
 or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
-print(knitr::kable(or,digits = 2))
+print(knitr::kable(or,digits = 3))
 
 #### Annually ---------
 
@@ -98,9 +98,13 @@ for (k in 1:length(U)){
   tmp<-dplyr::filter(df,YEAR_5==U[k])
   #t <- table(tmp$authortype,tmp$Gender)
   # Explicit order of the rows and columns!
+  # t <- table(
+  #   factor(df$authortype, levels = c("Single", "Other")),
+  #   factor(df$Gender,levels = c("female","male"))
+  # )
   t <- table(
-    factor(df$authortype, levels = c("Single", "Other")),
-    factor(df$Gender,levels = c("female","male"))
+    factor(tmp$authortype, levels = c("Single", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
   )
   output <- rbind(output,effectsize::oddsratio(t))
 }
@@ -120,9 +124,13 @@ for (k in 1:length(U)){
   tmp<-dplyr::filter(df,YEAR_5==U[k])
   #t <- table(tmp$authortype,tmp$Gender)
   # Explicit order of the rows and columns!
+  # t <- table(
+  #   factor(df$authortype, levels = c("First", "Other")),
+  #   factor(df$Gender,levels = c("female","male"))
+  # )
   t <- table(
-    factor(df$authortype, levels = c("First", "Other")),
-    factor(df$Gender,levels = c("female","male"))
+    factor(tmp$authortype, levels = c("First", "Other")),
+    factor(tmp$Gender,levels = c("female","male"))
   )
   output <- rbind(output,effectsize::oddsratio(t))
 }
@@ -142,9 +150,13 @@ for (k in 1:length(U)){
   tmp<-dplyr::filter(df,YEAR_5==U[k])
 #  t <- table(tmp$authortype,tmp$Gender)
   # Explicit order of the rows and columns!
+  # t <- table(
+  #   factor(df$authortype, levels = c("Coauthor", "Other")),
+  #   factor(df$Gender,levels = c("female","male"))
+  # )
   t <- table(
-    factor(df$authortype, levels = c("Coauthor", "Other")),
[…]

### 2025-09-12 analysis.qmd
Revised figure 2, removed unnecessary plots, added discussion to MS

@@
-# Analyse gender in music psychology
+---
+title: Analyse gender in music psychology from journal publication data
+date: "2025-08-01"
+author:
+  - "Tuomas Eerola"
+  - "Anna Czepiel"
+format:
+  html:
+    toc: true
+    number-sections: true
+    theme: materia
+---
 
-T. Eerola, 1/8/2025
-
-Gender analysis of the music psychology journals since 2000 (Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science).
-
-## Required libraries
+Gender analysis of the authors in music psychology journals since 2000.
+The journals included are _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_ and the data been obtained in June 2025 from Scopus.
 
@@
 
-## Load and preprocess data
+## Diagnostics
 
@@
 
-## Quantify
+## Quantify authorship types, citations and open access
 
@@
 source('scripts/country_table.R') # OK
-source('scripts/visualise_gender.R') # OK
+#source('scripts/visualise_gender.R') # OK
 
@@
 #| fig.height: 6
+source('scripts/count2category.R') 
 source('scripts/keyword_analysis.R') 

### 2025-09-12 manuscript/_figure2.qmd
Revised figure 2, removed unnecessary plots, added discussion to MS

@@
 #| label: fig2
-#| fig-width: 9
+#| fig-width: 13
 #| fig-height: 4
@@
 #| fig.cap: > 
-#|   Authorship types across author counts.
+#|   A) Number of authors and B) female odds ratio for authorship types across the number of authors.
 
@@
   scale_color_manual(name="Authorship\ntype",values = custom_palette) +
-  xlab('Author N')+
+  xlab('Number of authors/article')+
   ylab('Female Authorship Odds Ratio')+
   theme_classic(base_size = 15)
-print(fig2)
+#print(fig2)
+
+s <- summarise(group_by(df, BIBTEXKEY), n = n())
+S<-s
+S$n[S$n > 9]<-10
+fig2a <- ggplot(S,aes(x=n))+
+  geom_histogram(binwidth=1,fill='grey90',color='black')+
+  theme_classic(base_size = 15)+
+  labs(x='Number of authors/article',y='Count of articles')+
+  scale_x_continuous(breaks=seq(1,10,1),expand = c(0.001,0.001),limits = c(0,11),
+                     labels = c(seq(1,9,1),"10+"))+
+  scale_y_continuous(expand = c(0.001,0.001))
+#print(g)
+
+G<-cowplot::plot_grid(fig2a,fig2,nrow=1,labels=c('A','B'),rel_widths = c(0.4,1))
+print(G)
+
 ```

### 2025-09-12 manuscript/manuscript.qmd
Revised figure 2, removed unnecessary plots, added discussion to MS

@@
 
--> to add about topics/keywords:
+-> to add about topics/keywords/approaches/methods:
 - @bruck2023bibliometric : in an analysis of medical journals, research keyword varied by gender.
-For men, it was phase II-III trials, oncology, immunotherapy (etc..), while for mena, it was more related to healthcare-related themes like patient involvement, insurance, quality-of-care, access.
-- @sebo2020gender : women more liekly to publish qualitative studies compared to experimental and systematic review articles, perhaps as females demonstrate better relational skill sinvolved in qualitative studies, like prolongeed relationships and emotional ties with research participants.
+For men, it was phase II-III trials, oncology, immunotherapy (etc..), while for women, it was more related to healthcare-related themes like patient involvement, insurance, quality-of-care, access.
+- @sebo2020gender : women more liekly to publish qualitative studies compared to experimental and systematic review articles, perhaps as females demonstrate better relational skill involved in qualitative studies, like prolongeed relationships and emotional ties with research participants.
 Women were also more likely to publish in primary healthcare journals.
@@
 
-It is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
+As the topics, and approaches and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
 For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%).
@@
 To improve certain academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
-The past efforts have already provided a snapshot a rough overview of the state of affairs (e.g., @anglada2019visualizing) with suggestions of future directions to improve the field.
-[@eerola2023transparency] showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
+The past efforts such as @anglada2019visualizing have already provided a snapshot a rough overview of the state of affairs with suggestions of future directions to improve the field.
+@eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
 @jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is, which the view to see how – as a field – we need to diversify participant samples and stimuli to gain a more realistic understanding.
@@
 The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
-At the opposite end of this continuum, topics such as music analysis, machine learning, harmony, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+At the opposite end of this continuum, topics such as music analysis, harmony, machine learning, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
 
@@
 # Discussion
+
 Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
 However, there are several inequities within this aspect of academia, with women often being underrepresented.
-As such representations differ across different disciplines, it is worthwhile understanding how The goal of the current paper was to provide a snapshot of authorship across gender.
+As such representations differ across different disciplines, it is worthwhile to diagnose the situation in music psychology.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
 While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
-Overall, we see a relatively encouraging pattern within
+
+In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
+This proportion is higher than in musicology (34.8% [@eerola2026inprep]) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
+In terms of authorship roles, notable differences emerge.
+Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
+By contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+Given that last authorship in the social sciences is not universally defined but is increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution], this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
+This pattern is consistent with broader evidence showing that women remain underrepresented in senior academic positions, including professorial appointments [@bakker2016tenure], invitations to contribute to journals, and authorship in high-prestige outlets [@holman2018gender].
+Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
+
+There were substantial gender differences in citation patterns.
+One notable finding is that first-authored articles by women were significantly more likely to be published open access, with an odds ratio of 1.49, a result that warrants further exploration.
+This effect does not appear to be tied to specific journals.
+Although the odds ratio for women publishing open access is highest in _Music and Science_, the broader trend suggests a temporal shift: since around 2010, women have increasingly overtaken men in publishing open access.
+The data does not provide any reasons for this.
+While past surveys among UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
+
+Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
+Our summaries of gender patterns across continents and countries indicate clear cross-national variation.
+Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
+This pattern, however, is not universal across all countries examined [@chan2020].
+Prior work shows that women’s research success is positively correlated with national indicators of gender equity [@chan2020].
+This is likely to be related to a career length, which strongly correlates with overall productivity.
+Women exhibit higher dropout rates than men, particularly during the 5–10 years following their first publication [@huang2020], but policies geared towards equity and support for career breaks and caring responsibilities may well contribute towards the observed country-specific differences.
+Nevertheless, prior analyses demonstrate that even when women constitute a smaller share of researchers within a discipline or country, their contributions are often more impactful than those of their male colleagues [@chan2020].
+When examining countries where women have significantly lower odds of authorship compared to men (Japan, Belgium, Austria, Sweden, and Turkey), it is not straightforward to conclude that these nations fall consistently at the lower end of gender equality rankings.
+The notable exception is Turkey, which is positioned lower according to the Gender Inequality Index [@UNDPHDR2024].
+
+The keyword analysis of the articles indicated thematic differences in the topics studied by women and men.
+Research areas related to children, well-being, health, and mental health were more frequently associated with women.
+This pattern is consistent with observations in psychology more broadly, where women are more strongly represented in sub-disciplines that focus on care and nurturing [@gonzalez2020women].
+The higher likelihood of women studying topics related to singing may be connected to prevailing gender stereotypes, or to the early negative perceptions of singing among boys [@Warzecha2013] and men [@palkki2015gender].
+In contrast, topics related to cognition and computation were more often associated with men, which reflects a wider trend in psychology [@gonzalez2020women].
 

### 2025-09-12 scripts/OA_advantage.R

@@ -1,0 +1,48 @@
+# explore_OA_advantage.R
+
+
+#### Classify order: First author ----------
+df$authortype <- 'Other'
+df$authortype[df$author_order == 1] <- 'First'
+df$authortype[df$author_last == 1] <- 'Other'
+df$authortype[df$author_single == 1] <- 'Other' # Take out single authors from first authors!
+# t<-table(df$authortype,df$Gender)
+
+table(df$authortype,df$Gender)
+
+# Explicit order of the rows and columns!
+t <- table(
+  factor(df$authortype, levels = c("First", "Other")),
+  factor(df$Gender,levels = c("female","male"))
+)
+t
+print(effectsize::oddsratio(t)) # 1.41
+
+# Specific journals
+journal <- "Journal of New Music Research"
+t <- table(
+  factor(df$authortype[df$JOURNAL==journal], levels = c("First", "Other")),
+  factor(df$Gender[df$JOURNAL==journal],levels = c("female","male"))
+)
+t
+print(effectsize::oddsratio(t)) # 1.37
+journal <- "Psychology of Music" # 1.47
+journal <- "Music and Science" # 1.75
+journal <- "Music Perception" # 1.26
+journal <- "Musicae Scientiae" # 1.48
+
+
+table(df$JOURNAL,df$authortype)
+table(df$JOURNAL,df$OA)
+table(df$YEAR,df$OA)
+
+
+t <- table(
+  factor(df$authortype[df$YEAR>2000 & df$YEAR<2010], levels = c("First", "Other")),
+  factor(df$Gender[df$YEAR>2000 & df$YEAR<2010],levels = c("female","male"))
+)
+t
+print(effectsize::oddsratio(t)) # 1.51
+# 2019: 1.51
+# 2010-2019: 1.47
+# 2000-2009: 1.16

### 2025-09-12 scripts/attribute_gender_diagnostics.R

@@ -15,9 +15,9 @@ journal_with_total <- df %>%
   count(JOURNAL) %>%
   mutate(percentage = round(n / sum(n) * 100, 1)) %>%
   bind_rows(
-    summarise(., 
-              JOURNAL = "Total", 
-              n = sum(n), 
+    summarise(.,
+              JOURNAL = "Total",
+              n = sum(n),
               percentage = sum(percentage))
   )
 
@@ -31,15 +31,17 @@ cat("\n\n")
 #hist(df$Gender.Probability)
 
 # the majority is confidently attributed
-cat(paste("Prop. over .90 conf. =",sum(df$Gender.Probability>.90)/nrow(df),"/n"))
-cat(paste("Prop. over .95 conf. =",sum(df$Gender.Probability>.95)/nrow(df),"/n"))
-cat(paste("Prop. under .55 conf. =",sum(df$Gender.Probability<.55)/nrow(df),"/n"))
-cat(paste("Prop. under .55 conf. =",sum(df$Gender.Probability<.55),"/n/n"))
+cat("Confidence of gender attribution by the API:\n\n")
+cat(paste("Prop. over .90 conf. =",round(sum(df$Gender.Probability>.90)/nrow(df),3),"\n\n"))
+cat(paste("Prop. over .95 conf. =",round(sum(df$Gender.Probability>.95)/nrow(df),3),"\n\n"))
+cat(paste("Prop. under .55 conf. =",round(sum(df$Gender.Probability<.55)/nrow(df),3),"\n\n"))
+cat(paste("Prop. under .55 conf. =",round(sum(df$Gender.Probability<.55)/nrow(df),3),"\n\n"))
 
 cat("\n***")
 
 S <- dplyr::summarise(group_by(df,Gender),M=mean(Gender.Probability),Md=median(Gender.Probability))
-print(knitr::kable(S,digits = 2))
+print(knitr::kable(S,digits = 2,caption="API: Gender probability means/medians."))
+
 cat("\n***")
 
 upperCI<-function(x, conf.level = 0.95) {
@@ -56,11 +58,11 @@ lowerCI<-function(x, conf.level = 0.95) {
 }
 # citations across gender
 S <- dplyr::summarise(group_by(df,Gender),M=mean(Citations),Md=median(Citations),LCI=lowerCI(Citations),UCI=upperCI(Citations))
-print(knitr::kable(S,digits = 2))
+print(knitr::kable(S,digits = 2,caption="Citations across gender."))
 
[…]

### 2025-09-12 scripts/country_table.R

@@ -11,7 +11,7 @@ prominent <- df[df$Affiliation_country %in% names(index)[index==TRUE],]
 U<-unique(prominent$Affiliation_country)
 
 DATA <- NULL
-for (k in 1:length(U)) { 
+for (k in 1:length(U)) {
   prominent$target_country <- 'Untarget'
   prominent$target_country[prominent$Affiliation_country == U[k]] <- 'Target'
   t<-table(prominent$target_country,prominent$Gender)
@@ -40,11 +40,11 @@ print(g1)
 
 
 
-## By continent? 
+## By continent?
 
 
 # Assign continents
-library(countrycode) 
+library(countrycode)
 df$assigned_continent = countrycode(df$Affiliation_country, origin = "country.name", destination = "continent")
 
 
@@ -52,7 +52,7 @@ df$assigned_continent = countrycode(df$Affiliation_country, origin = "country.na
 continent_counts <- table(df$assigned_continent)
 #table(prominent$Affiliation_country)
 #length(table(prominent$Affiliation_country))
-continent_count = df 
+continent_count = df
 
 U<-unique(continent_count$assigned_continent)
 

### 2025-09-12 scripts/keyword_analysis.R

@@ -676,7 +676,7 @@ data = data %>% mutate(
                            "parent-child interactions" )
                  ~ "children",TRUE ~ KW))
 
-## Words related to infants / toddler 
+## Words related to infants / toddler
 idx = c( grep("infant", data$KW), grep("toddler", data$KW) )
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -827,7 +827,7 @@ data = data %>% mutate(
                            "play from memory"  ,
                            "memorization"  ,
                            "context-dependent memory" ,
-                           "memory span" )        
+                           "memory span" )
                  ~ "concert",TRUE ~ KW))
 
 
@@ -940,7 +940,7 @@ data = data %>% mutate(
                            "computer software design" )
                  ~ "computation",TRUE ~ KW))
 
-## Words related to cross modal 
+## Words related to cross modal
 idx = c(grep("crossmodal", data$KW),grep("cross-modal", data$KW))
 unique(data$KW[idx])
 data = data %>% mutate(
@@ -969,7 +969,7 @@ unique(data$KW[idx])
 data = data %>% mutate(
   KW = case_when(KW %in% c("human evolution",
                            "cultural evolution",
-                           "evolution" , 
+                           "evolution" ,
                            "evolution of music",
                            "hominin evolution" ,
                            "evolutionary optimisation" ,
@@ -1103,7 +1103,7 @@ data = data %>% mutate(
                            "biomechanical modelling",
                            "item response modelling",
                            "performance modelling" ,
-                           "melodic modelling",   
+                           "melodic modelling",
                            "segmentation modelling" ,
                            "structural equation modelling (sem)",
                            "structural equation modelling")
@@ -1185,7 +1185,7 @@ data<-dplyr::filter(data,KW!="music psychology")
 
 
 
-#kw_num = data %>% count(KW) %>% arrange(n) 
+#kw_num = data %>% count(KW) %>% arrange(n)
 kw_num = data %>% count(KW) %>% arrange(desc(n))
 
 
@@ -1227,10 +1227,10 @@ figure1 <- ggplot(data = x1[FROM:TO, ], aes(
[…]

### 2025-09-12 scripts/quantify_authorship.R

@@ -18,8 +18,8 @@ t <- table(
 )
 t
 or1 <- effectsize::oddsratio(t) # 0.83       | [0.71, 0.96]
-print(or1)
-fisher.test(t)
+#print(or1)
+#fisher.test(t)
 
 #### Classify order: First author ----------
 df$authortype <- 'Other'
@@ -36,8 +36,8 @@ t <- table(
 t
 
 or2 <- effectsize::oddsratio(t) # 1.27       | [1.03, 1.56]
-print(or2)
-fisher.test(t)
+#print(or2)
+#fisher.test(t)
 
 #### Classify order: Coauthor --------------
 df$authortype <- 'Coauthor'
@@ -54,8 +54,8 @@ t <- table(
 t
 
 or3 <- effectsize::oddsratio(t) # 1.06       | [0.77, 1.46]
-print(or3)
-fisher.test(t)
+#print(or3)
+#fisher.test(t)
 
 #### Classify order: Last -------------
 df$authortype <- 'Other'
@@ -77,7 +77,7 @@ or4 <- effectsize::oddsratio(t) # 0.73
 or <- rbind(or1,or2,or3,or4)
 or$name<-c("Single","First","Coauthor","Last")
 or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
-print(knitr::kable(or,digits = 3))
+print(knitr::kable(or,digits = 3,caption = "Female authorship odds ratios by authorship type."))
 
 #### Annually ---------
 
@@ -201,10 +201,10 @@ annual
 # Add overall numbers
 or$name = factor(or$name, levels = c('Single','First','Coauthor', 'Last'), labels = c('Single','First','Co', 'Last'))
 
-tmp <- data.frame(Odds_ratio = or$Odds_ratio, 
+tmp <- data.frame(Odds_ratio = or$Odds_ratio,
                   CI=0.95,
-                  CI_low = or$CI_low, 
-                  CI_high = or$CI_high, 
+                  CI_low = or$CI_low,
+                  CI_high = or$CI_high,
                   YearRange = "Overall",
                   Type = or$name)
 annual2<- rbind(tmp,annual)
@@ -292,10 +292,10 @@ last_gr <- last %>%
 last_gr
 last_gr_AAGR <- mean(last_gr$Rate_percent, na.rm = TRUE)
 
-print('5-year growth rate of female authorships:')
+#print('5-year growth rate of female authorships:')
[…]

### 2025-09-12 scripts/summarise_gender.R

@@ -10,9 +10,9 @@ gender_with_total <- df %>%
   count(Gender) %>%
   mutate(percentage = round(n / sum(n) * 100, 1)) %>%
   bind_rows(
-    summarise(., 
-              Gender = "Total", 
-              n = sum(n), 
+    summarise(.,
+              Gender = "Total",
+              n = sum(n),
               percentage = sum(percentage))
   )
 
@@ -42,4 +42,19 @@ cat(paste("\n mean:",round(mean(s$n),3)))
 cat(paste("\n sd:",round(sd(s$n),3)))
 cat(paste("\n max:",round(max(s$n),3)))
 cat("\n")
+s$n
+# create a histogram
+S<-s
+S$n[S$n > 9]<-10
+g<-ggplot(S,aes(x=n))+
+  geom_histogram(binwidth=1,fill='grey90',color='black')+
+  theme_classic()+
+  labs(x='Number of authors',y='Count of articles')+
+  scale_x_continuous(breaks=seq(1,10,1),expand = c(0.001,0.001),limits = c(0,11),
+                     labels = c(seq(1,9,1),"10+"))+
+  scale_y_continuous(expand = c(0.001,0.001))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=14))
+print(g)
+
 rm(s,gender_with_total)

### 2025-09-12 scripts/country_table.R
added eminent author idea, revised figure 3 (merged two plots)

@@ -34,9 +34,10 @@ g1<-ggplot(DATA,aes(x=reorder(Country,Odds_ratio),y=Odds_ratio,label=paste0("n="
   theme_classic(base_size = 14)+
   geom_hline(yintercept = 1,linetype='dashed',color='grey20')+
   xlab('Country') +
-  ylab('Female Authorship Odds Ratio (95%CI)') +
+  ylab('') +
+  scale_y_continuous(limits=c(0,2))+
   coord_flip()
-print(g1)
+#print(g1)
 
 
 
@@ -80,7 +81,11 @@ g2<-ggplot(DATA,aes(x=reorder(Continent,Odds_ratio),y=Odds_ratio,label=paste0("n
   geom_text(nudge_x = .30, size=2.25)+
   theme_classic(base_size = 14)+
   geom_hline(yintercept = 1,linetype='dashed',color='grey20')+
-  xlab('Country') +
+  scale_y_continuous(limits=c(0,2))+
+  xlab('Continent') +
   ylab('Female Authorship Odds Ratio (95%CI)') +
   coord_flip()
-print(g2)
+#print(g2)
+
+G <- cowplot::plot_grid(g1,g2,ncol=1,labels=c('A','B'),rel_heights = c(1,0.35),axis = "tblr",align = "hv")
+print(G)

### 2025-09-12 scripts/eminent_author_per_country_weights.R

@@ -1,0 +1,34 @@
+# eminent_author_per_country_weights.R
+#
+# Idea: To understand how much individuals contribute to the
+# country-specific stats, let's calculate the proportion of
+# the authorship the the top N authors carry in each country.
+
+U<-unique(prominent$Affiliation_country)
+U
+country_counts <- table(df$Affiliation_country)
+index <- country_counts > 60
+prominent <- df[df$Affiliation_country %in% names(index)[index==TRUE],]
+
+# Get detailed breakdown showing the top 3 names
+S <- prominent %>%
+  count(Affiliation_country, unique_name, name = "count") %>%
+  group_by(Affiliation_country) %>%
+  mutate(
+    total_country = sum(count),
+    proportion = count / total_country,
+    rank = rank(-count, ties.method = "first")
+  ) %>%
+  filter(rank <= 3) %>%
+  arrange(Affiliation_country, rank)
+
+  # Optional: create summary with top 3 proportion
+S %>%
+  group_by(Affiliation_country) %>%
+  mutate(
+    top3_proportion = sum(proportion)
+  ) %>%
+  ungroup() -> S2
+S2
+
+print(S2, n = 72)

### 2025-09-12 manuscript/manuscript.qmd
added eminent author idea, revised figure 3 (merged two plots)

@@
 #| fig-width: 9
-#| fig-height: 6
+#| fig-height: 8
 #| fig.cap: > 
@@
 ```
-
-- If we want to add the continents at this figure, that's possible.
 

### 2025-09-16 manuscript/manuscript.qmd
Converted bullet point draft/notes on geographical location and keyword/topic analysis into paragraph in manuscript intro

@@
 One aspect is gender inequality in academic authorship, where many studies show that women are often underrepresented.
-As there are authorship differences in across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
-A total of 3,373 papers published between 2000 and 2025 across five journals that represent the core journals in topics of music psychology were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
+Due to authorship differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
 In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
@@
 Academia suffers various equity issues, impacting productivity, innovation, and job satisfaction in the workplace, which-in turn-hugely impact academic progress.
-One aspect of this inequity is gender.
+Although there are several aspects, the focus of this paper is gender inequity.
 Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
@@
 
-Research across several scientific disciplines that show that women tend to be less represented in authors compared to men [@banks2025women; @son2022scientific].
+Research across several scientific disciplines show that women tend to be less represented as authors compared to men [@banks2025women; @son2022scientific].
 Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving.
 From around 1960 to 2021, studies show a general improvement [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender].
-Yet there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
+Yet, there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
 There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
@@
 
--> to add bits about gender authorships and country/continent: 
-- @ioannidis2023gender : compared high-income vs.
-non-high income countries.
-Higher number of authors in high-income countires, where men out-numbered women more so in high-income, though uncertain gender assignment more common in non-high income.
-- @sanchez2024analysis : looked at relationship between human development index (HDI) and female authorships
-- @son2022scientific: high female representation in Africa, lowest in Latin America and Caribbean.
-- @bruck2023bibliometric : proportion of women author varied by country (9.1% in Austria, 0% in South Korea)
+Taking a closer look at authorships, papers have explored certain aspects for such differences<to improve this last sentence to work way into this paragraph better>.
+One of these is different geographical regions.
+For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
+In physics journals, female representation was lowest for Latin America and the Caribbean (16%), but highest for Africa (19%) [@son2022scientific].
+In medical journals, women authors were least common in publications originating from South Korea, Japan, and Austria, but most common in South Africa and India [@bruck2023bibliometric].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship: Although female contribution to scientific output is lower in countries ranked in lowest quartile for human development and gender equity, this also occured in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
 
--> to add about topics/keywords/approaches/methods:
-- @bruck2023bibliometric : in an analysis of medical journals, research keyword varied by gender.
-For men, it was phase II-III trials, oncology, immunotherapy (etc..), while for women, it was more related to healthcare-related themes like patient involvement, insurance, quality-of-care, access.
-- @sebo2020gender : women more liekly to publish qualitative studies compared to experimental and systematic review articles, perhaps as females demonstrate better relational skill involved in qualitative studies, like prolongeed relationships and emotional ties with research participants.
-Women were also more likely to publish in primary healthcare journals.
-- @ashmos2011gendering : men are under-represented, but women are over-represented in qualitative studies. 
-- @thelwall2019gender : females more likely to use exploratory and qualitative rather than quantitative methods.
-Females more interested in cell biology and veterinary science compared to males, who prefer abstraction, patients, and power/control field
-- @vogel2012he : in computer science (computational linguistics), used XX to look at topics that female / male authors published on.
-Female more likely to publish on dialog, discourse, and sentiment, while men more likely to publish in formal semantic and finite state models.
+Another aspect that differs between distribution of gender in authorship is the research methodologies and topic.
+In an analyses of article keywords of medical journals, male-authored articles were more associated with keywords like phase II-III trials, oncology, immunotherapy, while female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care [@bruck2023bibliometric].
+In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+These trends could be explained by the idea that female researchers often value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+In terms of methodologies, women are are more likely to publish qualitative than quantitative/experimental studies [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
+Interestingly, such stereotypes are occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
+Overall, women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
 
-As the topics, and approaches and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
+As the topics, approaches, and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
 For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%).
@@
 To improve certain academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
-The past efforts such as @anglada2019visualizing have already provided a snapshot a rough overview of the state of affairs with suggestions of future directions to improve the field.
+The past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
 @eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
@@
 
+[@ashmos2011gendering; @sebo2020gender; @thelwall2019gender], In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+
 # Conclusions

### 2025-09-18 manuscript/manuscript.qmd
Added text to conclusions and to the results

@@
 
-Taking a closer look at authorships, papers have explored certain aspects for such differences<to improve this last sentence to work way into this paragraph better>.
-One of these is different geographical regions.
+Past studies have explored the gender distribution across a range of elements that could illuminate the reasons for the discrepancies.
+One of these is geographical regions.
 For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
@@
 As the topics, approaches, and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
-For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%).
+For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%), but not equal.
 In psychology, which is perhaps the closest benchmark for the present focus on music psychology, female authors accounted for 45.2% of the sample (N = 90,067).
 
-To improve certain academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
+To improve academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
 The past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
 @eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
-@jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is, which the view to see how – as a field – we need to diversify participant samples and stimuli to gain a more realistic understanding.
-In a similar way to these papers, the current paper aims to give a current overview of authorship and gender patterns in the field of music psychology.
+@jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is.
+In a similar way to these papers, the current paper aims to give a current overview of gender patterns in authorship in the field of music psychology.
 
@@
 Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
-What is the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
-Are there specific trends in terms of countries of the affiliations, topics, or time?
+More specifically, we want to identify what is the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
+Are there specific patterns in the authorship distributions across countries, topics of the studies, or temporal trends?
 
@@
 For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
-Our analysis focused on the female author proportion (FAP) and odds ratios (ORs) comparing the likelihood of occupying each authorship position by gender.
+Our analysis focused on the female author proportion (FAP) and utilised odds ratios (ORs) to compare the likelihood of occupying each authorship position by gender.
 
@@
 
+Panel A in Figure 3 shows a small number of countries, such as Australia, China, and Canada, with a high odds ratio for female authorship.
+At the other end of the spectrum, Japan, Belgium, Austria, Sweden, and Turkey all have significantly lower odds ratios for female authors.
+There is no simple rationale immediately apparent (e.g., country size, membership in a particular political bloc, gender equality index, or domination of articles by a single eminent scholar from one country) that would explain these differences, but this representation provides some starting points for discussion and may highlight the impact of academic career structures and principles implemented in these countries.
+Panel B shows data aggregated by continent, where Africa has meagre representation (52 authors), Oceania consists only of Australia and New Zealand, and the Americas are represented almost entirely by the US and Canada.
+While this level of presentation may be too coarse for detailed analysis, it does offer a broad summary of the current situation.
+
 ## Gendered topics through analysis of keywords
 
-To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (15,316 in total, 6730 unique).
+To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (15,316 in total, 6,730 unique).
 To reduce variant keywords, we manually simplified the variants (e.g., "arts in health", "healthy music use", "health musicking" were converted to "music health" and "music cognition" to "cognition" and "music perception" to "perception") and eliminated single use of the term "music".
@@
 At the opposite end of this continuum, topics such as music analysis, harmony, machine learning, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+There is much nuance and detail in the data that would warrant separate analyses; however, the purpose here is simply to provide reference points for comparing gender distributions of authorship with other disciplines, where thematic analyses have already been conducted (e.g., in sub-disciplines of psychology).
 
@@
 
-There were substantial gender differences in citation patterns.
-One notable finding is that first-authored articles by women were significantly more likely to be published open access, with an odds ratio of 1.49, a result that warrants further exploration.
+There were no substantial gender differences in the citation patterns.
+However, one notable observation is that first-authored articles by women were significantly more likely to be published open access, with an odds ratio of 1.49, a result that warrants further exploration.
 This effect does not appear to be tied to specific journals.
 Although the odds ratio for women publishing open access is highest in _Music and Science_, the broader trend suggests a temporal shift: since around 2010, women have increasingly overtaken men in publishing open access.
-The data does not provide any reasons for this.
-While past surveys among UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
+The data does not provide any direct reasons for this, although the legislation in few of the European nations (United Kingdom, EU) require open access publishing and when women have higher odds ratio for being first authors, this might be due to the high number of female from these countries.
+While past surveys among, for instance, UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
+But nothing in the extant literature would point out to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
 
@@
 In contrast, topics related to cognition and computation were more often associated with men, which reflects a wider trend in psychology [@gonzalez2020women].
+In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
 
-[@ashmos2011gendering; @sebo2020gender; @thelwall2019gender], In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+<!-- I did not know what these references signaled, so temporarily disabled them: -->
+<!-- [@ashmos2011gendering; @sebo2020gender; @thelwall2019gender], -->
 
@@
 The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.
-While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are more likely to hold first author positions in coauthored publications but less likely likely to hold last author positions than men.
+While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are significantly more likely to hold first author positions in coauthored publications but significantly less likely likely to hold last author positions than men.
 For solo publications and being a coauthor, there are no significant gender differences.
@@
 The aggregated results are in line with previous findings in psychology that favored men [@gonzalez2020women].
-We also observed that female first authored articles tend to be more likely to be published as open access as male first authors.
-The past surveys of experiences and attitudes towards open access publishing has revealed gender differences typically in the opposite direction [@zhu2017support].
+While the present analysis cannot determine the underlying causes of the observed trends, we can point to potential factors that may help explain why this discipline appears to be more equitably distributed by gender for the first authors than, for example, psychology.
+Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for Musicae Scientiae 1997-2009).
+However, this does not fully explain the rise of female first authors over the past 15 years, nor is it consistent with the reverse pattern observed in the last authorship position.
+A cautious interpretation is that the increase in first authorship may signal the discipline becoming more gender-balanced, but this shift has not yet extended to last authorship positions, which still reflect either career longevity or barriers to women attaining senior academic positions.
 
-While the present analysis cannot determine the underlying causes of the observed trends, we can point to several factors that may help explain why this discipline appears to be more equitably distributed by gender than, for example, psychology.
+While the analysis of gender of the authorships across the discipline-specific journal is relatively simple summary in the present study, there are more elaborate ways into obtaining insights into the gendered career progression in this disciple; for instance, future work could analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018].
+One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to pin down the factors contributing to the barriers experienced by women to take up senior author positions.
 
-(1) Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch and Edward Carterette for ICMPC in 1989, Irène Deliège for ESCOM in 1991), to a more recent ICMPC and ESCOM presidents?
-Editor-in-chiefs of the journals?).
-
-(2) Taking sub-disciplinary differences in psychology as an indicator of topic areas more often favoured by women, @gonzalez2020women observed that “nurturing” disciplines such as education and developmental psychology had higher proportions of female authors.
-In music psychology, however, a similar explanation may not apply.
-Keywords typically associated with nurturing domains (e.g., education, development, social, therapy, well-being, mental health, emotion regulation) account for only 17.9% of the total keyword counts among the top 40 keywords.
-
-- It would have been interesting to analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018]
-
-- Current findings present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
-While we have comparable accounts from psychology, musicology and other disciplines of music research (popular music, ethnomusicology) tends to have less equal division gender; a recent analysis of the last 25 years of articles in 15 musicology journals amounting to 3,515 articles suggests that female authors are far more infrequent (34.8% [33.3%-36.3%]) than males (65.2% [63.7%-66.7%]) [@eerola2026inprep].
-
-- Senior/last authorship position and why such a large inbalance exists.
+Current findings present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
+While the results are comparable to psychology in general, musicology (and the closely associated disciplines such as music theory, popular music and ethnomusicology) tend to have less equal division of gender in overall authorship; a recent analysis of the last 25 years of articles in 15 most eminent musicology journals amounting to 3,515 articles suggests that female authors are far more infrequent (34.8% [33.3%-36.3%]) than males (65.2% [63.7%-66.7%]) [@eerola2026inprep].
+The present findings should be used in conjuction of ...
 

### 2025-09-19 manuscript/manuscript.qmd
Worked on introduction: improved the flow of the text (though some terminology to be checked

@@
 In recent years, meta-scientific attention has increasingly turned toward music psychology, with a focus on how WEIRD [@jakubowski2024] and transparent [@eerola2023transparency] the discipline is.
-In this study, we examine gender equality in authorship as an indicator of how equitably gender is represented within the field.
+In this study, we examine in particular an indicator of equity within the field: gender equality in authorship.
+Authorship visibility is critical for academic career progression, influencing recognition, funding, and advancement opportunities.
 Disparities in gender distribution may signal structural barriers and disciplinary cultures that constrain the diversity of contributors and hinder optimal knowledge production [@ni2021gendered].
-Authorship visibility is also critical for academic career progression, influencing recognition, funding, and advancement opportunities.
 Establishing a clear understanding of gender representation in music psychology is therefore essential for supporting a more inclusive and sustainable academic community.
@@
 Research across several scientific disciplines show that women tend to be less represented as authors compared to men [@banks2025women; @son2022scientific].
-Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving.
-From around 1960 to 2021, studies show a general improvement [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender].
-Yet, there is evidence that this improvement is plateauing [@jemielniak2025gender] and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
-What counts as an authorship may differ between academic disciplines.
-Women are more likely to experience authorship disagreements, for example, when to decide authorship: women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
-Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
-Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
-
-Past studies have explored the gender distribution across a range of elements that could illuminate the reasons for the discrepancies.
-One of these is geographical regions.
-For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
-In physics journals, female representation was lowest for Latin America and the Caribbean (16%), but highest for Africa (19%) [@son2022scientific].
-In medical journals, women authors were least common in publications originating from South Korea, Japan, and Austria, but most common in South Africa and India [@bruck2023bibliometric].
-Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship: Although female contribution to scientific output is lower in countries ranked in lowest quartile for human development and gender equity, this also occured in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
-
-Another aspect that differs between distribution of gender in authorship is the research methodologies and topic.
-In an analyses of article keywords of medical journals, male-authored articles were more associated with keywords like phase II-III trials, oncology, immunotherapy, while female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care [@bruck2023bibliometric].
-In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
-These trends could be explained by the idea that female researchers often value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
-In terms of methodologies, women are are more likely to publish qualitative than quantitative/experimental studies [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
-Interestingly, such stereotypes are occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
-Overall, women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
-
 As the topics, approaches, and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
@@
 In psychology, which is perhaps the closest benchmark for the present focus on music psychology, female authors accounted for 45.2% of the sample (N = 90,067).
+Encouragingly, however, studies looking at various year ranges between 1960 until 2024 show a general trend that the ratio of women to men in authorships is improving [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+Yet, there is evidence that this improvement is plateauing [@jemielniak2025gender].
+Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
+There could be several reasons for such a disparity: two key aspects that are typically explored in such <article analysis> papers: geographical locations and keyword/topics.
 
-To improve academic equity issues, it is worthwhile focusing on our discipline and to assess the equity in music psychology.
-The past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
+Past studies have explored the gender distribution could by due to geographical regions.
+For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
+However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship, though again perhaps not as fully expected.
+Although female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
+Overall, while socioeconomic status of certain countries might be a reason for unbalanced female authorship [@ioannidis2023gender], human development might not fully explain such differences [@bruck2023bibliometric; @sanchez2024analysis; @son2022scientific], and further research is required to better understand the trends.
+
+Another aspect that differs between distribution of gender in authorship is the research methodologies and topics.
+In terms of methodologies, women are are less likely to publish quantitative and experimental studies, but rather publish qualitative work [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as they demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
+Interestingly, such stereotypes are occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
+In an analyses of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
+Similarly, in computational linguistics, woman are more likely to publish on speech, social, and conversational topics than males; the latter are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+These trends could be explained by the idea that female researchers often value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+Overall, research shows that women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
+
+Looking more broadly past the information within papers themselves, other research provides some further explanatins for discrepancies in female/male authorships.
+Despite efforts to give fair credit for author contributions [@ni2021gendered] using, for example, the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
+Women are more likely to experience authorship disagreements, for example, when to decide authorship: women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
+Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
+Additionally, women faculty perform more university, campus, and departmental services than men in general [@guarino2017faculty].
+Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
+To improve academic equity issues, it is worthwhile focusing on our discipline and to assess within subdisciplines to provide more 'localised' support to improve the equity in music psychology.
+Indeed, the past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
 @eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
@@
 
+->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
+
 Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.

### 2025-09-19 manuscript/manuscript.qmd
Worked on discussion: mirrored a few aspects from the introduction and added a few more sentences to improve flow.

@@
 However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
-Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship, though again perhaps not as fully expected.
-Although female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship.
+women’s academic success is correlated  positively with national indicators of gender equity [@chan2020].
+However, this does not seem to be as clearly reflected in authorships.
+Female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
 Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
@@
 In terms of methodologies, women are are less likely to publish quantitative and experimental studies, but rather publish qualitative work [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as they demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
-Interestingly, such stereotypes are occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
+Interestingly, such stereotypes occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
 In an analyses of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
 Similarly, in computational linguistics, woman are more likely to publish on speech, social, and conversational topics than males; the latter are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
-These trends could be explained by the idea that female researchers often value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+These trends could be explained by the idea that female researchers often conduct research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
 Overall, research shows that women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
 
-Looking more broadly past the information within papers themselves, other research provides some further explanatins for discrepancies in female/male authorships.
+Looking more broadly past the information within papers themselves, other research provides some further explanations for discrepancies in female/male authorships.
 Despite efforts to give fair credit for author contributions [@ni2021gendered] using, for example, the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
@@
 There is no simple rationale immediately apparent (e.g., country size, membership in a particular political bloc, gender equality index, or domination of articles by a single eminent scholar from one country) that would explain these differences, but this representation provides some starting points for discussion and may highlight the impact of academic career structures and principles implemented in these countries.
-Panel B shows data aggregated by continent, where Africa has meagre representation (52 authors), Oceania consists only of Australia and New Zealand, and the Americas are represented almost entirely by the US and Canada.
+Panel B shows data aggregated by continent, though we note that the representation is markedly different between continents.
+Africa has meagre representation of only 52 authors, Oceania consists only of Australia and New Zealand, and the Americas are represented almost entirely by the US and Canada.
 While this level of presentation may be too coarse for detailed analysis, it does offer a broad summary of the current situation.
+Overall, Oceanie and Americas have higher female author representation, while Asia and Europe have more a male representation.
 
@@
 Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
-However, there are several inequities within this aspect of academia, with women often being underrepresented.
-As such representations differ across different disciplines, it is worthwhile to diagnose the situation in music psychology.
+However, there are several inequities within this aspect of academia, with women often being underrepresented in aspects such as authorships.
 The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
 While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
+As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in music psychology.
 
@@
 Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
-By contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+This corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
 Given that last authorship in the social sciences is not universally defined but is increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution], this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
@@
 Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
+This suggests that the more 'leadership' author position is more likely to go to the male author.
 
@@
 While past surveys among, for instance, UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
-But nothing in the extant literature would point out to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
+The idea that publishing in a more open and transparent way benefit society in general [@pervsic2023open], and this fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+However, nothing in the extant literature would point out to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
 
@@
 Research areas related to children, well-being, health, and mental health were more frequently associated with women.
-This pattern is consistent with observations in psychology more broadly, where women are more strongly represented in sub-disciplines that focus on care and nurturing [@gonzalez2020women].
+This pattern is consistent with observations in psychology more broadly, where women are more strongly represented in sub-disciplines that focus on (health)care and nurturing [@bruck2023bibliometric; @gonzalez2020women].
 The higher likelihood of women studying topics related to singing may be connected to prevailing gender stereotypes, or to the early negative perceptions of singing among boys [@Warzecha2013] and men [@palkki2015gender].
@@
 In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+It perhaps might be interesting to note that 'emotion' has an equal balance between men and women.
+Although this may come as a surprise hen coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (toiviainen papers?), as well as being used in a more 'emotional' sense.
+this is the same for the keyword 'expression'.
+Although on first glance, it might fit the stereotype that this keyword might be more related to females, the 'expressivity' is again related to a more analytical coding aspect <cite Wanderley/thompson etc.>
+
+In summary, music psychology has similar groupings of topic for female and male authors as in other disciplines of psychology and medicines.
+Nonetheless, it should a positive trend.
+Below we also discuss how this information can help benefit the general field towards more equity with academia.
 
@@
 While the results are comparable to psychology in general, musicology (and the closely associated disciplines such as music theory, popular music and ethnomusicology) tend to have less equal division of gender in overall authorship; a recent analysis of the last 25 years of articles in 15 most eminent musicology journals amounting to 3,515 articles suggests that female authors are far more infrequent (34.8% [33.3%-36.3%]) than males (65.2% [63.7%-66.7%]) [@eerola2026inprep].
-The present findings should be used in conjuction of ...
+The present findings should be used in conjunction of ...
 

### 2025-09-19 manuscript/_figure1.qmd
Added a table and rewrote conclusions, polished keywords, and highlighted issues that need attention

@@
 
+#annual2$Type[1:4]<-NA
+annual3<-dplyr::filter(annual2,YearRange!="Overall")
+#annual3<-annual2
+#annual3$Type[1:4]<-NA
+
 fig1 <- ggplot(annual2,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator))+
@@
   geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
-#  geom_line(alpha = 0.2,position = position_dodge(width = 0.4))+
+  geom_line(data=annual3,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator),alpha = 0.75,position = position_dodge(width = 0.4))+
   scale_shape_manual(name="Authorship\ntype",values = c(22, 16),guide = "none") +

### 2025-09-19 manuscript/manuscript-apaquarto.qmd
Added a table and rewrote conclusions, polished keywords, and highlighted issues that need attention

--- 2025-08-05 manuscript/manuscript.qmd
@@
   - name          : "Tuomas Eerola"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
+    corresponding: yes
     address       : "Palace Green, Department of Music, Durham University, DH1 3RL Durham, United Kingdom"
     email         : "tuomas.eerola@durham.ac.uk"
+    url: "https://tuomaseerola.github.io/"
+    roles:
+      - conceptualization
+      - methodology
+      - formal Analysis
+      - writing
+    affiliations:
+      - id: id1
+        name: "Durham University"
+        department: Department of Music
+        address: Palace Green
+        city: Durham
+        country: United Kingdom
+        postal-code: DH1 3DA
   - name          : "Anna Czepiel"
-    affiliation   : "2"
-    address       : "Department of Psychology, University of Toronto, Mississauga, Canada"
-    email         : "a.czepiel@utoronto.ca"
-affiliation:
-  - id            : "1"
-    institution   : "Department of Music, Durham University"
-  - id            : "2"
-    institution   : "Department of Psychology, University of Toronto"
-
-author_note: |
-  Tuomas Eerola is at the Department of Music, Durham University, UK.
-  Anna Czepiel is at the Department of Psychology, University of Toronto, Canada.
-
-abstract: |
- blah blah blah.  
-keywords          : "gender, transparency, music psychology, open science, meta-research"
-wordcount         : "5555"
-figsintext        : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : no
-mask              : no
-csl: apa7.csl
-class             : "man"
+    roles:
+      - data curation
+      - formal analysis
+      - writing
+    affiliations   :
+      - id: id2
+        name       : "Department of Psychology, University of Toronto Mississauga"
+        city: Toronto
+        country: Canada
+author_note:
+      data-sharing: "https://tuomaseerola.github.io/gender_in_music_psych"
+abstract: "Academia suffers from various issues regarding equity.
+One aspect is gender inequality in academic authorship, where many studies show that women are often underrepresented.
+Due to authorship differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
+In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
+Overall, female authors were in the minority, accounting for 40.2% of all authors.
+However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
+These patterns have become more pronounced over the past 15 years.
+No substantial gender differences in citation counts were observed.
+Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
+Finally, an examination of keywords suggested broad thematic preferences differing by gender.
+Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture that hopefully inspires discussions and initiatives to further promote fairness and equity in academia."
+keywords          : "gender, transparency, music psychology, equality, meta-research"
+impact-statement: ~
+floatsintext: true
+# Numbered lines (.pdf and .docx only)
+numbered-lines: false
+# File with references
 bibliography: references.bib
+# Suppress title page
+suppress-title-page: false
+# Link citations to references
+link-citations: true
+# Masks references that appear in the masked-citations list
+mask: false
+# If true, adds today's date below author affiliations. If text, can be any value.
+# This is not standard APA format, but it is convenient.
+# Works with docx, html, and typst. 
+draft-date: false
+# Language options. See https://quarto.org/docs/authoring/language.html
+lang: en
+language:
+  citation-last-author-separator: "and"
+  citation-masked-author: "Masked Citation"
+  citation-masked-date: "n.d."
+  citation-masked-title: "Masked Title"
+  email: "Email"
+  title-block-author-note: "Author Note"
+  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
+  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
+  title-impact-statement: "Impact Statement"
+  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
 format:
-  html: default
+#  apaquarto-docx: default
+#  apaquarto-html: default
+  apaquarto-typst: 
+    keep-typ: true
+    list-of-figures: false
+    list-of-tables: false
 ---
@@
 
-Equality – Why Should We Care?
+Academia suffers various equity issues, impacting productivity, innovation, and job satisfaction in the workplace, which-in turn-hugely impact academic progress.
+Although there are several aspects, the focus of this paper is gender inequity.
+Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
+For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
+A similar situation holds for Italy (24% female professors) [@Filandri03082021], and to some extent for the United States (14%) [@spoon2023].
+A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the United Kingdom) [@quadlin2023higher].
+In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
+The number and quality (e.g., journal reputation) of authorships an individual determines crucial career development such as grant funding and future (permanent) positions.
 
-- Lack of equity impacts:
+Research across several scientific disciplines show that women tend to be less represented as authors compared to men [@banks2025women; @son2022scientific].
+As the topics, approaches, and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
+For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%), but not equal.
+In psychology, which is perhaps the closest benchmark for the present focus on music psychology, female authors accounted for 45.2% of the sample (N = 90,067).
+Encouragingly, however, studies looking at various year ranges between 1960 until 2024 show a general trend that the ratio of women to men in authorships is improving [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+Yet, there is evidence that this improvement is plateauing [@jemielniak2025gender].
+Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
+There could be several reasons for such a disparity: two key aspects that are typically explored in such <article analysis> papers: geographical locations and keyword/topics.
 
-- productivity
-- innovation
-- job satisfaction
+Past studies have explored the gender distribution could by due to geographical regions.
+For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
+However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship.
+women’s academic success is correlated  positively with national indicators of gender equity [@chan2020].
+However, this does not seem to be as clearly reflected in authorships.
+Female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
+Overall, while socioeconomic status of certain countries might be a reason for unbalanced female authorship [@ioannidis2023gender], human development might not fully explain such differences [@bruck2023bibliometric; @sanchez2024analysis; @son2022scientific], and further research is required to better understand the trends.
 
-Academia has acknowledged serious challenges in gender equity
+Another aspect that differs between distribution of gender in authorship is the research methodologies and topics.
+In terms of methodologies, women are are less likely to publish quantitative and experimental studies, but rather publish qualitative work [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as they demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
+Interestingly, such stereotypes occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
+In an analyses of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
+Similarly, in computational linguistics, woman are more likely to publish on speech, social, and conversational topics than males; the latter are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+These trends could be explained by the idea that female researchers often conduct research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+Overall, research shows that women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
 
-- women less present in positions of power (e.g. 30% UK profs)
-- pay gap is wider in academia (16%) than in other sectors (9%)
-- women represent fewer than 30% authorships globally
-
-Weave these studies into the narrative:
-
-- @ni2021gendered presented an overview of the gendered nature of authorship
-
-- @son2022scientific Scientific authorship by gender and they concluded that...
-
-- @shah2021gender showed how far off is the gender parity in scientific authorship using a bibliometric analysis.
-
-- @rock2021quantifying quantified the gender gap in authorship in herpetology, showing that .... (we can omit this)
-
-See @gonzalez2020women for a baseline from other disciplines:
-
-- Hard Sciences (N=119,592) = 0.148
-- Bio and Social Sciences (N=262,122) =  0.433
-- Psychology (N=90,067) =  0.452
-
-What is the situation in music psychology?
+Looking more broadly past the information within papers themselves, other research provides some further explanations for discrepancies in female/male authorships.
+Despite efforts to give fair credit for author contributions [@ni2021gendered] using, for example, the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
+Women are more likely to experience authorship disagreements, for example, when to decide authorship: women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
+Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
+Additionally, women faculty perform more university, campus, and departmental services than men in general [@guarino2017faculty].
+Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
+To improve academic equity issues, it is worthwhile focusing on our discipline and to assess within subdisciplines to provide more 'localised' support to improve the equity in music psychology.
+Indeed, the past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
+@eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
+@jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is.
+In a similar way to these papers, the current paper aims to give a current overview of gender patterns in authorship in the field of music psychology.
 
@@
 Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
-What is the proportion of different types of authorships (lead, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
-Are there specific trends in terms of countries of the affiliations, topics, or time?
+More specifically, we want to identify what is the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
+Are there specific patterns in the authorship distributions across countries, topics of the studies, or temporal trends?
 
@@
 ## Materials and analyses
-
-We retrieved bibliographic information for all articles published between 2000 and June 2025 from five specialist journals, resulting in 3,373 unique articles: Musicae Scientiae (N = 639), Psychology of Music (N = 1,231), Music Perception (N = 675), Journal of New Music Research (N = 563), and Music & Science (N = 265).
-These journals have also been used in previous meta-science studies to characterise research practices in music psychology [@jakubowski2024].
-Author affiliations were extracted automatically and converted into country-level data.
-However, these were not manually verified for each entry, as affiliations are not always clearly matched to individual authors due to variations in reporting conventions, such as multiple or partial affiliations.
-
-In total, the dataset included 9,066 authors, of whom 5,573 were unique.
-Author affiliations spanned 63 countries.
-We also extracted citation counts and Open Access status from Scopus.
-Information on joint first authorship was not available in the data.
-
-Gender attribution was initially based on first names.
-We recognise that treating gender as a binary category is inherently problematic.
-Gender is a complex and multidimensional social construct.
-However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
-We used the genderizeR API [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
-This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
-Only 89 names had a low attribution probability (< 0.55).
-Unattributed cases were then checked manually, resulting in 185 manual corrections.
-After this process, 32 names remained ambiguous and 27 were unknown—some likely due to data entry errors in Scopus (e.g.
-only initials or surnames).
-These 59 cases were excluded from the dataset.
-It is likely that the due to challenge of attributing gender correctly not all the gender attributions are correct.
-However, this error has been previously been shown to be non-biased, i.e.
-showing similar quantify of mistakes in both directions [ref].
-
-We defined three types of authorship positions: first authorship, coauthorship, and last authorship.
-Coauthorship includes all positions other than first and last.
-These positions carry different academic prestige: first authorship is typically associated with the primary contributor, while last authorship is often held by senior researchers with established reputations [@tscharntke2007author].
-
-For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
-Our analysis focused on the female author proportion (FAP) and odds ratios comparing the likelihood of occupying each authorship position by gender.
-
-- Manual corrections (to be done!).
-
-# Results
 
@@
 
-Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),2)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
-To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of occupying different authorship positions.
-We first investigate the authorship types across gender and then examine the author order and number in more detail.
+We retrieved bibliographic information for all articles published between 2000 and June 2025 from five specialist journals, resulting in 3,373 unique articles: Musicae Scientiae (N = 639), Psychology of Music (N = 1,231), Music Perception (N = 675), Journal of New Music Research (N = 563), and Music & Science (N = 265).
+These journals have also been used in previous meta-science studies to characterise research practices in music psychology [@jakubowski2024].
+Author affiliations were extracted automatically and converted into country-level data.
+However, these were not manually verified for each entry, as affiliations are not always clearly matched to individual authors due to variations in reporting conventions, such as multiple or partial affiliations.
+
+In total, the dataset included `{r} nrow(df)` authors, of whom `{r} length(unique(df$unique_name))` were unique.
+Author affiliations spanned `{r} length(unique(df$Affiliation_country_code))` countries.
+We also extracted citation counts and Open Access status from Scopus.
+Information on joint first authorship was not available in the data.
+
+Gender attribution was initially based on first names.
+We recognise that treating gender as a binary category is inherently problematic.
+Gender is a complex and multidimensional social construct.
+However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and that non-binary authorship generally is less than 0.1% [@son2022scientific], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
+We used the _genderize API_ [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
+This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
+Only `89` names had a low attribution probability (< 0.55).
+Unattributed cases were then checked manually, resulting in `185` manual corrections.
+After this process, `32` names remained ambiguous and `27` were unknown—some likely due to data entry errors in Scopus (e.g.
+only initials or surnames).
+These `59` cases were excluded from the dataset.
+It is likely that the due to challenge of attributing gender accurately not all the gender attributions are correct.
+Although genderize.io has been shown to achieve 96.6% accuracy in a diverse multinational test database without using a country of origin information -- 98% with the information included -- it also underperforms (82% accuracy) in Asian names [@vanhelene2024inferring].
+However, the error rates of these processes have been previously been shown to be non-biased, i.e.
+showing similar number of mistakes for both genders [@sebo2021performance;@vanhelene2024inferring;@thelwall2019gender].
+
+We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full first name coded with initials, or the second name used as the first name.
+This led to `53` corrections.
+
+We defined four types of authorship positions: single authorship, first authorship (which does not include papers with a single author), coauthorship, and last authorship.
+Coauthorship includes all positions other than first and last.
+These positions carry different academic prestige: first authorship is typically associated with the primary contributor, while last authorship is often held by senior researchers with established reputations [@tscharntke2007author].
+
+For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
+Our analysis focused on the female author proportion (FAP) and utilised odds ratios (ORs) to compare the likelihood of occupying each authorship position by gender.
+
+# Results
+
+Among all authors (N = `{r} nrow(df)`), `{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`% (N = `{r} nrow(dplyr::filter(df,Gender=='female'))`) were identified as female.
+To account for the unequal number of male and female authors across the dataset, we used odds ratios (ORs) to compare the relative likelihood of females occupying different authorship positions.
+We first investigate the authorship types across gender and then examine the author order and the overall number of coauthors in more detail.
 
@@
 
-The overall odds ratio for women as first authors was 1.32 [95% CI: 1.21–1.44], indicating that the odds of first authorship are 32% higher for females compared to the male authors.
-For coauthorship positions, the OR was 0.99 [0.90–1.08], suggesting no substantial gender difference.
-In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower in females.
-For the first and last authorship position, the odds ratios are significant at _p_ < .001 level.
+A summary of the authorship types and positions are given in Table 1.
+While the single-authored papers are in minority (8.6%) in this sample overall, there does not seem to be a significant difference between men and women authoring such articles (_p_ = .21); 38.05% of single-authored papers are women with an odds ratio of 0.90 [95% 0.78-1.05] suggesting no clear gender difference despite the figures leaning towards higher frequency of male authors.
+The overall OR for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
+For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
+In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
+
+```{r}
+#| id: table1
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| output: asis
+d <- read.csv('../data/authorship_counts_table.csv',header=TRUE)
+colnames(d)[1]<-"Author Type"
+colnames(d)[4]<-"Female %"
+colnames(d)[6]<-"OR (CI lower)"
+colnames(d)[7]<-"OR (CI higher)"
+print(knitr::kable(d,digits=2,caption="Summary of authorship types across gender."))
+```
 
 To illustrate the trends across time in the authorship positions across the gender, we looked at 5-year-windows of the authorship types across gender, shown in Figure 1.
+The overall numbers across the whole data are illustrated in the first column of the graph and the remaining five columns portray the chronology in 5-year windows.
 The higher odds for female authors at first author position seems to be established in 2010 onwards.
@@
 df <- mutate(group_by(df, BIBTEXKEY), author_N = max(author_order))
+
+s2 <- summarise(group_by(df, BIBTEXKEY,Gender), n = n(),.groups="drop")
+
+S <- s2 %>%
+  dplyr::group_by(Gender) %>%
+  dplyr::summarise(N=n(),m=mean(n,na.rm = TRUE),sd=sd(n,na.rm = TRUE),.groups = 'drop') %>%
+  dplyr::mutate(se=sd/sqrt(N),LCI=m+qnorm(0.025)*se,UCI=m+qnorm(0.975)*se)
+
+T <- t.test(s2$n~s2$Gender)
+
 ```
-The range of the number of authors in varies considerable.
-Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`,SD=`{r} round(sd(s$n),2)`).
+The variation of the number of authors across the papers is considerable.
+Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`, SD=`{r} round(sd(s$n),2)`).
+Previous research has calculated the collaboration index across gender [@gonzalez2020women], suggesting that male author tend to have larger number of coauthors (M = 1.5) than female (M = 1.34) in psychology.
+Here the average author numbers for male and female authors (disregarding authorship positions) indicate similar difference, higher mean author numbers for males (M = `{r} round(S$m[2],2)`, [`{r} round(S$LCI[2],2)`-`{r} round(S$UCI[2],2)`]) than for females (M = `{r} round(S$m[1],2)`, [`{r} round(S$LCI[1],2)`-`{r} round(S$UCI[1],2)`]), _t_(4640.58) = 7.21, _p_ < .001, Cohen's _d_ = -0.21, 95% CI
+[-0.26, -0.15]. This analysis, however, does not allow us to distinguish the authorship roles in the counts.
+
 To illustrate how the actual number of authors in a paper impacts female first author likelihood, we calculated the odds ratios across the number of authors in paper, pooling the papers with 7 or more authors together (N=`{r} sum(df$author_N>6)`,  `{r} round(sum(df$author_N>6)/nrow(df)*100,1)`%).
@@
 Here we tested whether the citations -- as indexed by Scopus -- show differences across gender.
-The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_=`{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
+The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_ = `{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
 The same comparison of citations for studies with last authors yields similar results `{r} citestats_last$Md[1]` [`{r} citestats_last$CI_lower[1]`–`{r} citestats_last$CI_upper[1]`], and for males, the numbers are similar (Md=`{r} citestats_last$Md[2]` [`{r} citestats_last$CI_lower[2]`–`{r} citestats_last$CI_upper[2]`]).
-Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_=`{r} round(stats_last$p.value,3)`).
-If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_=`{r} round(stats_all$p.value,4)`).
+Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_ = `{r} round(stats_last$p.value,3)`).
+If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_ < `{r} round(stats_all$p.value,4)`).
+For reference, prior study of citation patterns across gender in psychology has reported a small effect favoring men (M = 16.57) over women (M = 15.90) [@gonzalez2020women].
 
@@
 #| fig-width: 9
-#| fig-height: 6
+#| fig-height: 8
 #| fig.cap: > 
-#|   Female author odds ratio across more prolific countries publishing music psychology.
+#|   Female author odds ratio across the prolific countries publishing music psychology (with more than sixty publications).
 
@@
 
-OPTIONAL
+Panel A in Figure 3 shows a small number of countries, such as Australia, China, and Canada, with a high odds ratio for female authorship.
+At the other end of the spectrum, Japan, Belgium, Austria, Sweden, and Turkey all have significantly lower odds ratios for female authors.
+There is no simple rationale immediately apparent (e.g., country size, membership in a particular political bloc, gender equality index, or domination of articles by a single eminent scholar from one country) that would explain these differences, but this representation provides some starting points for discussion and may highlight the impact of academic career structures and principles implemented in these countries.
+Panel B shows data aggregated by continent, though we note that the representation is markedly different between continents.
+Africa has meagre representation of only 52 authors, Oceania consists only of Australia and New Zealand, and the Americas are represented almost entirely by the US and Canada.
+While this level of presentation may be too coarse for detailed analysis, it does offer a broad summary of the current situation.
+Overall, Oceanie and Americas have higher female author representation, while Asia and Europe have more a male representation.
 
-Table: Female author odds ratios across continents.
+## Gendered topics through analysis of keywords
 
-|Continent |    N| Odds_ratio| CI_low| CI_high|
-|:---------|----:|----------:|------:|-------:|
-|Europe    | 4985|       0.82|   0.75|    0.89|
-|Americas  | 2356|       1.12|   1.02|    1.23|
-|Asia      |  931|       0.89|   0.78|    1.02|
-|Oceania   |  742|       1.66|   1.43|    1.93|
-|Africa    |   52|       0.91|   0.52|    1.59|
+To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (41,438 in total, 5,960 unique).
+To reduce variant keywords, we manually simplified the variants (e.g., "arts in health", "healthy music use", "health musicking" were converted to "music health" and "music cognition" to "cognition" and "music perception" to "perception") and eliminated single use of the term "music".
+We aggregated the counts of keywords across all authors in the articles and Figure 4 shows the odds ratios for female authors to be associated (in any author position) with the 40 most frequently used keywords.
+The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
+At the opposite end of this continuum, topics such as music analysis, harmony, machine learning, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+There is much nuance and detail in the data that would warrant separate analyses; however, the purpose here is simply to provide reference points for comparing gender distributions of authorship with other disciplines, where thematic analyses have already been conducted (e.g., in sub-disciplines of psychology).
 
-## More analyses?
+```{r}
+#| id: keywords
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 9
+#| fig-height: 8
+#| fig.cap: > 
+#|   Female author odds ratio across 40 most frequent keywords in all articles as aggregated across all authors.
+library(stringr)
+source(here("scripts","count2category.R"))
+source(here("scripts","keyword_analysis.R"))
 
-- Association with the keywords (similar)
+```
+
+# Discussion
+
+Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
+However, there are several inequities within this aspect of academia, with women often being underrepresented in aspects such as authorships.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
+While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
+As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in music psychology.
+
+In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
+This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
+In terms of authorship roles, notable differences emerge.
+Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
+This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+This corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+Given that last authorship in the social sciences is not universally defined but is increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution], this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
+This pattern is consistent with broader evidence showing that women remain underrepresented in senior academic positions, including professorial appointments [@bakker2016tenure], invitations to contribute to journals, and authorship in high-prestige outlets [@holman2018gender].
+Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
+This suggests that the more 'leadership' author position is more likely to go to the male author.
+
+There were no substantial gender differences in the citation patterns.
+However, one notable observation is that first-authored articles by women were significantly more likely to be published open access, with an odds ratio of 1.49, a result that warrants further exploration.
+This effect does not appear to be tied to specific journals.
+Although the odds ratio for women publishing open access is highest in _Music and Science_, the broader trend suggests a temporal shift: since around 2010, women have increasingly overtaken men in publishing open access.
+The data does not provide any direct reasons for this, although the legislation in few of the European nations (United Kingdom, EU) require open access publishing and when women have higher odds ratio for being first authors, this might be due to the high number of female from these countries.
+While past surveys among, for instance, UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
+The idea that publishing in a more open and transparent way benefit society in general [@pervsic2023open], and this fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+However, nothing in the extant literature would point out to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
+
+->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
+[TE: is this supporting the idea that women are publishing more access than men? Or is this more about the lack of senior positions, which are probably not going to be achieved with internal service?]
+
+Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
+Our summaries of gender patterns across continents and countries indicate clear cross-national variation.
+Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
+This pattern, however, is not universal across all countries examined [@chan2020].
+Prior work shows that women’s research success is positively correlated with national indicators of gender equity [@chan2020].
+This is likely to be related to a career length, which strongly correlates with overall productivity.
+Women exhibit higher dropout rates than men, particularly during the 5–10 years following their first publication [@huang2020], but policies geared towards equity and support for career breaks and caring responsibilities may well contribute towards the observed country-specific differences.
+Nevertheless, prior analyses demonstrate that even when women constitute a smaller share of researchers within a discipline or country, their contributions are often more impactful than those of their male colleagues [@chan2020].
+When examining countries where women have significantly lower odds of authorship compared to men (Japan, Belgium, Austria, Sweden, and Turkey), it is not straightforward to conclude that the lower proportion of female author is related gender equality rankings.
+The notable exception is Turkey, which is positioned lower according to the Gender Inequality Index [@UNDPHDR2024].
+
+The keyword analysis of the articles indicated thematic differences in the topics studied by women and men.
+Research areas related to children, well-being, health, and mental health were more frequently associated with women.
+This pattern is consistent with observations in psychology more broadly, where women are more strongly represented in sub-disciplines that focus on (health)care and nurturing [@bruck2023bibliometric; @gonzalez2020women].
+The higher likelihood of women studying topics related to singing may be connected to prevailing gender stereotypes, or to the early negative perceptions of singing among boys [@Warzecha2013] and men [@palkki2015gender].
+In contrast, topics related to cognition and computation were more often associated with men, which reflects a wider trend in psychology [@gonzalez2020women].
+In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
+It perhaps might be interesting to note that 'emotion' has an equal balance between men and women.
+Although this may come as a surprise hen coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (top 3 authors for the keyword "emotion" are majority male, Saarikallio, Juslin, and Schubert, but "emotion regulation" top three is fully female), as well as being used in a more 'emotional' sense.
+This is the same for the keyword "expression".
+Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression"  is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).
+
+In summary, music psychology has similar groupings of topic for female and male authors as in other disciplines of psychology and medicines.
+Below we also discuss how this information can help benefit the general field towards more equity with academia.
+
+<!-- I did not know what these references signaled, so temporarily disabled them: -->
+
+<!-- @ashmos2011gendering : women publish more qualitative studies than men -->
+
+<!-- @sebo2020gender : women were more likely to publish qualitative studies and to publish in primary healthcare journals -->
+
+<!-- @thelwall2019gender : women are likely to use exploratory and qualitative methods rather than quantitative methods. -->
 
@@
 
-- Encouraging overall
+The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.
+While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are significantly more likely to hold first author positions in coauthored publications but significantly less likely likely to hold last author positions than men.
+For solo publications and being a coauthor, there are no significant gender differences.
 
-- Compare with sub-disciplines of psychology [@gonzalez2020women] Psychology: First authors (N=33,631)
+There was small but significant effect of gender on citations that favoured men when the citations were aggregated all authors in the papers.
+There was no, however, differences in citations for female first or last authors.
+The aggregated results are in line with previous findings in psychology that favored men [@gonzalez2020women].
+While the present analysis cannot determine the underlying causes of the observed trends, we can point to potential factors that may help explain why this discipline appears to be more equitably distributed by gender for the first authors than, for example, psychology.
+Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for _Musicae Scientiae_ 1997-2009).
+However, this does not fully explain the rise of female first authors over the past 15 years, nor is it consistent with the reverse pattern observed in the last authorship position.
+A cautious interpretation is that the increase in first authorship may signal the discipline becoming more gender-balanced, but this shift has not yet extended to last authorship positions, which still reflect either career longevity or barriers to women attaining senior academic positions.
 
-- Female role models (past and present) exist in music psychology (list names and their eminent roles, chairs of societies, editors of the main journals, etc.
-Diana Deutsch, Carol Krumhansl, Irene Deliege, to a more recent ICMPC and ESCOM presidents).
+While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same dataa could analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018].
+One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to identify the factors contributing to the barriers experienced by women to take up senior author positions.
 
-- Is this an area that appeals to women? (see comparison to sub-disciplines of psychology, where "softer" areas tend to have higher proportion)
-
-- Not matched by musicology: 25% female profs, 38% below??? (these are just some UK stats, I don't think the field-wide analysis has been done)
-
-- Senior authorship positions
+Current findings from authorship counts in music psychology present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia in general are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
+While the results are broadly comparable to psychology as a whole, musicology and its closely related disciplines such as music theory, popular music studies, and ethnomusicology tend to show less balanced gender representation in overall authorship.
+Gender, however, is only one dimension of equity; racial, ethnic, geographical, institutional, and other inequalities also merit attention in music psychology.
+Although women remain underrepresented in terms of absolute authorship counts in specialist music psychology journals, even a simple tally of contributions can help open discussion and provide evidence of progress toward greater inclusion and equality.
+To interpret these patterns more fully, one must examine how music psychology operates through its journals, conferences, training programmes, and curricula, as well as how the discipline presents itself both in countries with a long-standing tradition in the field and in regions where it has yet to become firmly established.
 

### 2025-09-19 manuscript/manuscript.qmd
Added a table and rewrote conclusions, polished keywords, and highlighted issues that need attention

@@
 keywords          : "gender, transparency, music psychology, open science, meta-research"
-wordcount         : "5555"
+wordcount         : "5578"
 figsintext        : yes
@@
 Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: two key aspects that are typically explored in such <article analysis> papers: geographical locations and keyword/topics.
+There could be several reasons for such a disparity: two key aspects that are typically explored in such **article analysis [TE: did you think citing specific studies here?]** papers: geographical locations and keyword/topics.
 
@@
 However, the error rates of these processes have been previously been shown to be non-biased, i.e.
-showing similar number of mistakes for both genders [@sebo2021performance;@vanhelene2024inferring].
+showing similar number of mistakes for both genders [@sebo2021performance;@vanhelene2024inferring;@thelwall2019gender].
 
@@
 
-The overall OR for women as first authors was 1.408 [95% CI: 1.284–1.544], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
-For coauthorship positions, the OR was 0.997 [0.913–1.090], suggesting no substantial gender difference (_p_ = .96).
-In contrast, the odds ratio for women as last authors was 0.733 [0.667–0.806], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
-Being a single author has odds ratio of 0.905 [0.778-1.052] for females, which suggests that while women are represented with relatively more single authored papers than men, the difference is not significant (_p_ = .21).
+A summary of the authorship types and positions are given in Table 1.
+While the single-authored papers are in minority (8.6%) in this sample overall, there does not seem to be a significant difference between men and women authoring such articles (_p_ = .21); 38.05% of single-authored papers are women with an odds ratio of 0.90 [95% 0.78-1.05] suggesting no clear gender difference despite the figures leaning towards higher frequency of male authors.
+The overall OR for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
+For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
+In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
+
+```{r}
+#| id: table1
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| output: asis
+d <- read.csv('../data/authorship_counts_table.csv',header=TRUE)
+colnames(d)[1]<-"Author Type"
+colnames(d)[4]<-"Female %"
+colnames(d)[6]<-"OR (CI lower)"
+colnames(d)[7]<-"OR (CI higher)"
+print(knitr::kable(d,digits=2,caption="Summary of authorship types across gender."))
+```
 
@@
 
-To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (15,316 in total, 6,730 unique).
+To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (41,438 in total, 5,960 unique).
 To reduce variant keywords, we manually simplified the variants (e.g., "arts in health", "healthy music use", "health musicking" were converted to "music health" and "music cognition" to "cognition" and "music perception" to "perception") and eliminated single use of the term "music".
@@
 In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology (34.8% [@eerola2026inprep]) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
+This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
 In terms of authorship roles, notable differences emerge.
@@
 
-->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
+**->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
+[TE: is this supporting the idea that women are publishing more access than men?
+Or is this more about the lack of senior positions, which are probably not going to be achieved with internal service?]**
 
@@
 Nevertheless, prior analyses demonstrate that even when women constitute a smaller share of researchers within a discipline or country, their contributions are often more impactful than those of their male colleagues [@chan2020].
-When examining countries where women have significantly lower odds of authorship compared to men (Japan, Belgium, Austria, Sweden, and Turkey), it is not straightforward to conclude that these nations fall consistently at the lower end of gender equality rankings.
+When examining countries where women have significantly lower odds of authorship compared to men (Japan, Belgium, Austria, Sweden, and Turkey), it is not straightforward to conclude that the lower proportion of female author is related gender equality rankings.
 The notable exception is Turkey, which is positioned lower according to the Gender Inequality Index [@UNDPHDR2024].
@@
 It perhaps might be interesting to note that 'emotion' has an equal balance between men and women.
-Although this may come as a surprise hen coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (toiviainen papers?), as well as being used in a more 'emotional' sense.
-this is the same for the keyword 'expression'.
-Although on first glance, it might fit the stereotype that this keyword might be more related to females, the 'expressivity' is again related to a more analytical coding aspect <cite Wanderley/thompson etc.>
+**Although this may come as a surprise coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (top 3 authors for the keyword "emotion" are majority male, Saarikallio, Juslin, and Schubert, but "emotion regulation" top three is fully female), as well as being used in a more 'emotional' sense.
+This is the same for the keyword "expression".
+Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression"  is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).**
 
 In summary, music psychology has similar groupings of topic for female and male authors as in other disciplines of psychology and medicines.
-Nonetheless, it should a positive trend.
 Below we also discuss how this information can help benefit the general field towards more equity with academia.
@@
 <!-- I did not know what these references signaled, so temporarily disabled them: -->
-<!-- [@ashmos2011gendering; @sebo2020gender; @thelwall2019gender], -->
+
+<!-- @ashmos2011gendering : women publish more qualitative studies than men -->
+
+<!-- @sebo2020gender : women were more likely to publish qualitative studies and to publish in primary healthcare journals -->
+
+<!-- @thelwall2019gender : women are likely to use exploratory and qualitative methods rather than quantitative methods. -->
 
@@
 While the present analysis cannot determine the underlying causes of the observed trends, we can point to potential factors that may help explain why this discipline appears to be more equitably distributed by gender for the first authors than, for example, psychology.
-Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for Musicae Scientiae 1997-2009).
+Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for _Musicae Scientiae_ 1997-2009).
 However, this does not fully explain the rise of female first authors over the past 15 years, nor is it consistent with the reverse pattern observed in the last authorship position.
@@
 
-While the analysis of gender of the authorships across the discipline-specific journal is relatively simple summary in the present study, there are more elaborate ways into obtaining insights into the gendered career progression in this disciple; for instance, future work could analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018].
-One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to pin down the factors contributing to the barriers experienced by women to take up senior author positions.
+While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same dataa could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
+One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to identify the factors contributing to the barriers experienced by women to take up senior author positions.
 
-Current findings present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
-While the results are comparable to psychology in general, musicology (and the closely associated disciplines such as music theory, popular music and ethnomusicology) tend to have less equal division of gender in overall authorship; a recent analysis of the last 25 years of articles in 15 most eminent musicology journals amounting to 3,515 articles suggests that female authors are far more infrequent (34.8% [33.3%-36.3%]) than males (65.2% [63.7%-66.7%]) [@eerola2026inprep].
-The present findings should be used in conjunction of ...
+Current findings from authorship counts in music psychology present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia in general are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
+While the results are broadly comparable to psychology as a whole, musicology and its closely related disciplines such as music theory, popular music studies, and ethnomusicology tend to show less balanced gender representation in overall authorship.
+Gender, however, is only one dimension of equity; racial, ethnic, geographical, institutional, and other inequalities also merit attention in music psychology.
+Although women remain underrepresented in terms of absolute authorship counts in specialist music psychology journals, even a simple tally of contributions can help open discussion and provide evidence of progress toward greater inclusion and equality.
+To interpret these patterns more fully, one must examine how music psychology operates through its journals, conferences, training programmes, and curricula, as well as how the discipline presents itself both in countries with a long-standing tradition in the field and in regions where it has yet to become firmly established.
 

### 2025-09-19 scripts/keyword_analysis.R

@@ -91,6 +91,45 @@ if(duplication==TRUE){
 kw_num = data %>% count(KW)
 
 
+# # # # # # instead of year, keep first author id
+alt <- TRUE
+
+if(alt==TRUE){
+
+  KW <- NULL
+  d_index <- NULL
+  GENDER_index <- NULL
+  year_index <- NULL # year
+
+for (k in 1:nrow(df)) {
+  tmp <- df$AUTHOR_KEYWORDS[k]
+  GENDER <- as.character(df$Gender[k])
+  y <- df$unique_name[k]
+  tmp <- stringi::stri_trans_general(tmp, "latin-ascii")
+  if (!is.na(tmp)) {
+    kw <- str_split(tmp, '[;,]', simplify = TRUE)
+    KW <- c(KW, kw)
+    year_index <- c(year_index, rep(y, length(kw)))
+    GENDER_index <- c(GENDER_index, rep(GENDER, length(kw)))
+  }
+}
+
+KW <- stringi::stri_trans_general(KW, "latin-ascii")
+KW<-stringr::str_trim(KW, side = "both")
+KW<-tolower(KW)
+length(KW)
+length(unique(KW))
+
+data <- data.frame(KW,
+                   GENDER_index,
+                   year_index)
+}
+
+# # # # # #
+
+
+
+
 # length(unique(data$KW))
 
 ## Words related to music cognition
@@ -1132,7 +1171,7 @@ data = data %>% mutate(
 
 
 dim(data)
-length(unique(data$KW)) # before: 6734
+length(unique(data$KW))
 
 
 
@@ -1177,16 +1216,75 @@ data = data %>% mutate(
                            "music emotions")
                  ~ "emotion",TRUE ~ KW))
 
+data = data %>% mutate(
+  KW = case_when(KW %in% c("emotions")
+                 ~ "emotion",TRUE ~ KW))
+
+# Music preference
+idx = grep("musical preference", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("musical preference",
+                           "musical preferences")
+                 ~ "music preference",TRUE ~ KW))
+# Music training
+idx = grep("musical training", data$KW)
+unique(data$KW[idx])
+data = data %>% mutate(
+  KW = case_when(KW %in% c("musical training",
+                           "school-based musical training","early musical training","age of musical training onset")
+                 ~ "music training",TRUE ~ KW))
+
[…]

### 2025-09-19 scripts/quantify_authorship.R

@@ -4,6 +4,10 @@
 # Gender analysis of MusicPsych authors
 # Use Odds ratios to quantify the probabilities of gendered authorships.
 
+# Store the raw counts in a separate table
+stored_table <- data.frame(matrix(data = NA,nrow=5,ncol=3)); colnames(stored_table)<-c("Author Type","Female","Male")
+
+
 #### Classify order: Single author ----------
 df$authortype <- 'Other'
 df$authortype[df$author_order == 1] <- 'Other'
@@ -18,6 +22,8 @@ t <- table(
 )
 t
 or1 <- effectsize::oddsratio(t) # 0.83       | [0.71, 0.96]
+stored_table[1,2:3] <- t[1,]; stored_table[1,1]<-"Single"
+#stored_table
 #print(or1)
 #fisher.test(t)
 
@@ -36,6 +42,7 @@ t <- table(
 t
 
 or2 <- effectsize::oddsratio(t) # 1.27       | [1.03, 1.56]
+stored_table[2,2:3] <- t[1,]; stored_table[2,1]<-"First"
 #print(or2)
 #fisher.test(t)
 
@@ -54,6 +61,7 @@ t <- table(
 t
 
 or3 <- effectsize::oddsratio(t) # 1.06       | [0.77, 1.46]
+stored_table[3,2:3] <- t[1,]; stored_table[3,1]<-"Co-author"
 #print(or3)
 #fisher.test(t)
 
@@ -72,6 +80,9 @@ t <- table(
 t
 
 or4 <- effectsize::oddsratio(t) # 0.73
+stored_table[4,2:3] <- t[1,]; stored_table[4,1]<-"Last"
+
+
 #print(or4)
 
 or <- rbind(or1,or2,or3,or4)
@@ -79,6 +90,22 @@ or$name<-c("Single","First","Coauthor","Last")
 or<-dplyr::select(or,name,Odds_ratio,CI_low,CI_high)
 print(knitr::kable(or,digits = 3,caption = "Female authorship odds ratios by authorship type."))
 
+
+t<-table(df$Gender)
+t
+stored_table[5,2:3] <- as.numeric(t); stored_table[5,1]<-"All"
+stored_table$Female_prc <- stored_table$Female/(stored_table$Female+stored_table$Male)*100
+
+stored_table$OR<-NA
+stored_table$OR_L<-NA
+stored_table$OR_U<-NA
+stored_table$OR[1:4]<-or$Odds_ratio
+stored_table$OR_L[1:4]<-or$CI_low
[…]

### 2025-09-25 manuscript/manuscript.qmd
Worked on discussion: first and second paragraph (added literature/comparisons) and added limitations paragraph

@@
 Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: two key aspects that are typically explored in such **article analysis [TE: did you think citing specific studies here?]** papers: geographical locations and keyword/topics.
+There could be several reasons for such a disparity.
+Two key aspects typically explored in bibliometric studies are geographical locations [e.g., @ioannidis2023gender] and keyword/topics [e.g., @bruck2023bibliometric].
 
@@
 Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
-However, there are several inequities within this aspect of academia, with women often being underrepresented in aspects such as authorships.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
-While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
-As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in music psychology.
+However, there are several inequities within this aspect of academia, with women often being underrepresented in authorships.
+As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in the field of music psychology.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in music psychology journals, specifically Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science.
+Overall in music psychology, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
+This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for a third of authors (33.2% [@huang2020]; 36.2% [@gonzalez2020women]).
+This representation is also much higher than in other STEM subjects of math, physics and computer (15% [@huang2020]) and physics topics like astronomy and bioscience (17.8%, @son2022scientific).
+However, overall women are still underrepresented.
+In exploring the nuances behind this, we further analysed differences between authorship role, number of citations, geography, and also keyword analysis.
 
-In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
-In terms of authorship roles, notable differences emerge.
+In terms of authorship roles, notable differences emerge between authorship positions.
 Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
@@
 Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
+This patterns is also consistent with the idea of a "leaky pipeline", where women are more likely to drop away from academic careers with increasing seniority, especially between docotral and (consolidated) researcher positions [@gonzalez2020women].
+
+Additionally, women are less likely to self-promote accomplishments [@rudman1998self], and therefore might not push for the more senior role 
  This suggests that the more 'leadership' author position is more likely to go to the male author.
@@
 Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression"  is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).**
+
+There are certain limitations of the current research.
+First, the gender of the authors are not part of the database - we infer the gender based on the name using XYZ.
+Despite this, because the field is relatively small, a check from both authors meant that a certain number of the more ambiguous names (in terms of their gender) could be fixed (see accompanying code).
+Additionally, the error (confidence) of the gender attribution was so small, that it is unlikely that the remaining names that were not necessarily attributed correctly would significantly change the main conclusions.
+Another limitation is that while we recognise that gender is a complex and multidimensional social construct, there are limitations to treating gender as a binary category.
+We used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and generally the number of non-binary authors is currently relatively low (0.8% @son2022scientific).
+Nonetheless, future research should also move towards using more inclusive, non-binary information.
 

### 2025-09-26 manuscript/manuscript.qmd
small tweaks in intro, much work on discussion.

--- 2025-09-19 manuscript/manuscript.qmd
@@
 abstract: "Academia suffers from various issues regarding equity.
-One aspect is gender inequality in academic authorship, where many studies show that women are often underrepresented.
+One such issue is gender inequality in authorship, where many studies show that women are often underrepresented.
 Due to authorship differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
-A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor).
+A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, and coauthor).
 In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
@@
 No substantial gender differences in citation counts were observed.
-Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
+Analyses of geographical variation revealed clear country and continental differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
 Finally, an examination of keywords suggested broad thematic preferences differing by gender.
-Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture that hopefully inspires discussions and initiatives to further promote fairness and equity in academia."
+Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture to inspire discussions and initiatives to further promote fairness and equity in academia."
 
@@
 Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: two key aspects that are typically explored in such **article analysis [TE: did you think citing specific studies here?]** papers: geographical locations and keyword/topics.
 
-Past studies have explored the gender distribution could by due to geographical regions.
-For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
+Past studies have explored the gender distribution according to geographical regions and in relationship to country demographics.
+In one study, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
 However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
-Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship.
-women’s academic success is correlated  positively with national indicators of gender equity [@chan2020].
-However, this does not seem to be as clearly reflected in authorships.
-Female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in academia.
+However, while national indicators of gender equity is positively correlated with women’s academic success [@chan2020], this does not seem to be as clearly reflected in authorships.
+Female scientific output is not only lower in countries ranked in lowest quartile for human development and gender equity, but also in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
 Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
@@
 Interestingly, such stereotypes occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
-In an analyses of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
+There are also key different when assessing article keywords.
+In one analysis of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
 Similarly, in computational linguistics, woman are more likely to publish on speech, social, and conversational topics than males; the latter are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
@@
 Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
-To improve academic equity issues, it is worthwhile focusing on our discipline and to assess within subdisciplines to provide more 'localised' support to improve the equity in music psychology.
+To improve academic equity issues, it is worthwhile focusing on subdiscipline to provide more 'localised' support to improve the equity in field.
+This was the aim of the current paper to look within our own field of music psychology.
 Indeed, the past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
@@
 
-Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
-However, there are several inequities within this aspect of academia, with women often being underrepresented in aspects such as authorships.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
-While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
-As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in music psychology.
+Authorship is a crucial marker of scholarly success in academia, which strongly influences academic career progression, recognition, funding, and advancement opportunities.
+However, there are several inequities within academic authorship, with women often being underrepresented.
+As male-female authorship representations differ across different academic disciplines, it is worthwhile diagnosing the situation in the field of music psychology.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science.
+Overall, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
+This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for a third of authors (33.2% [@huang2020]; 36.2% [@gonzalez2020women]).
+This representation is also much higher than in other STEM subjects of math, physics and computer (15% [@huang2020]) and physics topics like astronomy and bioscience (17.8%, @son2022scientific).
+While our current findings suggests that the field of music psychology is doing better in fostering gender equity than certain other fields, women are still overall underrepresented.
+In exploring the nuances behind this, we further analysed differences between authorship role, number of citations, geography, and article keywords.
 
-In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
 In terms of authorship roles, notable differences emerge.
-Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
+Women are more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
 This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
 In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
-This corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
-Given that last authorship in the social sciences is not universally defined but is increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution], this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
+This underrepresentation in woman as last authors corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+Although last authorship in the social sciences is not necessarily universally defined, it is nonetheless increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution].
+Therefore, this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
 This pattern is consistent with broader evidence showing that women remain underrepresented in senior academic positions, including professorial appointments [@bakker2016tenure], invitations to contribute to journals, and authorship in high-prestige outlets [@holman2018gender].
 Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
-This suggests that the more 'leadership' author position is more likely to go to the male author.
+This patterns is also consistent with the idea of a "leaky pipeline", where women are more likely to drop away from academic careers with increasing seniority, especially between doctoral and (consolidated) researcher positions [@gonzalez2020women] and generally have shorter academic careers [@sanchez2024analysis].
+On a broader level, this might highlight the need for more mentorship programs for (junior) women to senior positions.
+Indeed, mentorship programs have been shown support remaining in academia [@gardiner2007show].
+We note that within the broader music psychology community, there are some mentorship programs in place, for example, Women in Music Information Retrieval WIMIR (https://wimir.wordpress.com/) and the Society for Music Perception and Cognition has resources for trainees and a Mentorship committe (https://www.musicperception.org/committees) and often holds mentor-mentee meetings at conferences.
+Although these measures have been in places for some years, it might still take some time for the effect to translate and additional mentorship programs could be welcome.
 
@@
 
-**->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
-[TE: is this supporting the idea that women are publishing more access than men?
-Or is this more about the lack of senior positions, which are probably not going to be achieved with internal service?]**
-
 Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
-Our summaries of gender patterns across continents and countries indicate clear cross-national variation.
+Our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
 Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
 This pattern, however, is not universal across all countries examined [@chan2020].
-Prior work shows that women’s research success is positively correlated with national indicators of gender equity [@chan2020].
-This is likely to be related to a career length, which strongly correlates with overall productivity.
+Prior work shows that women’s research success can be positively correlated with national indicators of gender equity [@chan2020] or are related to countries with intermediate Human Development Index [@sanchez2024analysis; @sugimoto2015relationship].
+This discrepancy between author representation could likely to be related to a career length, which strongly correlates with overall productivity.
 Women exhibit higher dropout rates than men, particularly during the 5–10 years following their first publication [@huang2020], but policies geared towards equity and support for career breaks and caring responsibilities may well contribute towards the observed country-specific differences.
@@
 This is the same for the keyword "expression".
-Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression"  is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).**
+Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression" is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).**  The difference that female prefer 'caring' topics over male authors fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+This fits with another potential explanation for the disparity between authorship, namely that women might also lean toward and execute more university/campus/departmental services than men [@guarino2017faculty].
+Over time, performing such services to universities are time away from other work on acquiring funding and leading research, which would aggregate over time to time away from publishing research.
+It is nonetheless important for such services are completed; in some cases it could be there is more encouragement for males to take on these kinds of roles, but also it could be that success in academia could also be more societal defined as not only publications [@kiai2019protect], but as a whole should celebrate achievements that support internal, local and community services in a broader sense.
 
-In summary, music psychology has similar groupings of topic for female and male authors as in other disciplines of psychology and medicines.
-Below we also discuss how this information can help benefit the general field towards more equity with academia.
+There are certain limitations of the current research.
+The gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
+Nonetheless, we maintain that the conclusions may still hold.
+First, both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
+Second, there error (confidence) of the gender attribution was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
+Another limitation is treating gender as a binary category.
+While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low (0.8% @son2022scientific).
+Nonetheless, future research should also move towards using more inclusive, non-binary information.
+However, the main limitation is that such research indicates the symptoms of certain issues in academia, but not necessarily the causes.
+While we speculate on certain reasons for the current state of affairs, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promite better equity in academia.
 

### 2025-09-26 manuscript/manuscript-apaquarto.qmd
Updated text, mainly in-text citations and compiled pdf and word versions.

@@
 No substantial gender differences in citation counts were observed.
-Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
+Analyses of geographical variation revealed clear country and continental differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions.
 Finally, an examination of keywords suggested broad thematic preferences differing by gender.
-Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture that hopefully inspires discussions and initiatives to further promote fairness and equity in academia."
+Although this work overall shows that the field of music psychology is more balanced than some other (STEM) fields, it provides a picture to inspire discussions and initiatives to further promote fairness and equity in academia."
 keywords          : "gender, transparency, music psychology, equality, meta-research"
@@
 format:
-#  apaquarto-docx: default
+  apaquarto-docx: default
 #  apaquarto-html: default
-  apaquarto-typst: 
-    keep-typ: true
-    list-of-figures: false
-    list-of-tables: false
+#_  apaquarto-typst: 
+#_    keep-typ: true
+#_    list-of-figures: false
+#_    list-of-tables: false
 ---
@@
 Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
-There could be several reasons for such a disparity: two key aspects that are typically explored in such <article analysis> papers: geographical locations and keyword/topics.
 
-Past studies have explored the gender distribution could by due to geographical regions.
-For example, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
+Past studies have explored the gender distribution according to geographical regions and in relationship to country demographics.
+In one study, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
 However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
-Human development indicators and gender inequality indicators have also revealed systematic inequalities in authorship.
-women’s academic success is correlated  positively with national indicators of gender equity [@chan2020].
-However, this does not seem to be as clearly reflected in authorships.
-Female scientific output is lower in countries ranked in lowest quartile for human development and gender equity, female scientific output is also lower in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Human development indicators and gender inequality indicators have also revealed systematic inequalities in academia.
+However, while national indicators of gender equity is positively correlated with women’s academic success [@chan2020], this does not seem to be as clearly reflected in authorships.
+Female scientific output is not only lower in countries ranked in lowest quartile for human development and gender equity, but also in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
 Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
@@
 #| output: asis
+#| label: tbl-basics
+#| tbl-cap: Summary of authorship types across gender
 d <- read.csv('../data/authorship_counts_table.csv',header=TRUE)
@@
 colnames(d)[7]<-"OR (CI higher)"
-print(knitr::kable(d,digits=2,caption="Summary of authorship types across gender."))
+print(knitr::kable(d,digits=2))
 ```
@@
 
-Authorship is a crucial aspect in academia as a marker of scholarly success and strongly influences academic career progression, influencing recognition, funding, and advancement opportunities.
-However, there are several inequities within this aspect of academia, with women often being underrepresented in aspects such as authorships.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender.
-While we recognise that gender is a complex and multidimensional social construct, and there are limitations to treating gender as a binary category, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006].
-As male-female authorship representations differ across different academic disciplines, it is worthwhile to diagnose the situation in music psychology.
+Authorship is a crucial marker of scholarly success in academia, which strongly influences academic career progression, recognition, funding, and advancement opportunities.
+However, there are several inequities within academic authorship, with women often being underrepresented.
+As male-female authorship representations differ across different academic disciplines, it is worthwhile diagnosing the situation in the field of music psychology.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
+Overall, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
+This proportion is higher than in musicology [34.8% in @eerola2026inprep] and higher than in psychology more broadly, where women account for a third of authors [33.2% in @huang2020; 36.2%in @gonzalez2020women].
+This representation is also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
+While our current findings suggests that the field of music psychology is doing better in fostering gender equity than certain other fields, women are still overall underrepresented.
+In exploring the nuances behind this, we further analysed differences between authorship role, number of citations, geography, and article keywords.
 
-In music psychology, the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for 33.2% of authors [@huang2020].
 In terms of authorship roles, notable differences emerge.
-Women are relatively more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
+Women are more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
 This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
 In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
-This corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
-Given that last authorship in the social sciences is not universally defined but is increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution], this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
+This underrepresentation in woman as last authors corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+Although last authorship in the social sciences is not necessarily universally defined, it is nonetheless increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution].
+Therefore, this underrepresentation suggests that women are less likely to occupy leadership roles such as lab heads or holders of substantial research resources.
 This pattern is consistent with broader evidence showing that women remain underrepresented in senior academic positions, including professorial appointments [@bakker2016tenure], invitations to contribute to journals, and authorship in high-prestige outlets [@holman2018gender].
 Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
-This suggests that the more 'leadership' author position is more likely to go to the male author.
+This patterns is also consistent with the idea of a "leaky pipeline", where women are more likely to drop away from academic careers with increasing seniority, especially between doctoral and (consolidated) researcher positions [@gonzalez2020women] and generally have shorter academic careers [@sanchez2024analysis].
+On a broader level, this might highlight the need for more mentorship programs for (junior) women to senior positions.
+Indeed, mentorship programs have been shown support remaining in academia [@gardiner2007show].
+We note that within the broader music psychology community, there are some mentorship programs in place, for example, Women in Music Information Retrieval WIMIR (https://wimir.wordpress.com/) and the Society for Music Perception and Cognition has resources for trainees and a Mentorship committe (https://www.musicperception.org/committees) and often holds mentor-mentee meetings at conferences.
+Although these measures have been in places for some years, it might still take some time for the effect to translate and additional mentorship programs could be welcome.
 
@@
 
-->  (potential reasons: women faculty perform more university/campus/departmental services @guarino2017faculty )
-[TE: is this supporting the idea that women are publishing more access than men? Or is this more about the lack of senior positions, which are probably not going to be achieved with internal service?]
-
 Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
-Our summaries of gender patterns across continents and countries indicate clear cross-national variation.
+Our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
 Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
 This pattern, however, is not universal across all countries examined [@chan2020].
-Prior work shows that women’s research success is positively correlated with national indicators of gender equity [@chan2020].
-This is likely to be related to a career length, which strongly correlates with overall productivity.
+Prior work shows that women’s research success can be positively correlated with national indicators of gender equity [@chan2020] or are related to countries with intermediate Human Development Index [@sanchez2024analysis; @sugimoto2015relationship].
+This discrepancy between author representation could likely to be related to a career length, which strongly correlates with overall productivity.
 Women exhibit higher dropout rates than men, particularly during the 5–10 years following their first publication [@huang2020], but policies geared towards equity and support for career breaks and caring responsibilities may well contribute towards the observed country-specific differences.
@@
 In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
-It perhaps might be interesting to note that 'emotion' has an equal balance between men and women.
-Although this may come as a surprise hen coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (top 3 authors for the keyword "emotion" are majority male, Saarikallio, Juslin, and Schubert, but "emotion regulation" top three is fully female), as well as being used in a more 'emotional' sense.
-This is the same for the keyword "expression".
-Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression"  is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).
+The difference that female prefer 'caring' topics over male authors fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+This fits with another potential explanation for the disparity between authorship, namely that women might also lean toward and execute more university/campus/departmental services than men [@guarino2017faculty].
+Over time, performing such services to universities are time away from other work on acquiring funding and leading research, which would aggregate over time to time away from publishing research.
+It is nonetheless important for such services are completed; in some cases it could be there is more encouragement for males to take on these kinds of roles, but also it could be that success in academia could also be more societal defined as not only publications [@kiai2019protect], but as a whole should celebrate achievements that support internal, local and community services in a broader sense.
 
-In summary, music psychology has similar groupings of topic for female and male authors as in other disciplines of psychology and medicines.
-Below we also discuss how this information can help benefit the general field towards more equity with academia.
+There are certain limitations of the current research.
+The gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
+Nonetheless, we maintain that the conclusions may still hold.
+First, both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
+Second, there error (confidence) of the gender attribution was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
+Another limitation is treating gender as a binary category.
+While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.8% in @son2022scientific].
+Nonetheless, future research should also move towards using more inclusive, non-binary information.
+However, the main limitation is that such research indicates the symptoms of certain issues in academia, but not necessarily the causes.
+While we speculate on certain reasons for the current state of affairs, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promite better equity in academia.
 
@@
 
-While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same dataa could analyse the precarity of career options as judged from the years of authorship attributions [@lundine2018].
+While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same dataa could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
 One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to identify the factors contributing to the barriers experienced by women to take up senior author positions.

### 2025-09-26 manuscript/manuscript.qmd
Updated text, mainly in-text citations and compiled pdf and word versions.

@@
 As male-female authorship representations differ across different academic disciplines, it is worthwhile diagnosing the situation in the field of music psychology.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science.
+The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
 Overall, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology (34.8% @eerola2026inprep) and higher than in psychology more broadly, where women account for a third of authors (33.2% [@huang2020]; 36.2% [@gonzalez2020women]).
-This representation is also much higher than in other STEM subjects of math, physics and computer (15% [@huang2020]) and physics topics like astronomy and bioscience (17.8%, @son2022scientific).
+This proportion is higher than in musicology [34.8% in @eerola2026inprep] and higher than in psychology more broadly, where women account for a third of authors [33.2% in @huang2020; 36.2%in @gonzalez2020women].
+This representation is also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
 While our current findings suggests that the field of music psychology is doing better in fostering gender equity than certain other fields, women are still overall underrepresented.
@@
 In a similar trend in computational linguistics, woman are more likely to publish on speech, social and conversational topics than males, who are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
-It perhaps might be interesting to note that 'emotion' has an equal balance between men and women.
-**Although this may come as a surprise coming from different disciplines, music emotion can be a relatively 'analytical' aspect in music psychology, where emotion from music is attempted to be decoded in a more analytical way (top 3 authors for the keyword "emotion" are majority male, Saarikallio, Juslin, and Schubert, but "emotion regulation" top three is fully female), as well as being used in a more 'emotional' sense.
-This is the same for the keyword "expression".
-Although on first glance, it might fit the stereotype that this keyword might be more related to females, the "performance expression" is again related to a more analytical coding aspect (Marc Leman, Laura Bishop, and Roger Chaffin as top 3 authors).**  The difference that female prefer 'caring' topics over male authors fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
+The difference that female prefer 'caring' topics over male authors fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
 This fits with another potential explanation for the disparity between authorship, namely that women might also lean toward and execute more university/campus/departmental services than men [@guarino2017faculty].
@@
 Another limitation is treating gender as a binary category.
-While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low (0.8% @son2022scientific).
+While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.8% in @son2022scientific].
 Nonetheless, future research should also move towards using more inclusive, non-binary information.

### 2025-09-26 scripts/keyword_analysis.R

@@ -92,7 +92,7 @@ kw_num = data %>% count(KW)
 
 
 # # # # # # instead of year, keep first author id
-alt <- TRUE
+alt <- FALSE
 
 if(alt==TRUE){
 

### 2025-09-27 manuscript/manuscript.qmd
Changed title. Small tweaks in discussion.

@@
 ---
-title             : "Gender Distribution of Authors in Music Psychology"
+title             : "The Gendered Nature of Authorship in Music Psychology"
 shorttitle        : "Gender in Music Psychology"
@@
 
-Looking more broadly past the information within papers themselves, other research provides some further explanations for discrepancies in female/male authorships.
-Despite efforts to give fair credit for author contributions [@ni2021gendered] using, for example, the Contributor Roles Taxonomy (CRediT) system, assigning authorship can be unclear.
-Women are more likely to experience authorship disagreements, for example, when to decide authorship: women prefer to discuss in earlier stages, while men choose authorship at the final stage [@ni2021gendered].
-Another reason for the gender gap could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
-Additionally, women faculty perform more university, campus, and departmental services than men in general [@guarino2017faculty].
-Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
 To improve academic equity issues, it is worthwhile focusing on subdiscipline to provide more 'localised' support to improve the equity in field.
@@
 # Discussion
-
-Authorship is a crucial marker of scholarly success in academia, which strongly influences academic career progression, recognition, funding, and advancement opportunities.
-However, there are several inequities within academic authorship, with women often being underrepresented.
-As male-female authorship representations differ across different academic disciplines, it is worthwhile diagnosing the situation in the field of music psychology.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
-Overall, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology [34.8% in @eerola2026inprep] and higher than in psychology more broadly, where women account for a third of authors [33.2% in @huang2020; 36.2%in @gonzalez2020women].
-This representation is also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
-While our current findings suggests that the field of music psychology is doing better in fostering gender equity than certain other fields, women are still overall underrepresented.
+A crucial marker of scholarly success in academia is authorship, which strongly influences career progression, recognition, funding, and advancement opportunities.
+However, there are several inequities within academic authorship, where women are often underrepresented.
+As male-female authorship representations differ across different academic disciplines, the goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in the field of music psychology.
+An analysis of papers from five core music psychology journals across 25 years revealed that, overall, women represented 40.2% of authors.
+This proportion is higher than in musicology [34.8% in @eerola2026inprep], in psychology more broadly [33.2% in @huang2020; 36.2% in @gonzalez2020women], and also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
+Although our current findings suggests that the field of music psychology is doing better in fostering gender equity than some other fields, the distribution of authors by gender remains nonetheless unequal.
 In exploring the nuances behind this, we further analysed differences between authorship role, number of citations, geography, and article keywords.
 
-In terms of authorship roles, notable differences emerge.
+In terms of authorship roles, notable authorship differences emerge.
 Women are more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
-This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
-In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
-This underrepresentation in woman as last authors corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+This is in line with previous research showing that female authorship representation is increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+In contrast, we found that women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+This underrepresentation in woman as last authors is similar to the situation in medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
 Although last authorship in the social sciences is not necessarily universally defined, it is nonetheless increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution].
@@
 Related analyses from science and medicine further show that women are more often associated with conducting experiments, whereas men are disproportionately represented in authorship roles linked to leadership and oversight [@macaluso2016science].
-This patterns is also consistent with the idea of a "leaky pipeline", where women are more likely to drop away from academic careers with increasing seniority, especially between doctoral and (consolidated) researcher positions [@gonzalez2020women] and generally have shorter academic careers [@sanchez2024analysis].
-On a broader level, this might highlight the need for more mentorship programs for (junior) women to senior positions.
-Indeed, mentorship programs have been shown support remaining in academia [@gardiner2007show].
-We note that within the broader music psychology community, there are some mentorship programs in place, for example, Women in Music Information Retrieval WIMIR (https://wimir.wordpress.com/) and the Society for Music Perception and Cognition has resources for trainees and a Mentorship committe (https://www.musicperception.org/committees) and often holds mentor-mentee meetings at conferences.
-Although these measures have been in places for some years, it might still take some time for the effect to translate and additional mentorship programs could be welcome.
+This pattern is also consistent with finding that women are more likely to drop away from academic careers with increasing seniority [@gonzalez2020women; @sanchez2024analysis].
+Indeed, this gender gap could come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
+Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
+In particular, one crucial aspect that has been highlighted to is the need for more mentorship programs for (junior) women to senior positions.
+Indeed, mentorship programs can encourage mentees to remain in academia [@gardiner2007show].
+We note that within the broader music psychology community, there are already mentorship programs in place: for example, Women in Music Information Retrieval WIMIR (https://wimir.wordpress.com/) and the Society for Music Perception and Cognition (SMPC) has resources for trainees and a Mentorship committee (https://www.musicperception.org/committees) and often holds mentor-mentee meetings at conferences.
+Although these measures have been implemented for some years, it might still take some time for the effect to be reflected in authorship.
 
@@
 
-Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
-Our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
+In terms of eographical and cultural factors, our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
 Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
@@
 The difference that female prefer 'caring' topics over male authors fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
-This fits with another potential explanation for the disparity between authorship, namely that women might also lean toward and execute more university/campus/departmental services than men [@guarino2017faculty].
+This fits with another potential explanation for the disparity between authorship, namely that women faculty lean word and perform more university, campus, and departmental services than men in general [@guarino2017faculty].
 Over time, performing such services to universities are time away from other work on acquiring funding and leading research, which would aggregate over time to time away from publishing research.
@@
 There are certain limitations of the current research.
-The gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
-Nonetheless, we maintain that the conclusions may still hold.
-First, both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
-Second, there error (confidence) of the gender attribution was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
-Another limitation is treating gender as a binary category.
+First, the gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
+Nonetheless, we maintain that the current conclusions may still hold as both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
+Additionally, gender attribution error was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
+A second limitation is treating gender as a binary category.
 While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.8% in @son2022scientific].
 Nonetheless, future research should also move towards using more inclusive, non-binary information.
-However, the main limitation is that such research indicates the symptoms of certain issues in academia, but not necessarily the causes.
-While we speculate on certain reasons for the current state of affairs, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promite better equity in academia.
-
-<!-- I did not know what these references signaled, so temporarily disabled them: -->
-
-<!-- @ashmos2011gendering : women publish more qualitative studies than men -->
-
-<!-- @sebo2020gender : women were more likely to publish qualitative studies and to publish in primary healthcare journals -->
-
-<!-- @thelwall2019gender : women are likely to use exploratory and qualitative methods rather than quantitative methods. -->
+The third - and main - limitation is that such bibliometric research indicates the symptoms of academia issues, but not necessarily the causes.
+While we speculate on certain reasons for the current state of affairs as discussed, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promote better equity in academia.
 
 # Conclusions
-
 The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.
 While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are significantly more likely to hold first author positions in coauthored publications but significantly less likely likely to hold last author positions than men.
-For solo publications and being a coauthor, there are no significant gender differences.
+For solo publications and being a coauthor, there were no significant gender differences.
 
@@
 
-While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same dataa could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
+While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same data could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
 One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to identify the factors contributing to the barriers experienced by women to take up senior author positions.
 
-Current findings from authorship counts in music psychology present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia in general are [@bhopal2021competing;@harris2025gender;@Filandri03082021].
+Current findings from authorship counts in music psychology present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia in general are [@bhopal2021competing; @harris2025gender; @Filandri03082021].
 While the results are broadly comparable to psychology as a whole, musicology and its closely related disciplines such as music theory, popular music studies, and ethnomusicology tend to show less balanced gender representation in overall authorship.

### 2025-09-27 manuscript/manuscript.qmd
Fixed affliation for AMC

@@
     email         : "tuomas.eerola@durham.ac.uk"
-  - name          : "Anna Czepiel"
+  - name          : "Anna M. Czepiel"
     affiliation   : "2"
-    address       : "Department of Psychology, University of Toronto Mississauga, Canada"
+    address       : "Department of Psychology, University of Toronto Mississauga, L5L 1C6 Mississauga, Canada"
     email         : "a.czepiel@utoronto.ca"

### 2025-09-27 manuscript/_figure1.qmd
Presubmission version, ready for colleagues to read.

@@
 #| fig.cap: > 
-#|   Authorship types across time.
+#|   Authorship types across time
 

### 2025-09-27 manuscript/_figure2.qmd
Presubmission version, ready for colleagues to read.

@@
 #| fig.cap: > 
-#|   A) Number of authors and B) female odds ratio for authorship types across the number of authors.
+#|   A) Number of authors and B) female odds ratio for authorship types across the number of authors
 

### 2025-09-27 manuscript/manuscript-apaquarto.qmd
Presubmission version, ready for colleagues to read.

@@
 format:
-  apaquarto-docx: default
+#_  apaquarto-docx: default
 #  apaquarto-html: default
-#_  apaquarto-typst: 
-#_    keep-typ: true
-#_    list-of-figures: false
-#_    list-of-tables: false
+  apaquarto-typst: 
+    keep-typ: true
+    list-of-figures: false
+    list-of-tables: false
 ---
@@
 For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
-A similar situation holds for Italy (24% female professors) [@Filandri03082021], and to some extent for the United States (14%) [@spoon2023].
+A similar situation holds for Italy [24% female professors, @Filandri03082021], and to some extent for the United States [14%, @spoon2023].
 A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the United Kingdom) [@quadlin2023higher].
@@
 
-Authorship is a crucial marker of scholarly success in academia, which strongly influences academic career progression, recognition, funding, and advancement opportunities.
-However, there are several inequities within academic authorship, with women often being underrepresented.
-As male-female authorship representations differ across different academic disciplines, it is worthwhile diagnosing the situation in the field of music psychology.
-The goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in core music psychology journals, specifically _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_.
-Overall, we found that the distribution of authors by gender remains unequal, with women representing 40.2% of authors.
-This proportion is higher than in musicology [34.8% in @eerola2026inprep] and higher than in psychology more broadly, where women account for a third of authors [33.2% in @huang2020; 36.2%in @gonzalez2020women].
-This representation is also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
-While our current findings suggests that the field of music psychology is doing better in fostering gender equity than certain other fields, women are still overall underrepresented.
+A crucial marker of scholarly success in academia is authorship, which strongly influences career progression, recognition, funding, and advancement opportunities.
+However, there are several inequities within academic authorship, where women are often underrepresented.
+As male-female authorship representations differ across different academic disciplines, the goal of the current paper was to provide a snapshot of authorship prevalence and roles across gender in the field of music psychology.
+An analysis of papers from five core music psychology journals across 25 years revealed that, overall, women represented 40.2% of authors.
+This proportion is higher than in musicology [34.8% in @eerola2026inprep], in psychology more broadly [33.2% in @huang2020; 36.2% in @gonzalez2020women], and also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
+Although our current findings suggests that the field of music psychology is doing better in fostering gender equity than some other fields, the distribution of authors by gender remains nonetheless unequal.
 In exploring the nuances behind this, we further analysed differences between authorship role, number of citations, geography, and article keywords.
 
-In terms of authorship roles, notable differences emerge.
+In terms of authorship roles, notable authorship differences emerge.
 Women are more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
-This supports previous research showing that there is a positive trend with female representation increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
-In contrast, women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
-This underrepresentation in woman as last authors corresponds to the situation in other academic fields such as medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+This is in line with previous research showing that female authorship representation is increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+In contrast, we found that women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
+This underrepresentation in woman as last authors is similar to the situation in medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
 Although last authorship in the social sciences is not necessarily universally defined, it is nonetheless increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution].
@@
 
-Turning to potential drivers of these gender differences, geographical and cultural factors appear to play a major role.
-Our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
-Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
+In terms of eographical and cultural factors, our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
+Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a academic climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
 This pattern, however, is not universal across all countries examined [@chan2020].
@@
 There are certain limitations of the current research.
-The gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
-Nonetheless, we maintain that the conclusions may still hold.
-First, both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
-Second, there error (confidence) of the gender attribution was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
-Another limitation is treating gender as a binary category.
+First, the gender of the authors are not part of the database - gender was inferred based on author name using _genderize API_ [@wais2006].
+Nonetheless, we maintain that the current conclusions may still hold as both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
+Additionally, gender attribution error was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
+A second limitation is treating gender as a binary category.
 While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.8% in @son2022scientific].
 Nonetheless, future research should also move towards using more inclusive, non-binary information.
-However, the main limitation is that such research indicates the symptoms of certain issues in academia, but not necessarily the causes.
-While we speculate on certain reasons for the current state of affairs, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promite better equity in academia.
-
-<!-- I did not know what these references signaled, so temporarily disabled them: -->
-
-<!-- @ashmos2011gendering : women publish more qualitative studies than men -->
-
-<!-- @sebo2020gender : women were more likely to publish qualitative studies and to publish in primary healthcare journals -->
-
-<!-- @thelwall2019gender : women are likely to use exploratory and qualitative methods rather than quantitative methods. -->
+The third - and main - limitation is that such bibliometric research indicates the symptoms of academia issues, but not necessarily the causes.
+While we speculate on certain reasons for the current state of affairs as discussed, other kinds of research such as qualitative and longitudinal research would further elucidate the key problems, and where areas for improvements could be made to promote better equity in academia.
 
@@
 While overall women hold smaller number of authorship overall (`{r} round(nrow(dplyr::filter(df,Gender=='female'))/nrow(df),3)*100`%), they are significantly more likely to hold first author positions in coauthored publications but significantly less likely likely to hold last author positions than men.
-For solo publications and being a coauthor, there are no significant gender differences.
+For solo publications and being a coauthor, there were no significant gender differences.
 
@@
 
+### Acknowledgements
+
+We are grateful to Isabelle Peretz for suggesting a deeper look into the author types, and to our colleagues for their valuable input on the manuscript.
+
 # References

### 2025-09-27 manuscript/manuscript.qmd
Presubmission version, ready for colleagues to read.

@@
 # Discussion
+
 A crucial marker of scholarly success in academia is authorship, which strongly influences career progression, recognition, funding, and advancement opportunities.
@@
 In terms of eographical and cultural factors, our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
-Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a cultural climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
+Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a academic climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
 This pattern, however, is not universal across all countries examined [@chan2020].
@@
 # Conclusions
+
 The analysis of gender distributions in the five most prominent music psychology journals over the last 25 years presented a rich and evolving history of gender distribution in various authorship roles.

### 2025-12-12 manuscript/_figure1_presentation.qmd
First submission

--- 2025-09-27 manuscript/_figure1.qmd
@@
   #facet_wrap(~TypeGroup,nrow = 2)+
-  geom_hline(yintercept=1, linetype="dashed", color = "black") +
+  geom_hline(yintercept=1, linetype="dashed", color = "grey75") +
   geom_point(position = position_dodge(width = 0.4),show.legend=TRUE,size=3)+
@@
   xlab('Year range')+
-  ylab('Female Authorship Odds Ratio (95%CI)')+
+  ylab('Female Author Odds Ratio (95%CI)')+
   theme_classic(base_size = 15)
@@
 
+# apply dark theme
+fig1 <- fig1 + dark_theme_bw(base_size = 16) + theme(legend.position="top")
+
+# make transparent
+
+fig1<-fig1 + theme(
+  panel.background = element_rect(fill='transparent'),
+  plot.background = element_rect(fill='transparent', color=NA),
+  panel.grid.major = element_blank(),
+  panel.grid.minor = element_blank(),
+  legend.background = element_rect(fill='transparent',color=NA),
+  legend.box.background = element_rect(fill='transparent',color=NA)
+)
+fig1
+
+ggsave(fig1,filename = "figure1_authorship_oddsratio_dark.png",width = 7.8,height = 5,dpi = 300)
+
 ```

### 2025-12-12 manuscript/_figure2_presentation.qmd
First submission

--- 2025-08-26 manuscript/_figure2.qmd
@@
 #| fig.cap: > 
-#|   Authorship types across author counts.
+#|   A) Number of authors and B) female odds ratio for authorship types across the number of authors
 
@@
 custom_palette<-c(custom_palette[4],custom_palette[1:3])
-fig2 <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
-  geom_hline(yintercept=1, linetype="dashed", color = "black") +
+
+fig2b <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
+  geom_hline(yintercept=1, linetype="dashed", color = "gray50") +
   geom_point(position = position_dodge(width = 0.4),shape=15,size=3)+
@@
   scale_color_manual(name="Authorship\ntype",values = custom_palette) +
-  xlab('Author N')+
+  xlab('Number of authors/article')+
   ylab('Female Authorship Odds Ratio')+
   theme_classic(base_size = 15)
-print(fig2)
+#print(fig2)
+library(ggdark)
+fig2b <- fig2b + dark_theme_bw(base_size = 16) + theme(legend.position="top")
+
+# make transparent
+fig2b<-fig2b + theme(
+  panel.background = element_rect(fill='transparent'),
+  plot.background = element_rect(fill='transparent', color=NA),
+  panel.grid.major = element_blank(),
+  panel.grid.minor = element_blank(),
+  legend.background = element_rect(fill='transparent',color=NA),
+  legend.box.background = element_rect(fill='transparent',color=NA)
+)
+
+s <- summarise(group_by(df, BIBTEXKEY), n = n())
+S<-s
+S$n[S$n > 9]<-10
+fig2a <- ggplot(S,aes(x=n))+
+  geom_histogram(binwidth=1,fill='gold1',color='black')+
+  theme_classic(base_size = 15)+
+  labs(x='Number of authors/article',y='Count of articles')+
+  scale_x_continuous(breaks=seq(1,10,1),expand = c(0.001,0.001),limits = c(0,11),
+                     labels = c(seq(1,9,1),"10+"))+
+  scale_y_continuous(expand = c(0.001,0.001))
+#print(g)
+
+# apply dark theme
+fig2a <- fig2a + dark_theme_bw(base_size = 16) + theme(legend.position="top")
+
+# make transparent
+fig2a<-fig2a + theme(
+  panel.background = element_rect(fill='transparent'),
+  plot.background = element_rect(fill='transparent', color=NA),
+  panel.grid.major = element_blank(),
+  panel.grid.minor = element_blank(),
+  legend.background = element_rect(fill='transparent',color=NA),
+  legend.box.background = element_rect(fill='transparent',color=NA)
+)
+fig2a
+
+G<-cowplot::plot_grid(fig2a,fig2b,nrow=1,labels=c('A','B'),label_colour = "white",rel_widths = c(0.6,1))
+print(G)
+ggsave("figure2_authorship_by_number_of_authors_dark.png",
+  G,
+  width = 8.1,
+  height = 4,
+  dpi = 300
+)
+
 ```

### 2025-12-12 manuscript/manuscript-apaquarto.qmd
First submission

@@
       data-sharing: "https://tuomaseerola.github.io/gender_in_music_psych"
-abstract: "Academia suffers from various issues regarding equity.
-One such issue is gender inequality in authorship, where many studies show that women are often underrepresented.
-Due to authorship differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
-A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, and coauthor).
+abstract: "Academia suffers from various equity issues.
+One such issue is gender inequality in authorship, where many studies show that women are less represented compared to men.
+Due to authorship differences across disciplines, the goal of this study is to provide a current overview of the gender distribution of authorship within music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analysed with respect to author gender and authorship role (first, last, solo, and coauthor).
 In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
-Overall, female authors were in the minority, accounting for 40.2% of all authors.
-However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
+Overall, women authors were in the minority, accounting for 40.2% of all authors.
+However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (odds ratio, OR = 1.41), equally likely to be solo authors (OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
 These patterns have become more pronounced over the past 15 years.
@@
 # Suppress title page
-suppress-title-page: false
+suppress-title-page: true
 # Link citations to references
@@
 
-In recent years, meta-scientific attention has increasingly turned toward music psychology, with a focus on how WEIRD [@jakubowski2024] and transparent [@eerola2023transparency] the discipline is.
-In this study, we examine in particular an indicator of equity within the field: gender equality in authorship.
-Authorship visibility is critical for academic career progression, influencing recognition, funding, and advancement opportunities.
+In recent years, meta-scientific research has increasingly been taken up by music psychology, examining how Western, Educated, Industrialized, Rich and Democratic  [WEIRD, @jakubowski2024] and transparent [@eerola2023transparency] the discipline is.
+In the current article, we examine a particular indicator of equity within the field: gender equality in authorship.
+Authorship visibility is critical for academic career progression, as it influences recognition, funding, and advancement opportunities.
 Disparities in gender distribution may signal structural barriers and disciplinary cultures that constrain the diversity of contributors and hinder optimal knowledge production [@ni2021gendered].
-Establishing a clear understanding of gender representation in music psychology is therefore essential for supporting a more inclusive and sustainable academic community.
 Previous studies across various disciplines—including psychology—have consistently revealed persistent gender inequalities in authorship, though the extent of these disparities varies by field [@son2022scientific; @shah2021gender; @rock2021quantifying; @gonzalez2020women].
+To support a more inclusive and sustainable academic community, it is essential to establish a better understanding of gender representation in more localised subfield; in this case, in the field of music psychology.
 
-Academia suffers various equity issues, impacting productivity, innovation, and job satisfaction in the workplace, which-in turn-hugely impact academic progress.
-Although there are several aspects, the focus of this paper is gender inequity.
-Women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
-For instance, in the UK in 2016-17, 24.6% of professors are women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
-A similar situation holds for Italy [24% female professors, @Filandri03082021], and to some extent for the United States [14%, @spoon2023].
-A pay gap between men and women in academia is also significant, (e.g., female-male wage ratio is 0.85 in 2020 in the United Kingdom) [@quadlin2023higher].
+Academia across all disciplines suffers various equity issues, impacting productivity, innovation, and job satisfaction in the workplace, which—in turn—hugely impact academic progress, leadership opportunities, and visibility throughout a career [@allen2021work].
+Although there are several issues, this paper focuses on one particular issue: gender inequity.
+Although we recognise that gender is a multidimensional social construct, for the purposes of the current paper, and because non-binary authorship generally is less than 0.1% [@son2022scientific], we remain consistent and comparable gender and authorship research [@gonzalez2020women;@ni2021gendered;@rock2021quantifying;@shah2021gender;@son2022scientific;@wais2006] and generally adopt a binary (woman-man) classification.
+In disciplines overall, women are more at a disadvantage than men, with evidence suggesting that women are less present in positions of power.
+For instance, in the UK in 2016-17, 24.6% of professors were women [@bhopal2021competing] when compared to the 44% of all grades in the UK academia [@harris2025gender].
+A similar situation holds for Italy [24% female professors, @Filandri03082021], and to greater extent for the United States [14%, @spoon2023].
+A substantial pay gap exists between men and women in academia; for example, a woman-man wage ratio was 0.85 in 2020 in the United Kingdom [@quadlin2023higher].
 In a “publish or perish” culture of science and academia [@kiai2019protect], one crucial aspect to focus on is authorship.
-The number and quality (e.g., journal reputation) of authorships an individual determines crucial career development such as grant funding and future (permanent) positions.
+The number and quality (e.g., journal reputation) of authorships an individual has garnered can determine crucial career development such as grant funding and future (permanent) positions.
 
@@
 As the topics, approaches, and methods attract differential interest from men and women, it is important to recognise that fields of scholarship and academic disciplines vary considerably in terms of gender distribution [@huang2020].
-For example, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion was substantially higher (43.3%), but not equal.
-In psychology, which is perhaps the closest benchmark for the present focus on music psychology, female authors accounted for 45.2% of the sample (N = 90,067).
-Encouragingly, however, studies looking at various year ranges between 1960 until 2024 show a general trend that the ratio of women to men in authorships is improving [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
-Yet, there is evidence that this improvement is plateauing [@jemielniak2025gender].
-Further, although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
+When investigating the gender composition of publications, a large survey of authors by @gonzalez2020women showed that the hard sciences (N = 119,592) had the lowest prevalence of women (14.8%), whereas in the biological and social sciences (N = 262,122) the proportion of woman authors was substantially higher (43.3%), but not equal.
+In psychology, which is perhaps the closest benchmark for the present focus on music psychology, woman authors accounted for 45.2% of the sample (N = 90,067).
+Encouragingly, studies looking at various year ranges between 1960 until 2024 show a general trend that the ratio of women to men in authorships is improving [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+However, there is evidence that this improvement is plateauing [@jemielniak2025gender].
+Further, there are differences in terms of authorship position.
+Authorship position is important as it typically reflects career positions: primary contribution is typically associated with the primary contributor, which often are graduate students or postdocs, while last authorship is often held by senior researchers with established reputations [@tscharntke2007author].
+Although women are more likely to be the first author, women are still generally unfairly underrepresented as the last author [@bruck2023bibliometric; @gonzalez2020women; @rock2021quantifying; @shah2021gender].
+
+HERE
 
 Past studies have explored the gender distribution according to geographical regions and in relationship to country demographics.
-In one study, men out-numbered women more so in high-income compared to low-income countries [@ioannidis2023gender].
-However, there are also some unexpected trends: while female representation is lowest in countries like Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], females were most highly represented in countries like India and Africa [@bruck2023bibliometric; @son2022scientific].
-Human development indicators and gender inequality indicators have also revealed systematic inequalities in academia.
-However, while national indicators of gender equity is positively correlated with women’s academic success [@chan2020], this does not seem to be as clearly reflected in authorships.
-Female scientific output is not only lower in countries ranked in lowest quartile for human development and gender equity, but also in countries very high in gender equality; better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
-Another recent paper shows that female authorship rates are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
-Overall, while socioeconomic status of certain countries might be a reason for unbalanced female authorship [@ioannidis2023gender], human development might not fully explain such differences [@bruck2023bibliometric; @sanchez2024analysis; @son2022scientific], and further research is required to better understand the trends.
+In one study, men out-numbered women more in high-income compared to low-income countries [@ioannidis2023gender].
+However, there are also some unexpected trends: while female representation is lowest in Latin America, Caribbean [in physics journals, @son2022scientific] and in South Korea, Japan, and Austria [in medical journals, @bruck2023bibliometric], women were most highly represented in India and Africa [@bruck2023bibliometric; @son2022scientific].
 
-Another aspect that differs between distribution of gender in authorship is the research methodologies and topics.
-In terms of methodologies, women are are less likely to publish quantitative and experimental studies, but rather publish qualitative work [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as they demonstrate better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
-Interestingly, such stereotypes occur in the opposite direction: when assessing fictive research abstracts-one with quantitative and one with a qualitative design- scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender].
-There are also key different when assessing article keywords.
-In one analysis of article keywords of medical journals, female-authored articles were more associated with keywords more related to healthcare-related themes like patient involvement, insurance, and quality-of-care compared to males [@bruck2023bibliometric].
-Similarly, in computational linguistics, woman are more likely to publish on speech, social, and conversational topics than males; the latter are more likely to publish in formal mathematical, syntax, and semantic approaches [@vogel2012he].
-These trends could be explained by the idea that female researchers often conduct research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
-Overall, research shows that women choose qualitative research on social and societal topics, while males author papers using more quantitative methods on theoretical and analytical topics.
+Interestingly, while national indicators of gender equity are positively correlated with women’s academic success [@chan2020], this does not seem to be as clearly reflected in authorships.
+Scientific output from women is not only lower in countries ranked in the lowest quartile for human development and gender equity, but better equality seemed to occur in high and medium-developed countries [@sugimoto2015relationship].
+Another recent paper shows that authorship rates between men and women are most balanced in countries with low to intermediate Human Development Index [@sanchez2024analysis].
+Overall, while the socioeconomic status of certain countries might be a reason for unbalanced female authorship [@ioannidis2023gender], human development might not fully explain such differences [@bruck2023bibliometric; @sanchez2024analysis; @son2022scientific], and further research is required to better understand the trends.
 
-To improve academic equity issues, it is worthwhile focusing on subdiscipline to provide more 'localised' support to improve the equity in field.
-This was the aim of the current paper to look within our own field of music psychology.
-Indeed, the past efforts such as @anglada2019visualizing have already provided a current overview the publications in the music psychology field with suggestions of future directions to improve the field.
-@eerola2023transparency showcases how Open Science practices (e.g., preregistrations, sharing research materials, data, and analysis scripts) are relatively limited, and encourage authors in the field to start implementing such practices.
-@jakubowski2024 explores participant samples and musical stimuli used across a wide range of experiment to give an idea of how limited or generalizable the field is.
-In a similar way to these papers, the current paper aims to give a current overview of gender patterns in authorship in the field of music psychology.
+Another aspect that differs between distribution of gender in authorship is the research methodologies.
+Women are less likely to publish quantitative and experimental studies than qualitative work [@ashmos2011gendering; @nunkoo2020three; @sebo2020gender; @thelwall2019gender; @zhang2023gender], perhaps as they demonstrate the better relational skill involved in qualitative studies, like prolonged relationships and emotional ties with research participants [@sebo2020gender].
+When assessing fictive research abstracts—one with quantitative and one with a qualitative design—scientific quality of the qualitative abstract was ranked more trustworthy and accurate when assigned to a female author [@johansson2002gender], showing further evidence that of female stereotypes.
+
+Research also shows there that topics may vary by gender as shown by article keyword usage.
+In one analysis of article keywords of medical journals, articles authored by women were predominantly associated with keywords on healthcare-related themes like patient involvement, insurance, and quality-of-care, while articles authored by men were associated with keywords associated with higher publication citations, such as ‘phase II-III trials’, ‘anti-body based therapy’ males [@bruck2023bibliometric].
+Similarly, in computational linguistics, women are more likely to publish on speech, social, and conversational topics than men; men are more likely to publish on formal mathematical, syntax, and semantic approaches [@vogel2012he].
+These trends could be explained by evidence that research by women often conduct research on societal progress, while research by men is aimed more at scientific progress [@zhang2021gender].
+Overall, research shows that women tend to produce qualitative research on social and societal topics, while men use more quantitative methods on theoretical and analytical topics.
 
@@
 
-Our aim is to find out what is the gender distribution in the specialist journals of music psychology.
-More specifically, we want to identify what is the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years?
-Are there specific patterns in the authorship distributions across countries, topics of the studies, or temporal trends?
+To improve academic equity issues, it is worthwhile focusing on subdiscipline to provide more ‘localised’ support to improve the equity in field.
+Correspondingly, the aim of the current paper was to look within our own field of music psychology.
+This builds on past efforts in the music psychology field that have provided a current overview as well as future directions and improvements for: a) publications [@anglada2019visualizing], b) participant samples and stimuli for better generalization of findings [@jakubowski2024], and c) current practices and implementation of Open Science practices [e.g., preregistrations, sharing research materials, data, and analysis scripts, @eerola2023transparency].
+Our aim was to determine gender distribution in the specialist journals of music psychology.
+More specifically, we want to identify the proportion of different types of authorships (single, first, coauthors, and last authorships) for men and women in the published papers in the last 25 years and to explore specific patterns in the authorship distributions across countries, topics of the studies, or temporal trends.
+By giving a current overview of gender patterns in authorship in the field of music psychology, we hope this stimulates further discussion of identifying and implementing methods to combat such inequities for future generations.
 
@@
 
-Gender attribution was initially based on first names.
-We recognise that treating gender as a binary category is inherently problematic.
+Gender attribution was initially based on given names.
+As mentioned in the introduction, we recognise that treating gender as a binary category is inherently problematic.
 Gender is a complex and multidimensional social construct.
-However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and that non-binary authorship generally is less than 0.1% [@son2022scientific], we adopt a binary classification—male and female—for analytical purposes and assume that first names allow for a reasonable, though imperfect, attribution of gender.
-We used the _genderize API_ [@wais2006], which predicts gender from first names and can be supplemented with country information derived from author affiliations to improve accuracy.
+However, consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and the fact that non-binary authorship generally is less than 0.1% [@son2022scientific], although this is not necessarily the true picture of the non-binary authorship situation, we nevertheless adopt a binary classification—male and female—for analytical purposes and assume that given names allow for a reasonable, though imperfect, attribution of gender.
+We used the _genderize API_ [@wais2006], which predicts gender from given names and can be supplemented with country information derived from author affiliations to improve accuracy.
 This method resolved the gender of 89.3% of authors with a probability greater than 0.90.
-Only `89` names had a low attribution probability (< 0.55).
-Unattributed cases were then checked manually, resulting in `185` manual corrections.
-After this process, `32` names remained ambiguous and `27` were unknown—some likely due to data entry errors in Scopus (e.g.
-only initials or surnames).
-These `59` cases were excluded from the dataset.
+Only 89 names had a low attribution probability (< 0.55).
+Unattributed cases were then checked manually, resulting in 185 manual corrections.
+After this process, 32 names remained ambiguous and 27 were unknown—some likely due to data entry in Scopus such as erroneous switch between given names and surnames.
+Some errors likely also stemmed from initials for given names, although we recognise that certain authors might prefer to use initials for publications to avoid clear gender bias.
+Nonetheless, these 59 cases of these errors were excluded from the dataset.
 It is likely that the due to challenge of attributing gender accurately not all the gender attributions are correct.
@@
 
-We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full first name coded with initials, or the second name used as the first name.
-This led to `53` corrections.
+We carried out manual corrections of the gender attributions where we were familiar with the author or the database had the full given name coded with initials, or the second name used as the given name.
+This led to 53 corrections.
+For the purpose of the current study, we inferred authors based on outward gender presentation, though we acknowledge that for future studies, which are not database driven, a more accurate attribution would be to explicitly ask confirmation of such inferences.
 
@@
 These positions carry different academic prestige: first authorship is typically associated with the primary contributor, while last authorship is often held by senior researchers with established reputations [@tscharntke2007author].
+We note that there might be some ambiguity for two-author papers, where the second author could be either someone who contributed less significantly than the publication or a more senior researcher.
+For the scope of the current paper, as it is difficult from names alone to attribute career seniority and it is more common for last authors to be senior, we assume that last position is the more senior position.
 
-For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between male and female authors.
-Our analysis focused on the female author proportion (FAP) and utilised odds ratios (ORs) to compare the likelihood of occupying each authorship position by gender.
+For the analyses, we analysed gender disparities in authorship by comparing the frequency of these authorship types between man and woman authors.
+Our analysis focused on the proportion of woman authors and utilised odds ratios (ORs) to compare the likelihood of occupying each authorship position by gender.
 
@@
 A summary of the authorship types and positions are given in Table 1.
-While the single-authored papers are in minority (8.6%) in this sample overall, there does not seem to be a significant difference between men and women authoring such articles (_p_ = .21); 38.05% of single-authored papers are women with an odds ratio of 0.90 [95% 0.78-1.05] suggesting no clear gender difference despite the figures leaning towards higher frequency of male authors.
+While the single-authored papers are in minority (8.6%) in this sample, there does not seem to be a significant difference between men and women authoring such articles (_p_ = .21); 38.05% of single-authored papers are authored by women with an odds ratio of 0.90 [95% 0.78-1.05] suggesting no clear gender difference despite the figures leaning towards higher frequency of male authors.
 The overall OR for women as first authors was 1.41 [95% CI: 1.28–1.54], indicating that the odds of first authorship are 41% higher for females compared to the male authors (Fisher's exact test _p_ < .001).
 For coauthorship positions, the OR was 1.00 [0.91–1.09], suggesting no substantial gender difference (_p_ = .96).
-In contrast, the odds ratio for women as last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men and significantly different (_p_ < .001).
+In contrast, the odds ratio for women being as the last authors was 0.73 [0.67–0.81], indicating that the odds of last authorship are 27% lower for women compared to men, which was a significant difference (_p_ < .001).
 
@@
 Here we have the maximum of `{r} max(s$n)` authors in a paper and a median of `{r} median(s$n)` (M=`{r} round(mean(s$n),2)`, SD=`{r} round(sd(s$n),2)`).
-Previous research has calculated the collaboration index across gender [@gonzalez2020women], suggesting that male author tend to have larger number of coauthors (M = 1.5) than female (M = 1.34) in psychology.
+Previous research has calculated the collaboration index across gender [@gonzalez2020women], suggesting that male authors tend to have larger number of coauthors (M = 1.5) than female (M = 1.34) in psychology.
 Here the average author numbers for male and female authors (disregarding authorship positions) indicate similar difference, higher mean author numbers for males (M = `{r} round(S$m[2],2)`, [`{r} round(S$LCI[2],2)`-`{r} round(S$UCI[2],2)`]) than for females (M = `{r} round(S$m[1],2)`, [`{r} round(S$LCI[1],2)`-`{r} round(S$UCI[1],2)`]), _t_(4640.58) = 7.21, _p_ < .001, Cohen's _d_ = -0.21, 95% CI
@@
 The results are shown in Figure 2.
-We can observe that while female authors are not more likely to be single authors than men, they tend to be more likely to be first authors than males in papers with 2 or 3 authors; the odds ratios for first authorship position for female authors for papers with 2 authors is 1.55 [130-1.85], and for 3 authors the odds ratio is 1.61 [1.35-1.92], both significantly different from equal proportion.
-The last authorship positions for female authors are consequently lower, odds ratio of 0.65 [0.54–0.77] for 2 authors and 0.59 [0.49–0.71] for 3 authors, both also significant at _p_ < .001 level.
-At four and five authors, the higher odds ratio for female is still evident (OR of 1.37 both both), but the last authorship positions are no longer relative rare for papers with 4 and 5 authors (OR 0.90 and 0.95, respectively).
+We can observe that while women authors are not more likely to be single authors than men, they tend to be more likely to be first authors than men in papers with 2 or 3 authors; the odds ratios for first authorship position for women authors for papers with 2 authors is 1.55 [130-1.85], and for 3 authors the odds ratio is 1.61 [1.35-1.92], both significantly different from equal proportion.
+The last authorship positions for women authors are consequently lower, odds ratio of 0.65 [0.54–0.77] for 2 authors and 0.59 [0.49–0.71] for 3 authors, both also significant at _p_ < .001 level.
+At four and five authors, the higher odds ratio for women is still evident (OR of 1.37 both both), but the last authorship positions are no longer rare for papers with 4 and 5 authors (OR 0.90 and 0.95, respectively).
 At six authors and above, the odds ratios indicate the same trend (higher ratio for first author, lower for last) although due to lower number of observations, these are no longer significantly different.
@@
 
-For comparison, the closest disciplinary benchmark comes from psychology, where @gonzalez2020women analyzed the gender distribution of 74,413 authors publishing in psychology journals in 2009.
+For comparison, the closest disciplinary benchmark comes from psychology, where @gonzalez2020women analysed the gender distribution of 74,413 authors publishing in psychology journals in 2009.
 Women accounted for 45.2% of the sample, indicating near gender parity overall.
@@
 This pattern resembles the trends identified in the present study.
+
+To explore whether men and women authors transition different degrees to last authorship during their career, we analysed authors with at least four publications over a time span of at least seven years (213 in total).
+We compared the most common authorship position during the first and last halves of their careers.
+During the early part of the career, the odds ratio for women being last authors was 0.64 [0.34, 1.21] and during the later part 0.68 [0.39, 1.18].
+Comparing the early and late halves of the careers for men and women, there was a trend for a difference in the transition, but it was not statistically significant ($\chi^2$(1)=3.36, _p_ = 0.067).
 
@@
 Here we tested whether the citations -- as indexed by Scopus -- show differences across gender.
-The median citation for studies with female lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for males, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_ = `{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
-The same comparison of citations for studies with last authors yields similar results `{r} citestats_last$Md[1]` [`{r} citestats_last$CI_lower[1]`–`{r} citestats_last$CI_upper[1]`], and for males, the numbers are similar (Md=`{r} citestats_last$Md[2]` [`{r} citestats_last$CI_lower[2]`–`{r} citestats_last$CI_upper[2]`]).
+The median citation for studies with women lead authors is `{r} citestats_first$Md[1]` [`{r} citestats_first$CI_lower[1]`–`{r} citestats_first$CI_upper[1]`], and for men, the numbers are identical (Md=`{r} citestats_first$Md[2]` [`{r} citestats_first$CI_lower[2]`–`{r} citestats_first$CI_upper[2]`]) and the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_first$statistic,2)`, _p_ = `{r} round(stats_first$p.value,3)`) using rank-based Wilcoxon test.
+The same comparison of citations for studies with last authors yields similar results `{r} citestats_last$Md[1]` [`{r} citestats_last$CI_lower[1]`–`{r} citestats_last$CI_upper[1]`], and for men, the numbers are similar (Md=`{r} citestats_last$Md[2]` [`{r} citestats_last$CI_lower[2]`–`{r} citestats_last$CI_upper[2]`]).
 Again, the difference is not statistically significant ($\chi^2$(1)=`{r} round(stats_last$statistic,2)`, _p_ = `{r} round(stats_last$p.value,3)`).
-If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for female authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for males, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_ < `{r} round(stats_all$p.value,4)`).
+If we assume that we can weight in the gender contribution of all authors in a publication to its citation count, we observe a minor difference, where median citations for women authors is `{r} citestats_all$Md[1]` [`{r} citestats_all$CI_lower[1]`–`{r} citestats_all$CI_upper[1]`], and for men, the central measure are statistically significantly higher (Md=`{r} citestats_all$Md[2]` [`{r} citestats_all$CI_lower[2]`–`{r} citestats_all$CI_upper[2]`], $\chi^2$(1)=`{r} round(stats_all$statistic,2)`, _p_ < `{r} round(stats_all$p.value,4)`).
 For reference, prior study of citation patterns across gender in psychology has reported a small effect favoring men (M = 16.57) over women (M = 15.90) [@gonzalez2020women].
@@
 Out of `{r} as.numeric(sum(t))` articles, `{r} round(as.numeric(t[1]/sum(t))*100,2)`% are Open Access in this sample, as indexed by Scopus.
-For first-authored articles, female odds ratio for Open Access is `{r} round(author_OA$Odds_ratio[1],2)` [`{r} round(author_OA$CI_low[1],2)`–`{r} round(author_OA$CI_high[1],2)`], suggesting nearly 50% higher odds associated with publishing open access by women as compared to men.
+For first-authored articles, odds ratio of Open Access for women is `{r} round(author_OA$Odds_ratio[1],2)` [`{r} round(author_OA$CI_low[1],2)`–`{r} round(author_OA$CI_high[1],2)`], suggesting nearly 50% higher odds associated with publishing open access by women as compared to men.
 If we observe only the last authorship status of the publications, the difference vanishes with the odds ratio indisguishable from even division (1.00 and the confidence interval), OR = `{r} round(author_OA$Odds_ratio[3],2)` [`{r} round(author_OA$CI_low[3],2)`–`{r} round(author_OA$CI_high[3],2)`].
@@
 
-Past studies have identified consistent geographical patterns in female authorship.
+Past studies have identified consistent geographical patterns in woman authorship.
 For example, a large-scale analysis of psychology publications found that 46.5% of authorships were by women, with European countries showing a lower average proportion of 42.8% [@gonzalez2020women].
-To examine geographical differences more closely, we calculated the odds of female authorship by the countries that have published at least 60 publications in the sample, as shown in Figure 3.
+To examine geographical differences more closely, we calculated the odds of woman authorship by the countries that have published at least 60 publications in the sample, as shown in Figure 3.
 
@@
 #| fig.cap: > 
-#|   Female author odds ratio across the prolific countries publishing music psychology (with more than sixty publications).
+#|   Women author odds ratio across the prolific countries publishing music psychology (with more than sixty publications).
 
@@
 
-Panel A in Figure 3 shows a small number of countries, such as Australia, China, and Canada, with a high odds ratio for female authorship.
-At the other end of the spectrum, Japan, Belgium, Austria, Sweden, and Turkey all have significantly lower odds ratios for female authors.
+Panel A in Figure 3 shows a small number of countries, such as Australia, China, and Canada, with a high odds ratio for woman authorship.
+At the other end of the spectrum, Japan, Belgium, Austria, Sweden, and Turkey all have significantly lower odds ratios for women authors.
 There is no simple rationale immediately apparent (e.g., country size, membership in a particular political bloc, gender equality index, or domination of articles by a single eminent scholar from one country) that would explain these differences, but this representation provides some starting points for discussion and may highlight the impact of academic career structures and principles implemented in these countries.
@@
 While this level of presentation may be too coarse for detailed analysis, it does offer a broad summary of the current situation.
-Overall, Oceanie and Americas have higher female author representation, while Asia and Europe have more a male representation.
+Overall, Oceanie and the Americas have higher female author representation, while Asia and Europe have more a male representation.
 
@@
 
-To explore the potential differences within music psychology topics of choice between female and males, we analysed the keywords in the articles (41,438 in total, 5,960 unique).
+To explore the potential differences within topics of choice between men and women, we analysed the keywords in the articles (41,438 in total, 5,960 unique).
 To reduce variant keywords, we manually simplified the variants (e.g., "arts in health", "healthy music use", "health musicking" were converted to "music health" and "music cognition" to "cognition" and "music perception" to "perception") and eliminated single use of the term "music".
-We aggregated the counts of keywords across all authors in the articles and Figure 4 shows the odds ratios for female authors to be associated (in any author position) with the 40 most frequently used keywords.
-The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for females in comparison to overall gender ratios in the data.
-At the opposite end of this continuum, topics such as music analysis, harmony, machine learning, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent male authorships (female OR $\le$ 0.75).
+We aggregated the counts of keywords across all authors in the articles.
+Figure 4 shows the odds ratios for female authors to be associated (in any author position) with the 40 most frequently used keywords.
+The keywords signifying related to children, development, emotion regulation, well-being, music performance anxiety, motivation, expertise, music listening, mental health, and music therapy show significantly higher odds ratios (OR $\ge$ 1.50) for women in comparison to overall gender ratios in the data.
+At the opposite end of this continuum, topics such as music analysis, harmony, machine learning, computation, cognition, and performance expression show the opposite trend, suggesting that the articles with these keywords tend to have overall more prevalent  authorships by men (women OR $\le$ 0.75).
 There is much nuance and detail in the data that would warrant separate analyses; however, the purpose here is simply to provide reference points for comparing gender distributions of authorship with other disciplines, where thematic analyses have already been conducted (e.g., in sub-disciplines of psychology).
@@
 #| fig.cap: > 
-#|   Female author odds ratio across 40 most frequent keywords in all articles as aggregated across all authors.
+#|   Women author odds ratio across 40 most frequent keywords in all articles as aggregated across all authors.
 library(stringr)
@@
 An analysis of papers from five core music psychology journals across 25 years revealed that, overall, women represented 40.2% of authors.
-This proportion is higher than in musicology [34.8% in @eerola2026inprep], in psychology more broadly [33.2% in @huang2020; 36.2% in @gonzalez2020women], and also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
+This proportion is higher than in musicology (34.8% in @eerola2026inprep], in psychology more broadly [33.2% in @huang2020; 36.2% in @gonzalez2020women], and also much higher than in other STEM subjects of math, physics and computer science [15% in @huang2020] and topics such as astronomy and bioscience [17.8% in @son2022scientific].
 Although our current findings suggests that the field of music psychology is doing better in fostering gender equity than some other fields, the distribution of authors by gender remains nonetheless unequal.
@@
 Women are more likely than men to appear as first authors, a trend that has become characteristic of the field only in the past 15 years.
-This is in line with previous research showing that female authorship representation is increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
+This is in line with previous research showing that woman authorship representation is increasing [@gonzalez2020women; @ioannidis2023gender; @jemielniak2025gender; @sanchez2024analysis].
 In contrast, we found that women remain underrepresented in the last authorship position, and this trend has not changed substantially over the last 15 years.
-This underrepresentation in woman as last authors is similar to the situation in medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
+Interestingly, this changes depending on the number of authors on a paper overall, perhaps suggesting a greater ambiguity for the last author role for papers with fewer authors.
+Nonetheless, this lower representation of woman as last authors is similar to the situation in medicine [@bruck2023bibliometric; @shah2021gender], psychology [@gonzalez2020women], and biology [@rock2021quantifying].
 Although last authorship in the social sciences is not necessarily universally defined, it is nonetheless increasingly associated with seniority, supervisory responsibility, or principal investigator status [@drivas2024evolution].
@@
 Indeed, this gender gap could come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement [@nielsen2024getting].
-Fortunately, there are strides to improve the current situation, with opinion papers discussing way to improve receipt and reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
+Fortunately, there are strides to improve the current situation, with opinion papers discussing ways to improve reporting of intellectual credit [@banks2025women] as well as creating equitable environments in academic science [@martinez2024comparison].
 In particular, one crucial aspect that has been highlighted to is the need for more mentorship programs for (junior) women to senior positions.
@@
 We note that within the broader music psychology community, there are already mentorship programs in place: for example, Women in Music Information Retrieval WIMIR (https://wimir.wordpress.com/) and the Society for Music Perception and Cognition (SMPC) has resources for trainees and a Mentorship committee (https://www.musicperception.org/committees) and often holds mentor-mentee meetings at conferences.
-Although these measures have been implemented for some years, it might still take some time for the effect to be reflected in authorship.
+Although these measures have been implemented for some years, it might still take some time for the effect to be reflected in authorship roles.
 
@@
 Although the odds ratio for women publishing open access is highest in _Music and Science_, the broader trend suggests a temporal shift: since around 2010, women have increasingly overtaken men in publishing open access.
-The data does not provide any direct reasons for this, although the legislation in few of the European nations (United Kingdom, EU) require open access publishing and when women have higher odds ratio for being first authors, this might be due to the high number of female from these countries.
+The data does not provide any direct reasons for this, although the legislation in few of the European nations (United Kingdom, EU) require open access publishing and when women have higher odds ratio for being first authors the high number of women represented in publications these countries.
 While past surveys among, for instance, UK academics have not indicated substantial differences in experiences with open access [@zhu2017support], financial and other inequalities in emerging countries have been suggested to limit women's capacity to publish in open access journals [@vuong2021adopting].
-The idea that publishing in a more open and transparent way benefit society in general [@pervsic2023open], and this fits with the general idea that female researchers value research on societal progress, while males prefer research aimed at scientific progress [@zhang2021gender].
-However, nothing in the extant literature would point out to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
+The idea that publishing in a more open and transparent way benefit society in general [@pervsic2023open], and this fits with the general idea that women researchers value research on societal progress [@zhang2021gender].
+However, nothing in the extant literature would point to the reverse pattern, which requires further investigation about the reasons for adopting open access publishing models.
 
-In terms of eographical and cultural factors, our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequity may stem from national cultural factors [@bruck2023bibliometric].
-Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide a academic climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
+In terms of geographic and cultural factors, our summaries of gender patterns across continents and countries indicate clear cross-national variation, supporting the idea that certain inequities may stem from national cultural factors [@bruck2023bibliometric].
+Large Commonwealth nations, namely Australia, Canada, and the United Kingdom, seem to provide an academic climate more supportive of women in research, as reflected in the relatively high odds ratios of women appearing as authors (in any authorship position) in music psychology.
 This pattern, however, is not universal across all countries examined [@chan2020].
@@
 This fits with another potential explanation for the disparity between authorship, namely that women faculty lean word and perform more university, campus, and departmental services than men in general [@guarino2017faculty].
-Over time, performing such services to universities are time away from other work on acquiring funding and leading research, which would aggregate over time to time away from publishing research.
-It is nonetheless important for such services are completed; in some cases it could be there is more encouragement for males to take on these kinds of roles, but also it could be that success in academia could also be more societal defined as not only publications [@kiai2019protect], but as a whole should celebrate achievements that support internal, local and community services in a broader sense.
+Performing such services to universities are time away from other work on acquiring funding and leading research, which would aggregate over time to time away from publishing research.
+It is nonetheless important that services be completed; in some cases it could be there is more encouragement for men to take on these kinds of roles, but also it could be that success in academia could also be more societally defined as not only publications [@kiai2019protect], but as celebrating achievements that support internal, local, and community services in a broader sense.
 
@@
 Nonetheless, we maintain that the current conclusions may still hold as both authors checked whether any of the gender attribution of major names needed fixing (see accompanying code).
-Additionally, gender attribution error was relatively small (only `89` had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
+Additionally, gender attribution error was relatively small (only 89 had low (< 0.55) attribution probability, while 89.3% had high (> 0.90) attribution probability), so it is unlikely that the remaining names would significantly change the main conclusions.
 A second limitation is treating gender as a binary category.
-While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.8% in @son2022scientific].
+While we recognise that gender is a complex and multidimensional social construct, we used a binary male-female classification consistent with prior meta-science research on gender and authorship [@son2022scientific; @shah2021gender; @rock2021quantifying; @ni2021gendered; @gonzalez2020women; @wais2006] and because generally the number of non-binary authors is currently relatively low [0.01% in @son2022scientific].
 Nonetheless, future research should also move towards using more inclusive, non-binary information.
@@
 
-There was small but significant effect of gender on citations that favoured men when the citations were aggregated all authors in the papers.
-There was no, however, differences in citations for female first or last authors.
+There was a small but significant effect of gender on citations that favoured men when the citations were aggregated to include all authors in the papers.
+There were no differences, however, in citations for women first or last authors.
 The aggregated results are in line with previous findings in psychology that favored men [@gonzalez2020women].
-While the present analysis cannot determine the underlying causes of the observed trends, we can point to potential factors that may help explain why this discipline appears to be more equitably distributed by gender for the first authors than, for example, psychology.
-Female role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for _Musicae Scientiae_ 1997-2009).
-However, this does not fully explain the rise of female first authors over the past 15 years, nor is it consistent with the reverse pattern observed in the last authorship position.
+While the present analysis cannot determine the underlying causes of the observed trends, we can point to potential factors that may help explain why this discipline appears to be more equitably distributed by gender for the first authors than, for example, the field of general psychology.
+Women role models (past and present) exist in music psychology; many of the founding members of the societies have been women (e.g., Diana Deutsch for ICMPC in 1989, Irène Deliège for ESCOM in 1991, and similarly many founding editors of the journals have been women (e.g., Diana Deutsch for _Music Perception_ 1983-1995, and Irène Deliège for _Musicae Scientiae_ 1997-2009).
+Others have also pioneering certain aspects in the music psychology field, such as music and developmental (Laurel Trainor, Sandra Trehub), musical movement and performance (Jane Davidson, Caroline Palmer), and general music perception (Carol Krumhansl).
+However, this does not fully explain the rise of women first authors over the past 15 years, nor is it consistent with the reverse pattern observed in the last authorship position.
 A cautious interpretation is that the increase in first authorship may signal the discipline becoming more gender-balanced, but this shift has not yet extended to last authorship positions, which still reflect either career longevity or barriers to women attaining senior academic positions.
 
-While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways into obtaining insights into the gendered career progression in music psychology; for instance, future work with the same data could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
+While we present analysis of gender of the authorships across the discipline-specific journal in a relatively simple manner here, there would be more elaborate ways of obtaining insights into the gendered career progression in music psychology; for instance, future work with the same data could analyse the precarity of career options as judged from the years of authorship attributions [see @lundine2018].
 One could also analyse the last author positions in more detail using direct contacts to the authors (interviews and surveys) and try to identify the factors contributing to the barriers experienced by women to take up senior author positions.
+We also note that equity might not necessarily come from 50-50 representation in authors necessarily; there would be slight variations around this number, and perhaps there are also other ways to better define equity in the workplace that capture more equal opportunities in the workplace.
 
-Current findings from authorship counts in music psychology present a more optimistic outlook of female authors in comparison what the prevalence of female academics or professors in academia in general are [@bhopal2021competing; @harris2025gender; @Filandri03082021].
+Current findings from authorship counts in music psychology present a more optimistic outlook of women authors in comparison with the prevalence of women academics or professors in academia in general [@bhopal2021competing; @harris2025gender; @Filandri03082021].
 While the results are broadly comparable to psychology as a whole, musicology and its closely related disciplines such as music theory, popular music studies, and ethnomusicology tend to show less balanced gender representation in overall authorship.
-Gender, however, is only one dimension of equity; racial, ethnic, geographical, institutional, and other inequalities also merit attention in music psychology.
-Although women remain underrepresented in terms of absolute authorship counts in specialist music psychology journals, even a simple tally of contributions can help open discussion and provide evidence of progress toward greater inclusion and equality.
+Gender, however, is only one dimension of equity; racial, ethnic, geographical, institutional, and other inequalities and their intersections also merit attention in music psychology.
+Although women have a lower representation in terms of absolute authorship counts in specialist music psychology journals, even a simple tally of contributions can help open discussion and provide evidence of progress toward greater inclusion and equality.
 To interpret these patterns more fully, one must examine how music psychology operates through its journals, conferences, training programmes, and curricula, as well as how the discipline presents itself both in countries with a long-standing tradition in the field and in regions where it has yet to become firmly established.
@@
 
-Study data, analysis scripts and supporting information is available at GitHub, <https://tuomaseerola.github.io/gender_in_music_psych>.
+Study data, analysis scripts and supporting information is available at GitHub, [https://anonymous.4open.science/status/gender_in_music_psych-C6FE](https://anonymous.4open.science/status/gender_in_music_psych-C6FE).
 
@@
 We are grateful to Isabelle Peretz for suggesting a deeper look into the author types, and to our colleagues for their valuable input on the manuscript.
+We also thank Samuel P.
+Horlor, Alex Harrington, Alvaro Chang-Arana, and Emma Allingham for comments on a previous manuscript version.
 

### 2025-12-12 manuscript/manuscript.qmd
First submission

@@
 
-abstract: "Academia suffers from various issues regarding equity.
-One such issue is gender inequality in authorship, where many studies show that women are often underrepresented.
-Due to authorship differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authorship within music psychology.
-A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analyzed with respect to author gender and authorship role (first, last, solo, and coauthor).
+abstract: "Academia suffers from various equity issues.
+One such issue is gender inequality in authorship, where many studies show that women are less represented compared to men.
+Due to authorship differences across disciplines, the goal of this study is to provide a current overview of the gender distribution of authorship within music psychology.
+A total of 3,373 papers published between 2000 and 2025 across five core music psychology journals were analysed with respect to author gender and authorship role (first, last, solo, and coauthor).
 In addition, we explored gender patterns related to citations, open-access publishing, and keywords.
-Overall, female authors were in the minority, accounting for 40.2% of all authors.
-However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
+Overall, women authors were in the minority, accounting for 40.2% of all authors.
+However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (odds ratio, OR = 1.41), equally likely to be solo authors (OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73).
 These patterns have become more pronounced over the past 15 years.

### 2025-12-12 contents.R

@@ -34,3 +34,5 @@ source('scripts/visualise_gender.R') # Which representation we prefer?
 
 #source('scripts/visualise_names.R') # Bonus feature
 source('scripts/keyword_analysis.R') # creates 3 alternative figures
+
+# extra analysis: career changes in authorship positions

### 2025-12-12 scripts/career_publishing_changes.R

@@ -1,0 +1,107 @@
+# extra analysis: career changes in authorship positions
+source('scripts/load_gender_data.R') # Loads the data
+
+# take all unique author names
+U<-unique(df$unique_name)
+length(U) # 5312
+
+# minimun number of publications
+min_pubs <- 4
+
+# minim number of years between first and last publication
+min_years <- 7
+
+# create a data frame to store results
+career_changes<-NULL
+# loop through each unique author
+for (author in U) {
+  # subset data for the author
+  author_data <- df %>% filter(unique_name == author)
+  num_pubs <- nrow(author_data)
+  if (num_pubs >= min_pubs) {
+    first_year <- min(author_data$YEAR)
+    last_year <- max(author_data$YEAR)
+    if ((last_year - first_year) >= min_years) {
+      year0_position <- data.frame(year0=author_data$YEAR-first_year, author_order=author_data$author_order,authortype=author_data$authortype,Gender=author_data$Gender[1], stringsAsFactors=FALSE)
+      year0_position <- dplyr::arrange(year0_position, year0)
+      year0_position$authortype[year0_position$author_order==1]<-"First"
+     # year0_position$authortype[year0_position$author_order==2]<-"Other" # extra
+      year0_position
+      early<-filter(year0_position, year0 <= max(year0_position$year0)/2)
+      late<-filter(year0_position, year0 > max(year0_position$year0)/2)
+      early_counts <- table(early$authortype)
+      early_counts_max <- names(which.max(early_counts))
+      late_counts <- table(late$authortype)
+      late_counts_max <- names(which.max(late_counts))
+
+      career_changes <- rbind(career_changes, data.frame(author=author,
+                                                         gender=author_data$Gender[1],
+                                                           first_year=first_year,
+                                                           last_year=last_year,
+                                                           num_pubs=num_pubs,
+                                                           early_type_max=early_counts_max,
+                                                           late_type_max=late_counts_max,
+                                                           early_order_mean=mean(early$author_order),
+                                                           late_order_mean=mean(late$author_order),
+                                                         stringsAsFactors=FALSE))
+    }
+  }
+}
+
+dim(career_changes)
[…]

### 2025-12-12 scripts/country_table.R

@@ -83,7 +83,7 @@ g2<-ggplot(DATA,aes(x=reorder(Continent,Odds_ratio),y=Odds_ratio,label=paste0("n
   geom_hline(yintercept = 1,linetype='dashed',color='grey20')+
   scale_y_continuous(limits=c(0,2))+
   xlab('Continent') +
-  ylab('Female Authorship Odds Ratio (95%CI)') +
+  ylab('Women Authorship Odds Ratio (95%CI)') +
   coord_flip()
 #print(g2)
 

### 2025-12-12 scripts/keyword_analysis_dark.R

@@ -1385,21 +1385,21 @@ figure3 <- ggplot(data = x1[FROM:TO, ], aes(
   y = or,
   label = paste0("n=",Freq)
 )) +
-  geom_col(fill = 'grey80', color = 'grey10') +
-  geom_errorbar(aes(ymin=or_LCI, ymax=or_UCI), width=.2, alpha=0.6,color="grey40",linetype="solid") +
-  geom_text(y=.1,nudge_y = -.0997, size=2.25) +
-  theme(text = element_text(size = 16)) +
+  geom_col(fill = 'lightblue4', color = 'grey10') +
+  geom_errorbar(aes(ymin=or_LCI, ymax=or_UCI), width=.2, alpha=0.6,color="white",linetype="solid") +
+  geom_text(y=.1,nudge_y = -.0997, size=2.25,color='yellow2') +
+  theme(text = element_text(size = 14)) +
   scale_y_continuous(
-    breaks = seq(0, 3.78, .25),
+    breaks = seq(0, 3.78, .50),
     limits = c(0,3.78),
     expand = c(0.001,0.01)
   ) +
   scale_fill_grey() +
   coord_flip(ylim = c(0,3.78)) +
-  ylab("Women Author Odds Ratio (all authors)") +
+  ylab("OR Female (all authors)") +
   xlab("Keyword (ranked)") +
-  geom_hline(yintercept = 1.0, linetype = "dashed", color = "grey39") +
-  theme_classic(base_size = 15) +
+  geom_hline(yintercept = 1.0, linetype = "dashed", color = "yellow3") +
+  theme_classic(base_size = 14) +
   theme(
     panel.border = element_blank(),
     panel.grid.major = element_blank(),
@@ -1407,7 +1407,27 @@ figure3 <- ggplot(data = x1[FROM:TO, ], aes(
     axis.line = element_line(colour = "black")
   )
 
-print(figure3)
+library(ggdark)
+figure3 <- figure3 + dark_theme_bw(base_size = 16)
+
+# make transparent
+
+figure3<-figure3 + theme(
+  panel.background = element_rect(fill='transparent'),
+  plot.background = element_rect(fill='transparent', color=NA),
+  panel.grid.major = element_blank(),
+  panel.grid.minor = element_blank(),
+  legend.background = element_rect(fill='transparent',color=NA),
+  legend.box.background = element_rect(fill='transparent',color=NA)
+)
+
[…]
